\documentclass[12pt]{article}
\input{doc/header/header}
\input{doc/header/defs}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
% \usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage[margin=2.5cm,nohead]{geometry}
\usepackage{authblk}
\newcommand{\email}[1]{\href{mailto:#1}{#1}}

\title{RtEstim: Effective reproduction number estimation with trend filtering}

\author[a,1]{Jiaping Liu}
\author[b]{Zhenglun Cai}
\author[a]{Paul Gustafson}
\author[a]{Daniel J. McDonald}

\affil[a]{Department of Statistics, The University of British Columbia}
\affil[b]{Centre for Health Evaluation and Outcome Sciences, The University of
British Columbia}

\begin{document}

\maketitle

\begin{abstract}
  To understand the transmissibility and spread of infectious diseases,
  epidemiologists turn to estimates of the effective reproduction number.
  While many approaches exist, their utility may be limited by the
  challenges of surveillance data collection. Arbitrary model assumptions
  that are unverifiable with data alone and sophisticated, though
  computationally inefficient, frameworks that
  are critical limitations for many
  existing approaches. We propose a discrete spline-based approach 
  \RtEstim\ that solves a convex optimization problem --- Poisson trend filtering
  --- using the proximal Newton method. 
  It produces a locally adaptive estimator for effective reproduction number 
  estimation with heterogeneous smoothness. 
  \RtEstim\ remains accurate even under some process misspecifications and
  is computationally efficient, even for large-scale data. The 
  implementation is easily accessible in a lightweight \texttt{R} package
  \href{https://dajmcdon.github.io/rtestim/index.html}{\texttt{rtestim}}.

  {\bf Keywords:} tuning parameter selection $|$ heterogeneous smoothness
\end{abstract}

\renewcommand{\abstractname}{Author summary}
\begin{abstract}
  Effective reproduction number estimation faces many challenges, for example, 
  data is usually collected from multiple sources and there are reporting delays, that cause many problems, 
  such as missing values, underreporting and lack of standardization. 
  Such poor conditions of data hinder the accurate estimation of effective reproduction number.  
  Our motivation is to develop a model that returns accurate estimation, which is robust to violation of model misspecification, and a tool that is straightforward to use and computationally efficient, even for large-scale data. 
  Considering the temporal evolution of effective reproduction number for multiple degrees, we propose a convex optimization model with an $\ell_1$ trend filtering penalty for univariate data. We solve our model using the proximal Newton method, which converges rapidly and is numerically stable. 
  Our software, \texttt{R} package \RtEstim, can return the output in seconds for univariate incidence sequences with length 300 using candidate sets of length fifty and ten-fold cross validation. 
\end{abstract}

\footnotetext[1]{To whom correspondence should be
  addressed. E-mail: \email{jiaping.liu@stat.ubc.ca}} 
% Technical summary including research objective, method, results, and conclusions.

%\thispagestyle{empty}
%\pagenumbering{roman}
%\setcounter{page}{1}

%\printnotation
%\input{doc/notations.tex}
%\newpage
%\hypersetup{linkcolor=black}
\tableofcontents % for easy edits, remove later

%\listoffigures

\clearpage
%\pagenumbering{arabic}
%\setcounter{page}{1}

% research proposal
\input{doc/introduction}  
\input{doc/methods}
\input{doc/results}
\input{doc/discussion}

%\section*{Acknowledgement}


\addcontentsline{toc}{section}{References}

\bibliographystyle{doc/rss.bst}
\bibliography{doc/ptf}

%\appendix
%\titleformat{\chapter}{\normalfont\huge\bfseries}{Appendix \thechapter}{1em}{}

\end{document}
