\begin{thebibliography}{10}

\bibitem{diekmann1990definition}
Diekmann O, Heesterbeek JAP, Metz JA.
\newblock On the definition and the computation of the basic reproduction ratio
  R 0 in models for infectious diseases in heterogeneous populations.
\newblock Journal of mathematical biology. 1990;28:365--382.

\bibitem{dietz1993estimation}
Dietz K.
\newblock The estimation of the basic reproduction number for infectious
  diseases.
\newblock Statistical methods in medical research. 1993;2(1):23--41.

\bibitem{fine1993herd}
Fine PE.
\newblock Herd immunity: history, theory, practice.
\newblock Epidemiologic reviews. 1993;15(2):265--302.

\bibitem{delamater2019complexity}
Delamater PL, Street EJ, Leslie TF, Yang YT, Jacobsen KH.
\newblock Complexity of the basic reproduction number (R0).
\newblock Emerging infectious diseases. 2019;25(1):1.

\bibitem{brauer2019mathematical}
Brauer F, Castillo-Chavez C, Feng Z.
\newblock Mathematical models in epidemiology. vol.~32.
\newblock Springer; 2019.

\bibitem{anderson1982directly}
Anderson RM, May RM.
\newblock Directly transmitted infections diseases: control by vaccination.
\newblock Science. 1982;215(4536):1053--1060.

\bibitem{anderson1985vaccination}
Anderson RM, May RM.
\newblock Vaccination and herd immunity to infectious diseases.
\newblock Nature. 1985;318(6044):323--329.

\bibitem{heffernan2005perspectives}
Heffernan JM, Smith RJ, Wahl LM.
\newblock Perspectives on the basic reproductive ratio.
\newblock Journal of the Royal Society Interface. 2005;2(4):281--293.

\bibitem{cori2013new}
Cori A, Ferguson NM, Fraser C, Cauchemez S.
\newblock A new framework and software to estimate time-varying reproduction
  numbers during epidemics.
\newblock American journal of epidemiology. 2013;178(9):1505--1512.

\bibitem{abry2020spatial}
Abry P, Pustelnik N, Roux S, Jensen P, Flandrin P, Gribonval R, et~al.
\newblock Spatial and temporal regularization to estimate COVID-19 reproduction
  number R (t): Promoting piecewise smoothness via convex optimization.
\newblock Plos one. 2020;15(8):e0237901.

\bibitem{pascal2022nonsmooth}
Pascal B, Abry P, Pustelnik N, Roux S, Gribonval R, Flandrin P.
\newblock Nonsmooth convex optimization to estimate the Covid-19 reproduction
  number space-time evolution with robustness against low quality data.
\newblock IEEE Transactions on Signal Processing. 2022;70:2859--2868.

\bibitem{thompson2019improved}
Thompson RN, Stockwin JE, van Gaalen RD, Polonsky JA, Kamvar ZN, Demarsh PA,
  et~al.
\newblock Improved inference of time-varying reproduction numbers during
  infectious disease outbreaks.
\newblock Epidemics. 2019;29:100356.

\bibitem{ramdas2016fast}
Ramdas A, Tibshirani RJ.
\newblock Fast and flexible admm algorithms for trend filtering.
\newblock Journal of Computational and Graphical Statistics.
  2016;25(3):839--858.

\end{thebibliography}
