---
title: "Supplementary details on experiments"
author: "Olivia"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rtestim)
library(EpiEstim)
library(data.table)
library(tidyverse)
```

# Experimental design

## Synthetic data with Poisson incidence

Overall, we argue our estimator is accurate, robust in model misspecification and computationally efficient. We can do a series of tests for each property. We may consider the following curvature of efficient reproduction numbers, and test the accuracy of our estimators compared to EpiEstim and EpiLPS.

We may consider arbitrary reproduction numbers in a few scenarios: a) piecewise-constant epidemics with a drop at a certain time point to measure the effect of control measures, b) exponentially rising and falling epidemics with a change point, c) piecewise-constant with multiple segments to measure the initially controlled and resurged and the suppressed epidemics, d) periodic waves.

We may simulate the epidemics (with length T=300) 10 times for each scenario, estimate Rt, and compute the averaged KL divergence.

```{r evenly-spaced-Rt-functions}
# General settings: 
N1 = 2 # first incidence data
len = 300 # number of evenly spaced time points
library(rtestim)
# Get Poisson incidence cases: 
get_pois_incidence <- function(N1, Rt, gamma_pars = c(2.5, 2.5)){
  len <- length(Rt)
  incidence <- numeric(len) # N_1:n
  poisson_count <- numeric(len) # y_1:n
  incidence[1] <- N1
  
  poisson_count[1] <- rpois(1, N1)
  if(poisson_count[1] == 0) poisson_count[1]=1
  for(t in 2:len){
    pi <- discretize_gamma(1:(t-1), gamma_pars[1], gamma_pars[2])
    incidence[t] <- Rt[t] * sum(rev(pi) * poisson_count[1:(t-1)])
    poisson_count[t] <- rpois(1, incidence[t])
  }
  
  return(poisson_count)
}
# Display the synthetic data: 
library(ggplot2)
display_dat <- function(counts, Rt){
  len <- length(counts)
  if(length(counts) != length(Rt)) cli::cli_abort("Data lengths do not match.")
  dat <- data.frame(time = 1:len, count = counts, Rt = Rt)
  fig1 <- dat %>%
    ggplot(aes(y = Rt, x = time)) +
    geom_line() +
    theme_bw()
  print(fig1)
  fig2 <- dat %>%
    ggplot(aes(y = count, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    ylab("Poisson incidence data") + #(in log scale) 
    theme_bw()
  print(fig2)
}
# Check data quality:
check_dat <- function(incidence, Rt){
  if(min(Rt) < 0) {cli::cli_abort("`Rt` must be non-negative.")}
  if(min(incidence) < 0) {cli::cli_abort("`incidence` cases must be nonnegative.")}
  if(max(incidence) > 1e4L) {cli::cli_alert_warning("`incidence` cases are too large.")}
  if(sum(incidence == 0) > 30) {cli::cli_warn("`incidence` data has more than 10% 0s.")}
}
```

```{r Rt-Pois-case1}
# Scenario 1: two-stage piecewise constant with one dropping point (similar as in EpiFilter)
Rt1 <- c(rep(2, 70), rep(0.8, len-70)) # extend first segment from 50 to 70
# to avoid too many 0s in the tail
gamma_pars1 <- c(3, 3) # serial interval distribution parameters
seed <- 420
set.seed(seed)
incidence1 <- get_pois_incidence(N1, Rt1, gamma_pars1)
display_dat(incidence1, Rt1)
check_dat(incidence1, Rt1)
```

```{r Rt-Pois-case2}
# Scenario 2: two-stage exponential growth and decay (similar as in EpiFilter)
rate <- c(.015, -.005)
Rt2 <- numeric(len)
knot <- 50
Rt2[1:knot] <- exp(rate[1] * (1:knot)) 
Rt2[(knot+1):len] <- exp((rate[2]) * ((knot+1):len))* Rt2[knot]
gamma_pars2 <- c(2.5, 2.5) # serial interval distribution parameters
set.seed(seed)
incidence2 <- get_pois_incidence(N1, Rt2, gamma_pars2)
display_dat(incidence2, Rt2)
check_dat(incidence2, Rt2)
```

```{r Rt-Pois-case3}
# Scenario 3: multi-stage piecewise constant (similar as in EpiFilter)
Rt3 <- c(seq(2.5, 2, length.out = 60), seq(0.8, 0.6, length.out = 50), 
         seq(1.7, 2, length.out = 40), seq(0.9, 0.5, length.out = 150))
gamma_pars3 <- c(3.5, 3.5) # serial interval distribution parameters
set.seed(seed)
incidence3 <- get_pois_incidence(N1, Rt3, gamma_pars3)
display_dat(incidence3, Rt3) 
check_dat(incidence3, Rt3)
```

```{r Rt-Pois-case4}
# Scenario 4: multi-stage with exponential-sinusoidal growth and decay
x <- seq(0, 10, length.out = len)
Rt4 <- numeric(len)
components <- list(
  list(freq = 0.1, amp = 1),
  list(freq = 0.5, amp = 2),
  list(freq = 1.0, amp = 3)
)
for (component in components) {
  Rt4 <- Rt4 + 0.2 * (component$amp * sin(pi * component$freq * x / 1.2) + 
                        component$amp)
}
gamma_pars4 <- c(3.5, 3.5) # serial interval distribution parameters
set.seed(seed)
incidence4 <- get_pois_incidence(N1, Rt4, gamma_pars4)
display_dat(incidence4, Rt4)
check_dat(incidence4, Rt4)
```

Parameters of the serial interval distribution, i.e., shapes and scale/rates of Gamma distribution, can significantly influence the peak values of incidences and the smoothness of incidence curves. Here is a comparison of the densities of Gamma distribution with shape 2.5 and scale 2.5 and Gamma distribution with shape 5 and scale 5.
```{r Serial-Interval-density}
par(mfrow = c(1, 3))
plot(dgamma(1:50, 2.5, scale = 2.5), type="l", ylab="", xlab="", 
     main="Gamma(2.5, scale=2.5)")
  
plot(dgamma(1:50, 3, scale = 3), type="l", ylab="", xlab="", 
     main="Gamma(3, scale=3)")

plot(dgamma(1:50, 3.5, scale = 3.5), type="l", ylab="", xlab="", 
     main="Gamma(3.5, scale=3.5)")
par(mfrow = c(1, 1))
```

### Rt estimates using three methods
Fit our poisson trend filtering using `RtEstim`. See an example of the first
scenario. We first see an demonstration of estimates using 50 hyperparameters, 
and use cross validation to choose the "best" hyperparameter with the lowest 
score. The last plot of this scenario displays the Rt estimates using the chosen
hyperparameter. 

```{r Rt-Pois-RtEstim-case1}
rtestim_mod1 <- estimate_rt(incidence1, korder = 0, nsol = 50, 
                            dist_gamma = gamma_pars1)
sum(rtestim_mod1$convergence == FALSE) == 0
plot(rtestim_mod1)

# hyperparameter tuning using cross validation: 
cv_mod1 <- cv_estimate_rt(incidence1, korder = 0, nfold = 3, nsol = 50, 
                          dist_gamma = gamma_pars1, maxiter=1e7L)
plot(cv_mod1)

rtestim_tuned_mod1 <- cv_mod1$full_fit$Rt[, which.min(cv_mod1$cv_scores)]
rt_ci1 <- confband(cv_mod1, "lambda.min") # get 95% confidence band
rt_ci1 %>%
  ggplot(aes(x = 1:length(incidence1), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt with 95% CIs") + 
  theme_bw()
```

The Rt estimates for the other three scenarios are given in the following figures.

```{r Rt-Pois-RtEstim}
rtestim_mod2 <- estimate_rt(incidence2, korder=3, nsol=50, maxiter=1e7L, 
                            dist_gamma = gamma_pars2)
sum(rtestim_mod2$convergence == FALSE) == 0
plot(rtestim_mod2)
cv_mod2 <- cv_estimate_rt(incidence2, korder=3, nfold=10, nsol=50,  maxiter=1e7L, 
                          dist_gamma = gamma_pars2)
plot(cv_mod2)
rtestim_tuned_mod2 <- cv_mod2$full_fit$Rt[, which.min(cv_mod2$cv_scores)]
rt_ci2 <- confband(cv_mod2, "lambda.min") # get 95% confidence band
rt_ci2 %>%
  ggplot(aes(x = 1:length(incidence2), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt with 95% CIs") + 
  theme_bw()

rtestim_mod3 <- estimate_rt(incidence3, korder=1, nsol=50, maxiter=1e7L, 
                            dist_gamma = gamma_pars3)
sum(rtestim_mod3$convergence == FALSE) == 0
plot(rtestim_mod3)
cv_mod3 <- cv_estimate_rt(incidence3, korder=1, nfold=3, nsol=50, maxiter=1e7L, 
                          dist_gamma = gamma_pars3)
rtestim_tuned_mod3 <- cv_mod3$full_fit$Rt[, which.min(cv_mod3$cv_scores)]
rt_ci3 <- confband(cv_mod3, "lambda.min") # get 95% confidence band
rt_ci3 %>%
  ggplot(aes(x = 1:length(incidence2), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt with 95% CIs") + 
  theme_bw()

rtestim_mod4 <- estimate_rt(incidence4, korder=3, nsol=50, maxiter=1e7L, 
                            dist_gamma = gamma_pars4)
sum(rtestim_mod4$convergence == FALSE) == 0 # check the convergence
plot(rtestim_mod4)
cv_mod4 <- cv_estimate_rt(incidence4, korder=3, nfold=3, nsol=50, maxiter=1e7L, 
                          dist_gamma = gamma_pars4)
rtestim_tuned_mod4 <- cv_mod4$full_fit$Rt[, which.min(cv_mod4$cv_scores)]
rt_ci4 <- confband(cv_mod4, "lambda.min") # get 95% confidence band
rt_ci4 %>%
  ggplot(aes(x = 1:length(incidence2), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt with 95% CIs") + 
  theme_bw()
```

Fit EpiEstim with "true" serial interval.

```{r message=FALSE, Rt-Pois-EpiEstim}
library(EpiEstim)
method <- "non_parametric_si"
# Get "true" serial interval: get gamma probabilities with length 12; the 1st component is 0
Dmax1 <- floor(stats::qgamma(0.9999, 3, scale = 3)); Dmax1 < 250
prob_gamma1 <- discretize_gamma(1:Dmax1, shape = 3, scale = 3)

Dmax2 <- floor(stats::qgamma(0.9999, 2.5, scale = 2.5)); Dmax2 < 250
prob_gamma2 <- discretize_gamma(1:Dmax2, shape = 2.5, scale = 2.5)

Dmax3 <- floor(stats::qgamma(0.9999, 3.5, scale = 3.5)); Dmax3 < 250
prob_gamma3 <- discretize_gamma(1:Dmax3, shape = 3.5, scale = 3.5)
prob_gamma4 <- prob_gamma3 # set up for EpiLPS

# weekly sliding windows
t_week_start <- seq(2, len - 6) # starting at 2 as conditional on the past observations
t_week_end <- t_week_start + 6
            
config_week1 <- make_config(list(si_distr = c(0, prob_gamma1),
                                 t_start = t_week_start,
                                 t_end = t_week_end)) 
config_week2 <- make_config(list(si_distr = c(0, prob_gamma2),
                                 t_start = t_week_start,
                                 t_end = t_week_end)) 
config_week3 <- make_config(list(si_distr = c(0, prob_gamma3),
                                 t_start = t_week_start,
                                 t_end = t_week_end)) 
config_week4 <- config_week3

mod_epiEstim_week1 <- EpiEstim::estimate_R(incid = incidence1, 
                                      config = config_week1, 
                                      method = method)
plot(mod_epiEstim_week1, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_week2 <- EpiEstim::estimate_R(incid = incidence2, 
                                      config = config_week2, 
                                      method = method)
plot(mod_epiEstim_week2, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_week3 <- EpiEstim::estimate_R(incid = incidence3, 
                                      config = config_week3, 
                                      method = method)
plot(mod_epiEstim_week3, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_week4 <- EpiEstim::estimate_R(incid = incidence4, 
                                      config = config_week4, 
                                      method = method)
plot(mod_epiEstim_week4, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

# monthly sliding window
t_month_start <- seq(2, len - 29) # starting at 2 as conditional on the past observations
t_month_end <- t_month_start + 29
config_month1 <- make_config(list(si_distr = c(0, prob_gamma1),
                                  t_start = t_month_start,
                                  t_end = t_month_end)) 
config_month2 <- make_config(list(si_distr = c(0, prob_gamma2),
                                  t_start = t_month_start,
                                  t_end = t_month_end)) 
config_month3 <- make_config(list(si_distr = c(0, prob_gamma3), 
                                  t_start = t_month_start,
                                  t_end = t_month_end)) 
config_month4 <- config_month3

mod_epiEstim_month1 <- EpiEstim::estimate_R(incid = incidence1, 
                                            config = config_month1, 
                                            method = method)
plot(mod_epiEstim_month1, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_month2 <- EpiEstim::estimate_R(incid = incidence2, 
                                            config = config_month2, 
                                            method = method)
plot(mod_epiEstim_month2, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_month3 <- EpiEstim::estimate_R(incid = incidence3, 
                                            config = config_month3, 
                                            method = method)
plot(mod_epiEstim_month3, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")

mod_epiEstim_month4 <- EpiEstim::estimate_R(incid = incidence4, 
                                            config = config_month4, 
                                            method = method)
plot(mod_epiEstim_month4, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window and 95% CIs", title="") +
  theme_bw() + 
  theme(legend.position = "none")
```

Fit EpiLPS using "true" serial interval.

```{r Rt-Pois-EpiLPS}
library(EpiLPS)
#shape <- 3
#scale <- 3
#si1 <- Idist(mean = shape * scale, sd = sqrt(shape * scale^2), dist = #"gamma")$pvec[1:Dmax1]
#si1 <- si1 / sum(si1)
#si2 <- Idist(mean = 2.5^2, sd = sqrt(2.5^3), dist = "gamma")$pvec[1:Dmax2]
#si2 <- si2 / sum(si2)
#si3 <- Idist(mean = 3.5^2, sd = sqrt(3.5^3), dist = "gamma")$pvec[1:Dmax3]
#si3 <- si3 / sum(si3)
#si4 <- si3

mod_epilst_pois1 <- estimR(incidence = incidence1, si = prob_gamma1)
plot(mod_epilst_pois1) + 
  labs(y="EpiLPS Rt with 95% CIs", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_pois2 <- estimR(incidence = incidence2, si = prob_gamma2)
plot(mod_epilst_pois2) + 
  labs(y="EpiLPS Rt with 95% CIs", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_pois3 <- estimR(incidence = incidence3, si = prob_gamma3)
plot(mod_epilst_pois3) + 
  labs(y="EpiLPS Rt with 95% CIs", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_pois4 <- estimR(incidence = incidence4, si = prob_gamma4)
plot(mod_epilst_pois4) + 
  labs(y="EpiLPS Rt with 95% CIs", title="") + 
  theme_bw() + 
  theme(legend.position = "none")
```

## Synthetic data with negative Binomial incidence
```{r NB-samples}
get_nb_incidence <- function(N1, Rt, gamma_pars = c(2.5, 2.5)){
  len <- length(Rt)
  incidence <- numeric(len) # N_1:n
  NB_count <- numeric(len) # y_1:n
  incidence[1] <- N1
  size = 5
  
  NB_count[1] <- rnbinom(1, mu=N1, size=size)
  if(NB_count[1] == 0) NB_count[1] = 1
  for(t in 2:len){
    pi <- discretize_gamma(1:(t-1), gamma_pars[1], gamma_pars[2])
    incidence[t] <- Rt[t] * sum(rev(pi) * NB_count[1:(t-1)])
    NB_count[t] <- rnbinom(1, mu = incidence[t], size=size)
  }
  
  return(NB_count)
}

# case 1
seed <- 629
set.seed(seed)
nb_incidence1 <- get_nb_incidence(N1, Rt1, gamma_pars1)
display_dat(nb_incidence1, Rt1)
check_dat(nb_incidence1, Rt1)

# case 2
set.seed(seed)
nb_incidence2 <- get_nb_incidence(N1, Rt2, gamma_pars2)
display_dat(nb_incidence2, Rt2)
check_dat(nb_incidence2, Rt2)

# case 3
set.seed(seed)
nb_incidence3 <- get_nb_incidence(N1, Rt3, gamma_pars3)
display_dat(nb_incidence3, Rt3)
check_dat(nb_incidence3, Rt3)

# case 4
set.seed(seed)
nb_incidence4 <- get_nb_incidence(N1, Rt4, gamma_pars4)
display_dat(nb_incidence4, Rt4)
check_dat(nb_incidence4, Rt4)
```

Estimate Rt using our Poisson trend filtering using `RtEstim`.

```{r Rt-NB-RtEstim}
cv_mod_nb1 <- cv_estimate_rt(nb_incidence1, korder=0, nfold=3, nsol=50, 
                             dist_gamma = c(3, 3))
rtestim_tuned_mod_nb1 <- cv_mod_nb1$full_fit$Rt[ ,which.min(cv_mod_nb1$cv_scores)]
rt_ci_nb1 <- confband(cv_mod_nb1, "lambda.min") # get 95% confidence band
rt_ci_nb1 %>%
  ggplot(aes(x = 1:length(nb_incidence1), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt") + 
  theme_bw()

mod_nb2 <- estimate_rt(nb_incidence2, korder=3, nsol=50, 
                             dist_gamma = c(2.5, 2.5))
plot(mod_nb2)
cv_mod_nb2 <- cv_estimate_rt(nb_incidence2, korder=3, nfold=3, nsol=50, 
                             dist_gamma = c(2.5, 2.5))
rtestim_tuned_mod_nb2 <- cv_mod_nb2$full_fit$Rt[ ,which.min(cv_mod_nb2$cv_scores)]
rt_ci_nb2 <- confband(cv_mod_nb2, "lambda.min") # get 95% confidence band
rt_ci_nb2 %>%
  ggplot(aes(x = 1:length(nb_incidence2), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt") + 
  theme_bw()

cv_mod_nb3 <- cv_estimate_rt(nb_incidence3, korder=1, nfold=3, nsol=50, 
                             dist_gamma = c(3.5, 3.5))
rtestim_tuned_mod_nb3 <- cv_mod_nb3$full_fit$Rt[ ,which.min(cv_mod_nb3$cv_scores)]
rt_ci_nb3 <- confband(cv_mod_nb3, "lambda.min") # get 95% confidence band
rt_ci_nb3 %>%
  ggplot(aes(x = 1:length(nb_incidence3), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt") + 
  theme_bw()

cv_mod_nb4 <- cv_estimate_rt(nb_incidence4, korder=3, nfold=3, nsol=50, 
                             dist_gamma = c(3.5, 3.5)) # k=2 or 3?
rtestim_tuned_mod_nb4 <- cv_mod_nb4$full_fit$Rt[ ,which.min(cv_mod_nb4$cv_scores)]
rt_ci_nb4 <- confband(cv_mod_nb4, "lambda.min") # get 95% confidence band
rt_ci_nb4 %>%
  ggplot(aes(x = 1:length(nb_incidence4), y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Time", y = "RtEstim Rt") + 
  theme_bw()
```

Estimate Rt using `EpiEstim`.

```{r message=FALSE, Rt-NB-EpiEstim}
# weekly sliding window EpiEstim
mod_epi_nb_week1 <- EpiEstim::estimate_R(incid = nb_incidence1, 
                                         config = config_week1, 
                                         method = method)
plot(mod_epi_nb_week1, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_week2 <- EpiEstim::estimate_R(incid = nb_incidence2, 
                                         config = config_week2, 
                                         method = method)
plot(mod_epi_nb_week2, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_week3 <- EpiEstim::estimate_R(incid = nb_incidence3, 
                                         config = config_week3, 
                                         method = method)
plot(mod_epi_nb_week3, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_week4 <- EpiEstim::estimate_R(incid = nb_incidence4, 
                                         config = config_week4, 
                                         method = method)
plot(mod_epi_nb_week4, "R") + 
  labs(y="EpiEstim Rt with weekly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

# monthly sliding window EpiEstim
mod_epi_nb_month1 <- EpiEstim::estimate_R(incid = nb_incidence1, 
                                         config = config_month1, 
                                         method = method)
plot(mod_epi_nb_month1, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_month2 <- EpiEstim::estimate_R(incid = nb_incidence2, 
                                         config = config_month2, 
                                         method = method)
plot(mod_epi_nb_month2, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_month3 <- EpiEstim::estimate_R(incid = nb_incidence3, 
                                         config = config_month3, 
                                         method = method)
plot(mod_epi_nb_month3, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epi_nb_month4 <- EpiEstim::estimate_R(incid = nb_incidence4, 
                                         config = config_month4, 
                                         method = method)
plot(mod_epi_nb_month4, "R") + 
  labs(y="EpiEstim Rt with monthly sliding window", title="") + 
  theme_bw() + 
  theme(legend.position = "none")
```

Estimate Rt using `EpiLPS`.
```{r Rt-NB-EpiLPS}
mod_epilst_nb1 <- estimR(incidence = nb_incidence1, si = prob_gamma1)
plot(mod_epilst_nb1) + 
  labs(y="EpiLPS Rt", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_nb2 <- estimR(incidence = nb_incidence2, si = prob_gamma2)
plot(mod_epilst_nb2) + 
  labs(y="EpiLPS Rt", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_nb3 <- estimR(incidence = nb_incidence3, si = prob_gamma3)
plot(mod_epilst_nb3) + 
  labs(y="EpiLPS Rt", title="") + 
  theme_bw() + 
  theme(legend.position = "none")

mod_epilst_nb4 <- estimR(incidence = nb_incidence4, si = prob_gamma4)
plot(mod_epilst_nb4) + 
  labs(y="EpiLPS Rt", title="") + 
  theme_bw() + 
  theme(legend.position = "none")
```

## Display and save results for synthetic data

Create a results table to save all Rt and incidence samples, and Rt estimates
using three methods. Save the results table.
```{r Rt-results}
res_dat <- data.table(
  time = 1:len, Rt1 = Rt1, Rt2 = Rt2, Rt3 = Rt3, Rt4 = Rt4,
  Pois_count1 = incidence1, Pois_count2 = incidence2,
  Pois_count3 = incidence3, Pois_count4 = incidence4,
  NB_count1 = nb_incidence1, NB_count2 = nb_incidence2,
  NB_count3 = nb_incidence3, NB_count4 = nb_incidence4,
  Pois_Epi_week1 = c(rep(NA, 7), mod_epiEstim_week1$R$`Mean(R)`),
  Pois_Epi_week2 = c(rep(NA, 7), mod_epiEstim_week2$R$`Mean(R)`),
  Pois_Epi_week3 = c(rep(NA, 7), mod_epiEstim_week3$R$`Mean(R)`),
  Pois_Epi_week4 = c(rep(NA, 7), mod_epiEstim_week4$R$`Mean(R)`),
  Pois_Epi_month1 = c(rep(NA, 30), mod_epiEstim_month1$R$`Mean(R)`),
  Pois_Epi_month2 = c(rep(NA, 30), mod_epiEstim_month2$R$`Mean(R)`),
  Pois_Epi_month3 = c(rep(NA, 30), mod_epiEstim_month3$R$`Mean(R)`),
  Pois_Epi_month4 = c(rep(NA, 30), mod_epiEstim_month4$R$`Mean(R)`),
  Pois_Rtestim1 = rtestim_tuned_mod1,
  Pois_Rtestim2 = rtestim_tuned_mod2,
  Pois_Rtestim3 = rtestim_tuned_mod3,
  Pois_Rtestim4 = rtestim_tuned_mod4,
  Pois_EpiLPS1 = c(rep(NA, 7), mod_epilst_pois1$RLPS$R[-1:-7]),
  Pois_EpiLPS2 = c(rep(NA, 7), mod_epilst_pois2$RLPS$R[-1:-7]),
  Pois_EpiLPS3 = c(rep(NA, 7), mod_epilst_pois3$RLPS$R[-1:-7]),
  Pois_EpiLPS4 = c(rep(NA, 7), mod_epilst_pois4$RLPS$R[-1:-7]),
  
  NB_Epi_week1 = c(rep(NA, 7), mod_epi_nb_week1$R$`Mean(R)`),
  NB_Epi_week2 = c(rep(NA, 7), mod_epi_nb_week2$R$`Mean(R)`),
  NB_Epi_week3 = c(rep(NA, 7), mod_epi_nb_week3$R$`Mean(R)`),
  NB_Epi_week4 = c(rep(NA, 7), mod_epi_nb_week4$R$`Mean(R)`),
  NB_Epi_month1 = c(rep(NA, 30), mod_epi_nb_month1$R$`Mean(R)`),
  NB_Epi_month2 = c(rep(NA, 30), mod_epi_nb_month2$R$`Mean(R)`),
  NB_Epi_month3 = c(rep(NA, 30), mod_epi_nb_month3$R$`Mean(R)`),
  NB_Epi_month4 = c(rep(NA, 30), mod_epi_nb_month4$R$`Mean(R)`),
  NB_Rtestim1 = rtestim_tuned_mod_nb1,
  NB_Rtestim2 = rtestim_tuned_mod_nb2,
  NB_Rtestim3 = rtestim_tuned_mod_nb3,
  NB_Rtestim4 = rtestim_tuned_mod_nb4,
  NB_EpiLPS1 = c(rep(NA, 7), mod_epilst_nb1$RLPS$R[-1:-7]), 
  NB_EpiLPS2 = c(rep(NA, 7), mod_epilst_nb2$RLPS$R[-1:-7]),
  NB_EpiLPS3 = c(rep(NA, 7), mod_epilst_nb3$RLPS$R[-1:-7]),
  NB_EpiLPS4 = c(rep(NA, 7), mod_epilst_nb4$RLPS$R[-1:-7])
)

# save Rt, incidence and estimates
saveRDS(res_dat, here::here("dat/def_examples.RDS"))
```

Display results of Poisson incidence.
```{r Pois-res-display}
cbPalette <- c("#999999", "#E69F00", "#009E73", "#0072B2", "#CC79A7", 
               "#56B4E9", "#D55E00", "#F0E442")
fig1 <- res_dat %>%
  select(time, Rt1, Pois_Epi_week1, Pois_Rtestim1, Pois_EpiLPS1) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, "True Rt" = "Rt1",
                              "RtEstim" = "Pois_Rtestim1", 
                              "EpiEstim" = "Pois_Epi_week1",
                              "EpiLPS" = "Pois_EpiLPS1"
                              )) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt",
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color = "Rt methods", x="", title="Scenario 1: 2-segment constant line") + 
  theme_bw()

fig2 <- res_dat %>%
  select(time, Rt2, Pois_Epi_week2, Pois_Rtestim2, Pois_EpiLPS2) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, "True Rt" = "Rt2",
                              "RtEstim" = "Pois_Rtestim2", 
                              "EpiEstim" = "Pois_Epi_week2",
                              "EpiLPS" = "Pois_EpiLPS2"
                              )) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt",
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 2: 2-segment exponential curve") + 
  theme_bw()

fig3 <- res_dat %>%
  select(time, Rt3, Pois_Epi_week3, Pois_Rtestim3, Pois_EpiLPS3) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, "True Rt" = "Rt3",
                              "RtEstim" = "Pois_Rtestim3", 
                              "EpiEstim" = "Pois_Epi_week3",
                              "EpiLPS" = "Pois_EpiLPS3"
                              )) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt" ,
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  #filter(Rt_type != "EpiLPS") %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 3: 4-segment linear line") + 
  theme_bw()

fig4 <- res_dat %>%
  select(time, Rt4, Pois_Epi_week4, Pois_Rtestim4, Pois_EpiLPS4) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, "True Rt" = "Rt4",
                              "RtEstim" = "Pois_Rtestim4", 
                              "EpiEstim" = "Pois_Epi_week4",
                              "EpiLPS" = "Pois_EpiLPS4"
                              )) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt" ,
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  #filter(Rt_type != "EpiLPS") %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 4: periodic curve") + 
  theme_bw()

library(ggpubr)
fig_Pois_res <- ggpubr::ggarrange(
  fig1, fig2, fig3, fig4, ncol=2, nrow=2, 
  common.legend = TRUE, legend = "bottom",
  font.label = list(size = 14)
  ) 

fig_Pois_res <- ggpubr::annotate_figure(
  fig_Pois_res, 
  left = grid::textGrob("Rt estimates for Poisson incidences", 
                        rot = 90, vjust = 1, gp = grid::gpar(cex = 1)))

ggsave(here::here("fig/Pois-res-plot.png"), fig_Pois_res, width = 8.17, height = 4.71)
```

Display results of negative Binomial incidence.
```{r NB-res-display}
fig21 <- res_dat %>%
  select(time, Rt1, NB_Epi_week1, NB_Rtestim1, NB_EpiLPS1) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, 
                              "EpiEstim" = "NB_Epi_week1",
                              "EpiLPS" = "NB_EpiLPS1",
                              "True Rt" = "Rt1",
                              "RtEstim" = "NB_Rtestim1")) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt",
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 1: 2-segment constant line") + 
  theme_bw()

fig22 <- res_dat %>%
  select(time, Rt2, NB_Epi_week2, NB_Rtestim2, NB_EpiLPS2) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type,
                              "EpiEstim" = "NB_Epi_week2", 
                              "RtEstim" = "NB_Rtestim2", 
                              "EpiLPS" = "NB_EpiLPS2",
                              "True Rt" = "Rt2")) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt" ,
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 2: 2-segment exponential curve") + 
  theme_bw()

fig23 <- res_dat %>%
  select(time, Rt3, NB_Epi_week3, NB_Rtestim3, NB_EpiLPS3) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, 
                              "EpiEstim" = "NB_Epi_week3",
                              "RtEstim" = "NB_Rtestim3", 
                              "EpiLPS" = "NB_EpiLPS3",
                              "True Rt" = "Rt3")) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt" ,
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 3: 4-segment linear line") + 
  theme_bw()

fig24 <- res_dat %>%
  select(time, Rt4, NB_Epi_week4, NB_Rtestim4, NB_EpiLPS4) %>%
  pivot_longer(!time, names_to = "Rt_type", values_to = "Rt_value") %>%
  mutate(Rt_type = fct_recode(Rt_type, 
                              "EpiEstim" = "NB_Epi_week4",
                              "RtEstim" = "NB_Rtestim4", 
                              "EpiLPS" = "NB_EpiLPS4",
                              "True Rt" = "Rt4")) %>%
  mutate(Rt_type = fct_relevel(Rt_type, "True Rt" ,
                              "EpiEstim",
                              "EpiLPS",
                              "RtEstim"
                              )) %>% 
  group_by(Rt_type) %>%
  ggplot(aes(x = time, y = Rt_value, col=Rt_type)) + 
  geom_line(na.rm=TRUE) + 
  scale_colour_manual(values = cbPalette) + 
  labs(y="", color="Rt methods", x="", title="Scenario 4: periodic curve") + 
  theme_bw()

fig_NB_res <- ggarrange(fig21, fig22, fig23, fig24, ncol=2, nrow=2, 
          common.legend = TRUE, legend = "bottom",
          font.label = list(size = 14))
fig_NB_res <- ggpubr::annotate_figure(
  fig_NB_res, 
  left = grid::textGrob("Rt estimates for negative Binomial incidences", 
                        rot = 90, vjust = 1, gp = grid::gpar(cex = 1)))

ggsave(here::here("fig/NB-res-plot.png"), fig_NB_res, width = 8.17, height = 4.71)
```

Save graphical display of estimates across all Rt scenarios for both incidence
distributional assumptions. 
```{r save-res-plot}
fig_res <- ggarrange(fig1, fig21,
          fig2, fig22,
          fig3, fig23,
          fig4, fig24, ncol=2, nrow=4, 
          common.legend = TRUE, legend = "bottom")
ggsave(here::here("fig/res-plot.png"), fig_res, width = 9.77, height = 10.39)
```

Save graphical display of all Rt and incidence samples across both incidence 
distributional assumptions.

```{r save-example-plot}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

save_example <- function(res_dat, palette = cbPalette){
  len <- dim(res_dat)[1]
  fig1 <- res_dat %>%
    ggplot(aes(y = Rt1, x = time)) +
    geom_line() +
    labs(y = "Scenario 1", title = "Rt", x="") + #2-segment constant line
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_bw()
  fig2 <- res_dat %>%
    ggplot(aes(y = Pois_count1, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(title = "Poisson incidence", y=" ", x="") + #(in log scale) 
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_bw()
  fig3 <- res_dat %>%
    ggplot(aes(y = NB_count1, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(title = "Negative Binomial incidence", y=" ", x="") + #(in log scale) 
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_bw()
  
  fig21 <- res_dat %>%
    ggplot(aes(y = Rt2, x = time)) +
    geom_line() +
    labs(y = "Scenario 2", x="") + #2-segment exponential curve
    theme_bw()
  fig22 <- res_dat %>%
    ggplot(aes(y = Pois_count2, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y=" ", x="") +  #(in log scale) 
    theme_bw()
  fig23 <- res_dat %>%
    ggplot(aes(y = NB_count2, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y=" ", x="") +  #(in log scale) 
    theme_bw()
  
  fig31 <- res_dat %>%
    ggplot(aes(y = Rt3, x = time)) +
    geom_line() +
    labs(y = "Scenario 3", x="") + #4-segment linear line
    theme_bw()
  fig32 <- res_dat %>%
    ggplot(aes(y = Pois_count3, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y =" ", x="") +  #(in log scale) 
    theme_bw()
  fig33 <- res_dat %>%
    ggplot(aes(y = NB_count3, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y =" ", x="") +  #(in log scale) 
    theme_bw()

  fig41 <- res_dat %>%
    ggplot(aes(y = Rt4, x = time)) +
    geom_line() +
    labs(y = "Scenario 4") + #periodic curve
    theme_bw()
  fig42 <- res_dat %>%
    ggplot(aes(y = Pois_count4, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y=" ") +  #(in log scale) 
    theme_bw()
  fig43 <- res_dat %>%
    ggplot(aes(y = NB_count4, x = time)) +
    geom_line() +
    #scale_y_log10() + # axis of incidence in log scale
    labs(y=" ") +  #(in log scale) 
    theme_bw()
  
  figfull <- ggpubr::ggarrange(fig1, fig2, fig3, 
            fig21, fig22, fig23,
            fig31, fig32, fig33,
            fig41, fig42, fig43,
            ncol=3, nrow=4, 
          common.legend = TRUE, legend = "bottom")
  ggplot2::ggsave(here::here("fig/plot_samples.png"), figfull, 
                  width = 9.04, height = 10.47)
  print(figfull)
}

head(res_dat)
save_example(res_dat)
```


# Real case 

## Covid-19 Canada

Get latest Covid-19 incidence data in Canada. 

```{r Covid-data}
library(CanCovidData)
dat <- get_british_columbia_case_data() %>%
  rename(HA = `Health Authority`, Date = `Reported Date`,
         Age = `Age group`) %>%
  filter(Date >= ymd("2020-03-01"))

cases <- dat %>%
  count(Date, name = "Cases")

time_diff <- difftime(cases$Date, min(cases$Date), units = "days")
natural_numbers <- as.numeric(time_diff) + 1

covid_cases <- data.table(cases, x = natural_numbers)
covid_dat <- covid_cases %>%
  ggplot(aes(x=Date)) + 
  geom_line(aes(y=Cases), col="black") + 
  labs(x="Date", y='Observed counts') +
  scale_x_date(date_breaks = "6 month") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(hjust = 0.5),
        panel.spacing = unit(0.8, "lines"),
        text = element_text(size = 12))
ggsave(here::here("fig/covid_dat.png"), covid_dat, width = 6.67, height = 4.06)

n <- length(covid_cases$Cases)
```

Estimate Rt using `RtEstim`.

```{r covid-Rt-est}
# rtestim
mod_rt_covid <- estimate_rt(x=cases$Date, covid_cases$Cases, korder = 1, nsol = 50,
                            maxiter=1e7L)
covid_fig1 <- plot(mod_rt_covid) + labs(y="", x="Date")
cv_mod_covid <- cv_estimate_rt(x=cases$Date, covid_cases$Cases, korder = 1, nfold = 3, 
                               maxiter=1e7L, nsol = 50)
mod_rtestim_covid <- cv_mod_covid$full_fit$Rt[ ,which.min(cv_mod_covid$cv_scores)]
rt_ci_covid <- confband(cv_mod_covid, "lambda.min") # get 95% confidence band
covid_fig_cv1 <- rt_ci_covid %>%
  ggplot(aes(x = covid_cases$Date)) +
  geom_line(aes(y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise linear Rt") + 
  theme_bw()

# k=2
mod_rt_covid <- estimate_rt(x=cases$Date, covid_cases$Cases, korder = 2, nsol = 50,
                            maxiter=1e7L)
covid_fig2 <- plot(mod_rt_covid) + labs(y="", x="Date")
cv_mod_covid <- cv_estimate_rt(x=cases$Date, covid_cases$Cases, korder = 2, nfold = 3, 
                               maxiter=1e7L, nsol = 50)
mod_rtestim_covid <- cv_mod_covid$full_fit$Rt[ ,which.min(cv_mod_covid$cv_scores)]
rt_ci_covid <- confband(cv_mod_covid, "lambda.min") # get 95% confidence band
covid_fig_cv2 <- rt_ci_covid %>%
  ggplot(aes(x = covid_cases$Date)) +
  geom_line(aes(y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise quadratic Rt") + 
  theme_bw()

# k=3
mod_rt_covid <- estimate_rt(x=cases$Date, covid_cases$Cases, korder = 3, nsol = 50,
                            maxiter=1e7L)
covid_fig3 <- plot(mod_rt_covid) + labs(y="", x="Date")
cv_mod_covid <- cv_estimate_rt(x=cases$Date, covid_cases$Cases, korder = 3, nfold = 3, 
                               maxiter=1e7L, nsol = 50)
mod_rtestim_covid <- cv_mod_covid$full_fit$Rt[ ,which.min(cv_mod_covid$cv_scores)]
rt_ci_covid <- confband(cv_mod_covid, "lambda.min") # get 95% confidence band
covid_fig_cv3 <- rt_ci_covid %>%
  ggplot(aes(x = covid_cases$Date)) +
  geom_line(aes(y = Rt)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), 
              fill = "gray", alpha = 0.5) +
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise cubic Rt") + 
  theme_bw()

covid_all <- ggarrange(covid_fig_cv1, covid_fig1, covid_fig_cv2, covid_fig2,
          covid_fig_cv3, covid_fig3, ncol=2, nrow=3, 
          common.legend = TRUE, legend = "bottom")

ggsave(here::here("fig/covid_full_res.png"), covid_all, width = 6.91, height = 6.17)
```

## 1918 H1N1 influenza in the USA

Obtain 1918 H1N1 influenza data in the US from `EpiEtim` package.

```{r h1n1-data}
# grab the flu data from EpiEstim package
data("Flu1918")
n <- length(Flu1918$incidence)
flu_dat <- data.frame(Flu_incidence = Flu1918$incidence, Time = 1:n)
flu_fig <- flu_dat %>%
  ggplot(aes(y=Flu_incidence, x=Time)) + 
  geom_line() +
  theme_bw()
ggsave(here::here("fig/flu_dat.png"), flu_fig, width = 5.18, height = 3.34)

```

```{r flu-Rt-est}
mod_rt <- estimate_rt(Flu1918$incidence, korder = 3, nsol = 50)
fig_flu3 <- plot(mod_rt) + labs(y="", x="Date")
cv_mod_flu <- cv_estimate_rt(Flu1918$incidence, korder = 3, 
                             nfold = 3, nsol = 50)
mod_rtestim_flu <- cv_mod_flu$full_fit$Rt[ ,which.min(cv_mod_flu$cv_scores)]

rt_ci_flu <- confband(cv_mod_flu, "lambda.1se") # get 95% confidence band
fig_flu_cv3 <- rt_ci_flu %>%
  ggplot(aes(x = 1:n, y = Rt)) +
  geom_line() +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "gray", alpha = 0.5) +
  
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise cubic Rt") + 
  theme_bw()

# k=2
mod_rt <- estimate_rt(Flu1918$incidence, korder = 2, nsol = 50)
fig_flu2 <- plot(mod_rt) + labs(y="", x="Date")
cv_mod_flu <- cv_estimate_rt(Flu1918$incidence, korder = 2, 
                             nfold = 3, nsol = 50)
mod_rtestim_flu <- cv_mod_flu$full_fit$Rt[ ,which.min(cv_mod_flu$cv_scores)]

rt_ci_flu <- confband(cv_mod_flu, "lambda.1se") # get 95% confidence band
fig_flu_cv2 <- rt_ci_flu %>%
  ggplot(aes(x = 1:n, y = Rt)) +
  geom_line() +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "gray", alpha = 0.5) +
  
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise quadratic Rt") + 
  theme_bw()

mod_rt <- estimate_rt(Flu1918$incidence, korder = 1, nsol = 50)
fig_flu1 <- plot(mod_rt) + labs(y="", x="Date")
cv_mod_flu <- cv_estimate_rt(Flu1918$incidence, korder = 1, 
                             nfold = 3, nsol = 50)
mod_rtestim_flu <- cv_mod_flu$full_fit$Rt[ ,which.min(cv_mod_flu$cv_scores)]

rt_ci_flu <- confband(cv_mod_flu, "lambda.1se") # get 95% confidence band
fig_flu_cv1 <- rt_ci_flu %>%
  ggplot(aes(x = 1:n, y = Rt)) +
  geom_line() +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "gray", alpha = 0.5) +
  
  geom_hline(yintercept = 1, linetype = "dotted") + 
  labs(x="Date", y = "Piecewise linear Rt") + 
  theme_bw()

flu_all <- ggarrange(fig_flu_cv1, fig_flu1, fig_flu_cv2, fig_flu2,
          fig_flu_cv3, fig_flu3, ncol=2, nrow=3, 
          common.legend = TRUE, legend = "bottom")

ggsave(here::here("fig/flu_full_res.png"), flu_all, width = 6.91, height = 6.17)
```
