---
title: "Supplementary details of time-varying reproduction number estimation with trend filtering"
author: "Jiaping Liu, Zhenglun Cai, Paul Gustafson, and Daniel J. McDonald"
date: ""
output: 
  bookdown::pdf_document2: 
    number_sections: true
    fig_caption: true
    df_print: kable
    toc: false
header-includes:
  - \input{header}
  - \input{defs}
  - \def\meas{\texttt{measles}}
  - \def\sars{\texttt{SARS}}
  - \def\flu{\texttt{flu}}
  - \def\meass{\texttt{adj\_measles}}
  - \def\flus{\texttt{adj\_flu}}
  - \def\RtEstim{\texttt{RtEstim}}
  - \def\EpiEstim{\texttt{EpiEstim}}
  - \def\EpiLPS{\texttt{EpiLPS}}
  - \def\EpiFilter{\texttt{EpiFilter}}
  - \def\EpiNow2{\texttt{EpiNow2}}
  - \renewcommand{\thesection}{A.\arabic{section}}
  - \counterwithin{figure}{section}
  - \usepackage{lipsum}
  - \pagenumbering{arabic}
bibliography: ptf.bib  
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "", fig.align="center")
library(rtestim)
library(EpiEstim)
library(EpiLPS)
library(data.table)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(forcats)
library(lubridate)
library(ggpubr)
library(forcats)
library(kableExtra)
source(here::here("src/exp_cluster/generate_data.R"))
source(here::here("src/exp_cluster_ci/design_algos_ci.R"))
source(here::here("src/exp_cluster/epiFilter.R"))
source(here::here("src/exp_cluster/epiSmoother.R"))
```

# Derivation of Kullback Leibler divergence for accuracy comparison

We provide the detailed derivation of the Kullback Leibler (KL) divergence in 
(11) in the manuscript that is used to compare the accuracy of the estimated 
time-varying instantaneous reproduction number with the true ones. Given the 
total infectiousness $\eta$, we compare the distance between the Poisson 
distributions $y\sim Pois(\eta \hat{\calR})$ and $y\sim Pois(\eta \calR)$, 
where $y,\calR\in\bbN_0^n$ are natural numbers including 0, $\eta\in\bbR^n$, 
and $f_0(y;\eta,\calR) = \prod_{t=1}^n \frac{(\eta_t\calR_t)^{y_t} e^{-\eta_t\calR_t}}{y_t!}$,
$f_1(y;\eta,\hat{\calR}) = \prod_{t=1}^n \frac{(\eta_t\hat{\calR}_t)^{y_t} e^{-\eta_t\hat{\calR}_t}}{y_t!}$ are the corresponding density mass functions 
for independent $y_t,t=1,\cdots,n$. 
Then, the KL divergence between them is defined as 
\begin{equation*}
  \begin{split}
    D_{KL}(\calR || \hat{\calR}) &= D_{KL}(f_0(y;\eta,\calR) || f_1(y;\eta,\hat{\calR})) \\
    &= \sum_{y\in\bbN_0^n} f_0(y;\eta,\calR) \log \frac{f_0(y;\eta,\calR)}{f_1(y;\eta,\hat{\calR})} \\
    &= \sum_{y\in\bbN_0^n} \prod_{t=1}^{n} \frac{(\eta_t\calR_t)^{y_t}e^{-\eta_t\calR_t}}{y_t!} 
      \log\prod_{t=1}^{n} \frac{\calR_t^{y_t} e^{-\eta_t\calR_t}}{\hat{\calR}_t^{y_t} e^{-\eta_t\hat{\calR}_t}} \text{ for independent $y_t$, $t=1,...,n$}\\
    &= \sum_{y_n=0}^{\infty}\cdots \sum_{y_1=0}^{\infty} 
      \prod_{t=1}^{n} \frac{(\eta_t\calR_t)^{y_t}e^{-\eta_t\calR_t}}{y_t!} 
      \sum_{t=1}^n \Big(y_t\log \frac{\calR_t}{\hat{\calR}_t} - \eta_t(\calR_t - \hat{\calR}_t)\Big) \\
    &= \sum_{y_n=0}^{\infty}\frac{(\eta_n\calR_n)^{y_n}e^{-\eta_n\calR_n}}{y_n!}\cdots 
      \sum_{y_1=0}^{\infty}\frac{(\eta_1\calR_1)^{y_1}e^{-\eta_1\calR_1}}{y_1!} 
      \Big(y_1\log \frac{\calR_1}{\hat{\calR}_1} + \sum_{t=2}^n y_t\log \frac{\calR_t}{\hat{\calR}_t} - 
      \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t) \Big) \\
    &= \sum_{y_n=0}^{\infty}\frac{(\eta_n\calR_n)^{y_n}e^{-\eta_n\calR_n}}{y_n!}\cdots
      \sum_{y_2=0}^{\infty}\frac{(\eta_2\calR_2)^{y_2}e^{-\eta_2\calR_2}}{y_2!} 
      \Big( \eta_1\calR_1\log\frac{\calR_1}{\hat{\calR}_1} + \sum_{t=2}^n y_t\log 
      \frac{\calR_t}{\hat{\calR}_t} - \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t)\Big) \\
    &= \sum_{y_n=0}^{\infty}\frac{(\eta_n\calR_n)^{y_n}e^{-\eta_n\calR_n}}{y_n!} 
      \Big(\sum_{t=1}^{n-1} \eta_t\calR_t\log \frac{\calR_t}{\hat{\calR}_t} + y_n\log 
      \frac{\calR_n}{\hat{\calR}_n} - \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t) \Big) \\
    &= \sum_{t=1}^n \eta_t\lr{\calR_t \log\frac{\calR_t}{\hat{\calR}_t} + \hat{\calR}_t - \calR_t},
  \end{split}
\end{equation*}
where 
\begin{equation*}
  \begin{split}
    &\sum_{y_1=0}^{\infty}\frac{(\eta_1\calR_1)^{y_1}e^{-\eta_1\calR_1}}{y_1!} 
      \Big(y_1\log \frac{\calR_1}{\hat{\calR}_1} + \sum_{t=2}^n y_t\log \frac{\calR_t}{\hat{\calR}_t} - 
      \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t) \Big) \\
    = & \lr{\sum_{y_1=1}^{\infty}\frac{(\eta_1\calR_1)^{y_1-1}e^{-\eta_1\calR_1}}{(y_1-1)!} 
      \eta_1\calR_1\log \frac{\calR_1}{\hat{\calR}_1} } + 
      \sum_{t=2}^n y_t\log \frac{\calR_t}{\hat{\calR}_t} - 
      \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t), \\
      & \text{ as } \sum_{y_1=0}^{\infty}\frac{(\eta_1\calR_1)^{y_1}e^{-\eta_1\calR_1}}{y_1!} = 1 
      \text{ and }  \frac{(\eta_1\calR_1)^{y_1}e^{-\eta_1\calR_1}}{y_1!} y_1
      \log \frac{\calR_1}{\hat{\calR}_1} = 0 \text{ when } y_1 = 0\\
    = & \eta_1\calR_1\log \frac{\calR_1}{\hat{\calR}_1} + 
      \sum_{t=2}^n y_t\log \frac{\calR_t}{\hat{\calR}_t} - 
      \sum_{t=1}^n \eta_t(\calR_t - \hat{\calR}_t), \text{ as } \sum_{y_1=1}^{\infty}\frac{(\eta_1\calR_1)^{y_1-1}e^{-\eta_1\calR_1}}{(y_1-1)!} = 1.
  \end{split}
\end{equation*}
We use mean KL divergence (denoted, 
$\overline{D_{KL}}(\calR || \hat{\calR}) := D_{KL}(\calR || \hat{\calR}) / n$, 
which is the KL divergence divided by the sequence length) in experiments for 
accuracy comparison.


# Supplmentary details on experimental settings

We compare the accuracy of the estimated instantaneous reproduction numbers using 
the mean Kullback Leibler (KL) divergence with Poisson distributional assumption 
on incidence (we say (mean) KL divergence for short in the following) in (11) across 
our \RtEstim and several alternative 
methods, including \EpiEstim with weekly and monthly sliding windows, \EpiLPS, 
\EpiFilter, \EpiNow2, and \RtEstim with degrees k=0,1,2,3, which yields $9$ 
methods in total. We consider two lengths of epidemics with $n=50$ or $n=300$ 
timepoints respectively. Since \EpiNow2 takes too long to converge (e.g., for 
a long \meas epidemic, it takes almost 2 hours (specifically, 115 minutes 
computed on Cedar cluster provided by Compute Canada)), we only compare it with 
other methods for short \flu epidemics. 

We consider the serial interval (SI) distributions of \meas and \sars to generate 
long synthetic epidemics, and \flu for short epidemics, inspired by @cori2013new
which used SI from real epidemics to illustrate the performance of their method.
The means and standard deviations of SI distributions are estimated by existing 
studies; specifically, $(14.9, 3.9)$ for \meas (@groendyke2011bayesian), 
$(8.4,3.8)$ for \sars (@lipsitch2003transmission), and $(2.6,1.5)$ for 
\flu (@ferguson2005strategies, @boelle2011transmission). 
Incident cases in synthetic \meas epidemics are relatively low (within $1000$ at 
the peak overall), and \sars incident cases are relatively large (between $15000$ 
and $20000$ at the peak overall). We consider a reasonably large 
overdispersion level of negative binomial incidence with size 5. Figure 
\@ref(fig:NB-overdispersion) displays the ratio of standard deviation over 
mean (called, sigma to mean ratio) of incidence across different settings 
using the same set of sample epidemics in Fig 5 and Fig 6, and all figures in 
Section \@ref(sec:samples). Compared to the counterpart of Poisson incidence 
(which decreases quickly to 0 and remains to be under 0.25) per $\calR_t$ 
scenario for each epidemic, the negative binomial incidence appears to have an 
apparently larger sigma to mean ratio (staying at around 0.5 or above), which 
implies a distinguishable overdispersion level. 

```{r NB-overdispersion, echo=FALSE, fig.height=4.4, fig.width=6, fig.cap = "Dispersion level of incidence of sample epidemics."}
dat_example <- readRDS(here::here("dat/data_example.RDS"))
size <- 5
dat_example_NB_SMR <- dat_example %>%
  filter(dist == "Negative Binomial") %>%
  group_by(Rt_case, si_type) %>%
  mutate(smr = sqrt(mean + mean^2 / size) / mean)
dat_example_pois_SMR <- dat_example %>%
  filter(dist == "Poisson") %>%
  group_by(Rt_case, si_type) %>%
  mutate(smr = sqrt(mean) / mean)
dat_example_SMR <- rbind(dat_example_pois_SMR, dat_example_NB_SMR)

# plot the SNR (mean/sd) ratio
fig_smr <- dat_example_SMR %>%
  group_by(dist, Rt_case, si_type) %>%
  ggplot(aes(y = smr, x = time)) + 
  geom_line(aes(color = si_type)) + 
  facet_grid(dist ~ Rt_case) + #, scales = "free"
  scale_color_brewer(palette = "Set1") +
  labs(y = "Sigma to mean ratio", x = "Time") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL))
print(fig_smr)
```

In model fitting, we use both true and misspecified serial interval (SI) 
distributions to test the robustness of our method, compared to other alternatives. 
The misspecification of serial interval distributions are either 
"mild" or ``major", where, in the major misspecification, we use a completely different 
pair of SI parameters, e.g., we use SI of \sars to solve measles epidemics,
and SI of measles to solve short \flu epidemics. While, in the mild SI 
misspecification, we consider slightly adjusted parameters for both \meas and 
\flu epidemics, where the mean is decreased by $2$ for \meas and increased by $2$ 
for \flu and the standard deviation is increased by $10\%$, denoted as \flus and 
\meass respectively. These settings result in 7 pairs of SI distributions 
(for epidemic generating, model fitting), i.e., (\meas, \meas), (\sars, \sars), 
(\meas, \meass), (\meas, \sars) for long epidemics and (\flu, \flu), 
(\flu, \flus), (\flu, \meas) for short epidemics. Figure \@ref(fig:si-dist) 
displays all SI distributions (\meas, \meass, \sars, \flu, and \flus) used in 
the experiments. 

```{r si-dist, echo=FALSE, fig.height=4.4, fig.width=6.2, fig.cap = "Density curves of serial interval distributions used in the experiments."}
dat_si <- data.table(
  x = 0:40,
  measles = dgamma(0:40L, shape = 14.5963, scale = 1.0208),
  SARS = dgamma(0:40L, shape = 4.8864, scale = 1.7190),
  flu = dgamma(0:40L, shape = 3.0044, scale = 0.8654),
  measles_ss = dgamma(0:40L, shape = 9.0420, scale = 1.4267),
  flu_ss = dgamma(0:40L, shape = 7.7723, scale = 0.5918)
)

dat_si %>%
  pivot_longer(!x, values_to = "prob", names_to = "epidemic") %>%
  mutate(epidemic = fct_recode(epidemic,
    "adjusted flu" = "flu_ss",
    "adjusted measles" = "measles_ss"
  )) %>%
  group_by(epidemic) %>%
  ggplot(aes(x = x, y = prob)) + 
  geom_line(aes(col = epidemic)) + 
  geom_ribbon(aes(ymin = 0, ymax = prob, fill = epidemic), alpha = 0.3) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(y = "Probability", x = "Index") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  theme_bw() + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL), fill = "none")
```

Table \@ref(tab:df-exp) summarizes the aforementioned experimental setting for 
accuracy comparison. Poisson and negative binomial (NB) distributions for 
incidence and four $\calR_t$ scenarios are used for all long epidemics. We only
consider one $\calR_t$ scenario (Scenario 3: piecewise linear $\calR_t$) for 
short epidemics. Each experimental setting is replicated for $50$ times, which 
yields $12800$ experiments for long epidemics and $2700$ for short epidemics. 

```{r df-exp, echo=FALSE, message=FALSE, warning=FALSE}
df_exp <- data.table(
  len = c(300, 300, 50),
  si = c("measles", "SARS", "flu"),
  rt = c("1-4", "1-4", "3"),
  inci = rep("Poisson, NB", 3),
  si_pars = c("measles, adj_measles, SARS", "SARS", "flu, adj_flu, measles"),
  method = c(rep("8 methods", 2), "9 methods")
)

kbl(df_exp, booktabs = T, caption = "Summary of experimental settings on accuracy comparison.",
    col.names = c("Length", "SI", "Rt scenario", "Incidence",
                  "SI for modelling", "Method")) %>%
  kable_styling(latex_options = c("repeat_header"))
```

We visualize the selected key results of the accuracy comparison using long synthetic 
epidemics in Section 3.2 in the manuscript. Other main experimental results are displayed in 
Section \@ref(sec:exp-accuracy).

# Supplementary experimental results on accuracy comparison {#sec:exp-accuracy}

## Long epidemics

```{r load-simulation-results, include=FALSE}
cbPalette <- c(
  "#E69F00", "#F0E442", "#009E73", "#FDBE85", "#D55E00", 
  "#CC79A7", "#0072B2", "#b44582", "#56B4E9"
)
Rt_result <- readRDS(here::here("dat/Rt_reduced_results_all.RDS")) 
Rt_result <- Rt_result %>%
  select(method, Rt_case, dist, runtime, Rt_kl, Rt_kl_month, si_type, si_pars, len) %>%
  mutate(method = fct_recode(method,
    "EpiEstim (weekly)" = "EpiEstim(week)",
    "EpiEstim (monthly)" = "EpiEstim(month)",
    "RtEstim (k=0)" = "RtEstim(k=0)",
    "RtEstim (k=1)" = "RtEstim(k=1)",
    "RtEstim (k=2)" = "RtEstim(k=2)",
    "RtEstim (k=3)" = "RtEstim(k=3)"
  )) %>%
  mutate(method = fct_relevel(
    method, 
    "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS",
    "EpiFilter", "EpiNow2",
    "RtEstim (k=0)", "RtEstim (k=1)",
    "RtEstim (k=2)", "RtEstim (k=3)"
  )) %>%
  mutate(dist = fct_recode(dist,
    "Poisson" = "poisson",
    "Negative Binomial" = "NB"
  )) %>%
  mutate(dist = fct_relevel(dist, "Poisson", "Negative Binomial")) %>%
  mutate(Rt_case = as.character(Rt_case)) %>% 
  mutate(Rt_case = fct_recode(Rt_case,
    "Scenario 1" = "1",
    "Scenario 2" = "2",
    "Scenario 3" = "3",
    "Scenario 4" = "4"
  ))
```

We have displayed the accuracy of all methods (where EpiEstim uses weekly 
sliding window) for measles and SARS sample epidemics using KL divergence 
excluding the first weeks since EpiEstim does not provide estimates in the first 
weeks in Fig 3 and Fig 4 in the manuscript, where we exclude the outliers. 
A full visualization including the outliers is in Figure 
\@ref(fig:display-KL-weekly-long). 

Figure \@ref(fig:display-KL-monthly-long) compares \EpiEstim with ***monthly*** 
sliding windows with other methods. We average the KL divergence per coordinate 
excluding the timepoints in the first months for all approaches, since \EpiEstim 
estimates with the monthly sliding windows are not available until the second months. 
The $y$-axis is displayed on a logarithmic scale for a better visualization. 

The relative performance of \EpiEstim with monthly sliding windows, in general, 
is not as good as its weekly sliding window based on the relative positions of 
its boxes and the counterparts of the other methods. It can be explained that 
\EpiEstim with longer sliding windows assume similarity of neighbouring $\calR_t$ 
across longer periods, and thus, is smoother and less accurate compared to the 
one with shorter sliding windows. 

```{r display-KL-weekly-long, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first weeks for measles and SARS epidemics, since EpiEstim with the weekly sliding window does not provide estimates for the first week. Y-axis is on a logarithmic scale."}
Rt_res_long_week <- Rt_result %>% 
  filter(len == 300, !method %in% c("EpiEstim (monthly)", "EpiNow2")) %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method)

fig_kl_week_long <- Rt_res_long_week %>% #
  mutate(mis_si = si_type != si_pars) %>%
  filter(!mis_si) %>%
  group_by(Rt_case, dist, method, si_type) |>
  ggplot(aes(x = method, y = Rt_kl)) + 
  stat_boxplot(aes(col = method)) +
  facet_grid(si_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  scale_colour_manual(values = cbPalette[c(1,3,4,6:9)]) +
  labs(x = "Method", y = "KL divergence (excluding the first weeks)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_week_long)
```

```{r display-KL-monthly-long, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first months for measles and SARS epidemics, since EpiEstim with the monthly sliding window does not provide estimates for the first month. Y-axis is on a logarithmic scale."}
Rt_res_long_month_no_outliers <- Rt_result %>% 
  filter(len == 300, !method %in% c("EpiEstim (weekly)", "EpiNow2")) %>% 
  select(Rt_case, dist, method, Rt_kl_month, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method) %>% 
  mutate(big_out = Rt_kl_month > fivenum(Rt_kl_month, na.rm = TRUE)[4] +
    1.5 * IQR(Rt_kl_month, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out)

Rt_res_long_month <- Rt_result %>% 
  filter(len == 300, !method %in% c("EpiEstim (weekly)", "EpiNow2")) %>% 
  select(Rt_case, dist, method, Rt_kl_month, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method)

fig_kl_month_long <- Rt_res_long_month %>% #
  mutate(mis_si = si_type != si_pars) %>%
  filter(!mis_si) %>%
  group_by(Rt_case, dist, method, si_type) |>
  ggplot(aes(x = method, y = Rt_kl_month)) + 
  stat_boxplot(aes(col = method)) +
  facet_grid(si_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  scale_colour_manual(values = cbPalette[c(2:4,6:9)]) +
  labs(x = "Method", y = "KL divergence (excluding the first months)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_month_long)
```

## Short epidemics

Figures \@ref(fig:display-KL-weekly-flu) and \@ref(fig:display-KL-monthly-flu)
display the KL divergence for short epidemics aggregated over per coordinate 
excluding the first weeks and months respectively to compare EpiEstim with weekly
and monthly sliding windows with other methods including \EpiNow2. 
The difference in accuracy is more obvious given Poisson distributional assumption 
in incidence. To estimate "true" piecewise linear $\calR_t$, piecewise constant and linear 
\RtEstim\ (with $k=0,1$) are most accurate given Poisson incidence, \RtEstim 
($k=2,3$), \EpiLPS\ and \EpiFilter\ are accurate as well with median KL estimates around 1.
Given negative binomial incidence, the advantage of \RtEstim\ is less obvious, 
but \RtEstim\ with all degrees still have the lowest median with a short IQR. 

```{r display-KL-weekly-flu, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first weeks for flu epidemics, since EpiEstim with the weekly sliding window does not provide estimates for the first week. Y-axis is on a logarithmic scale."}
Rt_res_short_week_no_outliers <- Rt_result %>% 
  filter(len == 50, !method %in% c("EpiEstim (monthly)")) %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method) %>% 
  mutate(big_out = Rt_kl > fivenum(Rt_kl, na.rm = TRUE)[4] +
    1.5 * IQR(Rt_kl, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out)
Rt_res_short_week <- Rt_result %>% 
  filter(len == 50, !method %in% c("EpiEstim (monthly)")) %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method)

fig_kl_short_week <- Rt_res_short_week %>% 
  filter(si_type == "flu", si_pars == "flu") |> 
  select(Rt_case, dist, method, Rt_kl) |>
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl)) + # averaged over KL per coordinate
  stat_boxplot(aes(col = method)) +
  facet_grid(~dist, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[-2]) +
  labs(x = "Method", y = "KL divergence (excluding the first weeks)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_short_week)
```

```{r display-KL-monthly-flu, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first months for flu epidemics, since EpiEstim with the monthly sliding window does not provide estimates for the first month. Y-axis is on a logarithmic scale."}
Rt_res_short_month_no_outliers <- Rt_result %>% 
  filter(len == 50, method != "EpiEstim (weekly)") %>% 
  select(Rt_case, dist, method, Rt_kl_month, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method) %>% 
  mutate(big_out = Rt_kl_month > fivenum(Rt_kl_month, na.rm = TRUE)[4] +
    1.5 * IQR(Rt_kl_month, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out)

Rt_res_short_month <- Rt_result %>% 
  filter(len == 50, method != "EpiEstim (weekly)") %>% 
  select(Rt_case, dist, method, Rt_kl_month, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method)

fig_kl_short_month <- Rt_res_short_month %>%#
  filter(si_type == "flu", si_pars == "flu") |> 
  select(Rt_case, dist, method, Rt_kl_month) |>
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl_month)) +
  stat_boxplot(aes(col = method)) +
  facet_grid(~dist, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[-1]) +
  labs(x = "Method", y = "KL divergence (excluding the first months)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_short_month)
```

# Experimental results on accuracy under misspecification of serial interval distributions

## SI misspecification for long epidemics

Figures \@ref(fig:display-KL-weekly-mis-si-measles) and
\@ref(fig:display-KL-monthly-mis-si-measles) display KL divergence (excluding 
the first weeks and the first months respectively) for all 8 methods with "mild" 
misspecification (using shaped and scaled \meas SI parameters) and "major" 
misspecification (using \sars SI parameters) for long \meas epidemics across all 
settings. \RtEstim\ performs robust to misspecification of SI parameters, since 
the median for each problem design is almost always the lowest with the lowest 
IQR. \EpiLPS\ is a strong competitor given negative binomial incidence, since it
assumes incidence to follow negative binomial distributions. \EpiFilter\ is also
quite robust to SI misspecification given Poisson incidence. 

```{r display-KL-weekly-mis-si-measles, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first weeks for measles epidemics with SI misspecification, since EpiEstim with the weekly sliding window does not provide estimates for the first week. Y-axis is on a logarithmic scale."}
Rt_res_long_week_no_outliers <- Rt_result %>% 
  filter(len == 300, method != "EpiEstim (monthly)") %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method) %>% 
  mutate(big_out = Rt_kl > fivenum(Rt_kl, na.rm = TRUE)[4] +
    1.5 * IQR(Rt_kl, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out)

Rt_res_long_week <- Rt_result %>% 
  filter(len == 300, method != "EpiEstim (monthly)") %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method)
Rt_res_long_week_mis_si <- Rt_res_long_week %>%
  mutate(mis_si = si_pars != si_type) %>%
  filter(mis_si) %>%
  select(!mis_si)

fig_kl_full_week_si <- Rt_res_long_week_mis_si %>%
  mutate(mis_type = ifelse(si_pars == "SARS", "Major", "Minor")) %>%
  mutate(mis_type = fct_relevel(mis_type, "Minor", "Major")) %>%
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl)) + 
  stat_boxplot(aes(col = method)) +
  facet_grid(mis_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[c(1,3,4,6:9)]) +
  labs(x = "Method", y = "KL divergence (excluding the first weeks)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_full_week_si)
```

```{r display-KL-monthly-mis-si-measles, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first months for measles epidemics with SI misspecification, since EpiEstim with the monthly sliding window does not provide estimates for the first month. Y-axis is on a logarithmic scale."}
#Rt_res_long_month <- Rt_result %>% 
#  filter(len == 300, method != "EpiEstim (weekly)") %>% 
#  select(Rt_case, dist, method, Rt_kl_month, si_type, si_pars) %>% 
#  group_by(Rt_case, dist, method)
Rt_res_long_month_mis_si <- Rt_res_long_month %>%
  mutate(mis_si = si_pars != si_type) %>%
  filter(mis_si) %>%
  select(!mis_si)

fig_kl_full_month_si <- Rt_res_long_month_mis_si %>%
  mutate(mis_type = ifelse(si_pars == "SARS", "Major", "Minor")) %>%
  mutate(mis_type = fct_relevel(mis_type, "Minor", "Major")) %>%
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl_month)) + 
  stat_boxplot(aes(col = method)) +
  facet_grid(mis_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[c(2:4,6:9)]) +
  labs(x = "Method", y = "KL divergence (excluding the first months)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_full_month_si)
```

## SI misspecification for short epidemics

Figures \@ref(fig:display-KL-weekly-mis-si-flu) and 
\@ref(fig:display-KL-monthly-mis-si-flu) display KL divergence (excluding the 
first weeks and the first months respectively) for all 9 methods with "minor" 
misspecification (using shaped and scaled \flu SI parameters) and "major" 
misspecification (using \meas parameters) for short \flu epidemics across all 
settings. We yield similar conclusion in short epidemics. We also note that 
\EpiNow2 is quite robust to major misspecification in SI parameters, while 
\EpiLPS\ is less satisfactory in major misspecification excluding the first 
weeks in KL computation. It might be due to the large estimates at the beginning
of the epidemics beyond the first weeks, but eliminated within the first months. 

```{r display-KL-weekly-mis-si-flu, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first weeks for flu epidemics with SI misspecification, since EpiEstim with the weekly sliding window does not provide estimates for the first week. Y-axis is on a logarithmic scale."}
Rt_res_short_week_mis_si <- Rt_res_short_week %>%
  mutate(mis_si = si_pars != si_type) %>%
  filter(mis_si) %>%
  select(!mis_si)

fig_kl_short_week_si <- Rt_res_short_week_mis_si %>%
  mutate(mis_type = ifelse(si_pars == "measles", "Major", "Minor")) %>%
  mutate(mis_type = fct_relevel(mis_type, "Minor", "Major")) %>%
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl)) + # averaged over KL per coordinate
  stat_boxplot(aes(col = method)) +
  facet_grid(mis_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[-2]) +
  labs(x = "Method", y = "KL divergence (excluding the first weeks)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_short_week_si)
```

```{r display-KL-monthly-mis-si-flu, echo=FALSE, fig.height=6, fig.width=6.2, fig.cap = "The mean KL divergence excluding the first months for flu epidemics with SI misspecification, since EpiEstim with the monthly sliding window does not provide estimates for the first month. Y-axis is on a logarithmic scale."}
Rt_res_short_month_mis_si <- Rt_res_short_month %>%
  mutate(mis_si = si_pars != si_type) %>%
  filter(mis_si) %>%
  select(!mis_si)

fig_kl_short_month_si <- Rt_res_short_month_mis_si %>%
  mutate(mis_type = ifelse(si_pars == "measles", "Major", "Minor")) %>% 
  mutate(mis_type = fct_relevel(mis_type, "Minor", "Major")) %>%
  group_by(Rt_case, dist, method) |>
  ggplot(aes(x = method, y = Rt_kl_month)) + # averaged over KL per coordinate
  stat_boxplot(aes(col = method)) +
  facet_grid(mis_type * dist ~ Rt_case, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[-1]) +
  labs(x = "Method", y = "KL divergence (excluding the first months)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_short_month_si)
```

# Time comparisons of all methods

Figures \@ref(fig:display-time-long) show the time comparisons across all methods 
for long (\meas and \sars) epidemics. 
\EpiEstim with both sliding windows are very fast and converge in less than 0.1 
seconds. Piecewise constant \RtEstim (with k=0) estimates can be generated 
within 0.1 seconds as well. \EpiLPS is slightly slower, but still very fast and 
within 1 second for all experiments. \EpiFilter\ is in a similar scale of our method
with higher than $0$ degrees. Piecewise linear and cubic \RtEstim (with 
$k=1$ and $k=3$ respectively) are slower, but mostly within 10 seconds.
We also provide an alternative view with the running time of each case in a 
separate panel in Figures \@ref(fig:display-time-meas) and 
\@ref(fig:display-time-sars) for \meas and \sars epidemics respectively. 
We find similar results as in Figure \@ref(fig:display-time-long) in all panels. 

It is remarkable that our \RtEstim computes 50 lambda values with 10-fold CV for 
each experiment, which results in $550$ times of modelling per experiment 
(including modelling for all folds). The running times are no more than 10 seconds 
for most of the experiments, which means the running time for each time of modelling 
is very fast, and on average can be less than 0.02 seconds. The other two methods 
only run once for a fixed set of hyperparameters for each experiment. 

```{r display-time-long, echo = FALSE, fig.height = 5, fig.width = 5, fig.cap = "Running time comparison of all methods for long (measles and SARS) epidemics across all cases. Y-axis is on a logarithmic scale."}
runtime_epi_long <- Rt_result %>% #RtEstim with k=1,2,3 sometimes runs much slower
  filter(len != 50, si_type == "measles") |>
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(x = method, y = runtime / 1e6)) +
  geom_boxplot(aes(col = method)) +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  labs(y = "Running time in seconds", x = "Methods") +
  labs(color = "Methods") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(runtime_epi_long)
```

```{r display-time-meas, echo = FALSE, fig.height = 7.5, fig.width = 6.6, fig.cap = "Running time comparison of all methods for measles epidemics with each pair of SI parameters (measles, adjusted measles, and SARS) for modelling per incidence distribution per Rt scenario (excluding outliers for better illustration). Y-axes are on a logarithmic scale."}
runtime_epi_meas <- Rt_result %>%
  mutate(dist = fct_recode(dist, "NB" = "Negative Binomial")) %>%
  mutate(dist = fct_relevel(dist, "Poisson", "NB")) %>%
  mutate(si_pars = fct_recode(si_pars, "adj_measles" = "measles_ss")) %>%
  filter(len != 50, si_type == "measles") |>
  group_by(dist, Rt_case, method, si_pars) %>%
  mutate(big_out = runtime > fivenum(runtime, na.rm = TRUE)[4] +
    1.5 * IQR(runtime, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out) %>% 
  ggplot(aes(x = method, y = runtime / 1e6)) +
  geom_boxplot(aes(col = method)) +
  facet_grid(si_pars * dist ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  labs(y = "Running time in seconds", x = "Methods") +
  labs(color = "Methods") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(runtime_epi_meas)
```

```{r display-time-sars, echo = FALSE, fig.height = 4.5, fig.width = 6.6, fig.cap = "Running time comparison of all methods for SARS epidemics with each choice of SI parameter for modelling per incidence distribution per Rt scenario. Y-axes are on a logarithmic scale."}
runtime_epi_sars <- Rt_result %>%
  filter(len != 50, si_type == "SARS") |>
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(x = method, y = runtime / 1e6)) +
  geom_boxplot(aes(col = method)) +
  facet_grid(dist ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  labs(y = "Running time in seconds", x = "Methods") +
  labs(color = "Methods") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(runtime_epi_sars)
```

Figure \@ref(fig:display-time-flu-all) displays the running time of all methods 
for short (\flu) epidemics. All methods except \EpiNow2 can converge with in 
around 1 second (within 10 seconds). Figure \@ref(fig:display-time-flu) displays 
the running times for each setting separately, and finds similar results as in 
the overall running time comparison. 

```{r display-time-flu-all, echo = FALSE, fig.height = 4.5, fig.width = 6.6, fig.cap = "Time comparisons of methods for short (flu) epidemics across all cases. Y-axis is on a logarithmic scale."}
runtime_epi_flu_all <- Rt_result %>% 
  filter(len == 50) |>
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(x = method, y = runtime / 1e6)) +
  geom_boxplot(aes(col = method)) +
  facet_wrap(vars(dist)) +
  scale_colour_manual(values = cbPalette) +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  labs(y = "Running time in seconds", x = "Methods") +
  labs(color = "Methods") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(runtime_epi_flu_all)
```

```{r display-time-flu, echo = FALSE, fig.height = 6.5, fig.width = 6.6, fig.cap = "Time comparisons of methods for short (flu) epidemics for pieciewise linear Rt (Scenario 3) for different pairs of SI parameters (flu, adjusted flu, and measles) and incidence distributions in different panels. Y-axes are on a logarithmic scale."}
runtime_epi_flu <- Rt_result %>%
  filter(len == 50) |>
  mutate(dist = fct_recode(dist, "NB" = "Negative Binomial")) %>%
  mutate(dist = fct_relevel(dist, "Poisson", "NB")) %>%
  mutate(si_pars = fct_recode(si_pars, "adjusted flu" = "flu_ss")) %>%
  group_by(si_pars, method) %>%
  ggplot(aes(x = method, y = runtime / 1e6)) +
  geom_boxplot(aes(col = method)) +
  facet_grid(dist ~ si_pars, scales = "free") +
  scale_colour_manual(values = cbPalette) +
  scale_y_log10(breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  labs(y = "Running time in seconds", x = "Methods") +
  labs(color = "Methods") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(runtime_epi_flu)
```

# Confidence interval coverage

## Display estimates and confidence intervals for sample epidemics {#sec:samples}

Fig 5 and Fig 6 in the manuscript provided $\calR_t$ estimates by all methods on 
sample \meas epidemics with Poisson incidence and \sars epidemics with negative 
binomial incidence respectively. Figures \@ref(fig:ci-example-measles-pois) and 
\@ref(fig:ci-example-sars-nb) provide a clearer view of each method with its 
95\% confidence interval in a separate panel. The full display of sample epidemics
for other settings are visualized in Figures \@ref(fig:ci-example-measles-nb) and 
\@ref(fig:ci-example-sars-pois).

All methods (except EpiEstim with the monthly sliding window) fit the epidemics 
with Poisson incidence well with estimate $\hat{\calR}_t$ close to the true 
$\calR_t$ and 95\% CI covering the true value at most timepoints. 
While, given negative binomial incidence, \RtEstim with k=0 misses to recover the
curvature in $\calR_t$ curves, especially in the exponential and periodic scenarios.
\EpiEstim\ with weekly sliding windows and \EpiFilter\ are more wiggly, and 
\EpiLPS\ has wider confidence intervals given negative binomial incidence 
compared to Poisson incidence. For large incidence with negative binomial distribution, 
\EpiFilter\ is extremely wiggly and it is difficult for \RtEstim\ (k=0) to recover 
many changepoints and the curvature especially in exponential and periodic scenarios.
\EpiLPS\ performs well overall, but returns large estimates at the beginning of
the epidemics, which remains after the first week. Overall, our method with different
degrees can recover the changepoints and graphical curvature of $\calR_t$ in all 
scenarios, except in the case of the periodic $\calR_t$ curve with large incidence 
from negative binomial distribution, where \EpiLPS\ has a clear win ignoring the large
estimates at the early stage. 
The performance of sample curves across different settings by different methods 
generally coincides with the findings in the KL divergence estimates. 

```{r ci-example, echo = FALSE}
#res_dat <- readRDS(here::here("dat/data_example_estimates.RDS"))
#
#res_Rt <- res_dat %>%
#  select(incidence:RtEstim3_Rt) %>%
#  pivot_longer(EpiWeek_Rt:RtEstim3_Rt, names_to = "method", values_to = "Rt_value") %>%
#  mutate(method = fct_recode(method,
#    "EpiEstim (weekly)" = "EpiWeek_Rt",
#    "EpiEstim (monthly)" = "EpiMonth_Rt",
#    "EpiLPS" = "EpiLPS_Rt",
#    "EpiFilter" = "EpiFilter_Rt",
#    "RtEstim (k=0)" = "RtEstim0_Rt",
#    "RtEstim (k=1)" = "RtEstim1_Rt",
#    "RtEstim (k=2)" = "RtEstim2_Rt",
#    "RtEstim (k=3)" = "RtEstim3_Rt"
#  ))
#resLo95 <- res_dat %>%
#  select(time, Rt_case, dist, si_type, EpiWeek_Lo95:RtEstim3_Lo95) %>%
#  pivot_longer(EpiWeek_Lo95:RtEstim3_Lo95, names_to = "method", values_to = "Lower_bound")  %>%
#  mutate(method = fct_recode(method,
#    "EpiEstim (weekly)" = "EpiWeek_Lo95",
#    "EpiEstim (monthly)" = "EpiMonth_Lo95",
#    "EpiLPS" = "EpiLPS_Lo95",
#    "EpiFilter" = "EpiFilter_Lo95",
#    "RtEstim (k=0)" = "RtEstim0_Lo95",
#    "RtEstim (k=1)" = "RtEstim1_Lo95",
#    "RtEstim (k=2)" = "RtEstim2_Lo95",
#    "RtEstim (k=3)" = "RtEstim3_Lo95"
#  ))
#resUp95 <- res_dat %>%
#  select(time, Rt_case, dist, si_type, EpiWeek_Up95:RtEstim3_Up95) %>%
#  pivot_longer(EpiWeek_Up95:RtEstim3_Up95, names_to = "method", values_to = "Upper_bound")  %>%
#  mutate(method = fct_recode(method,
#    "EpiEstim (weekly)" = "EpiWeek_Up95",
#    "EpiEstim (monthly)" = "EpiMonth_Up95",
#    "EpiLPS" = "EpiLPS_Up95",
#    "EpiFilter" = "EpiFilter_Up95",
#    "RtEstim (k=0)" = "RtEstim0_Up95",
#    "RtEstim (k=1)" = "RtEstim1_Up95",
#    "RtEstim (k=2)" = "RtEstim2_Up95",
#    "RtEstim (k=3)" = "RtEstim3_Up95"
#  ))
##res_dat_long <- ljoin(res_Rt, resLo95, by = c("time", "Rt_case", "dist", "si_type", "method"))
#res_dat_long <- data.table(res_Rt, Lower_bound = resLo95$Lower_bound, Upper_bound = #resUp95$Upper_bound)
#
#res_dat_long <- res_dat_long %>%
#  mutate(method = fct_relevel(
#    method, 
#    "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter", 
#    "RtEstim (k=0)", "RtEstim (k=1)", "RtEstim (k=2)", "RtEstim (k=3)"
#  )) %>%
#  mutate(dist = fct_relevel(dist, "Poisson", "Negative Binomial")) %>%
#  filter(Rt_value > 0)  
#
#saveRDS(res_dat_long, here::here("dat/data_example_estimates_long.RDS"))
res_dat_long <- readRDS(here::here("dat/data_example_estimates_long.RDS"))
```

```{r ci-example-measles-pois, echo = FALSE, fig.height=7, fig.width=6.5, fig.cap = "Example measles epidemics with Poisson incidence."}
fig_res_full_pois_measles <- res_dat_long %>%
  filter(si_type == "measles", dist == "Poisson") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_ribbon(aes(ymax = Upper_bound, ymin = Lower_bound, fill = method), alpha = 0.5) +
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_grid(method ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for measles epidemics with Poisson incidence") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 5)) + # truncate y due to large EpiLPS at the first few timepoints
  theme_bw() +
  theme(legend.position = "bottom", strip.text.y = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow = TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_full_pois_measles)
```

```{r ci-example-measles-nb, echo = FALSE, fig.height=7, fig.width=6.5, fig.cap = "Example measles epidemics with negative binomial incidence."}
fig_res_full_nb_measles <- res_dat_long %>%
  filter(si_type == "measles", dist == "Negative Binomial") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_ribbon(aes(ymax = Upper_bound, ymin = Lower_bound, fill = method), alpha = 0.5) +
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_grid(method ~ Rt_case, scales = "free_y") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for measles epidemics with negative Binomial incidence") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 5)) + 
  theme_bw() +
  theme(legend.position = "bottom", strip.text.y = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow = TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_full_nb_measles)
```

```{r ci-example-sars-pois, echo = FALSE, fig.height=7, fig.width=6.5, fig.cap = "Example SARS epidemics with Poisson incidence."}
fig_res_full_pois_sars <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Poisson") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_ribbon(aes(ymax = Upper_bound, ymin = Lower_bound, fill = method), alpha = 0.5) +
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_grid(method ~ Rt_case, scales = "free_y") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for SARS epidemics with Poisson incidence") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 5)) + 
  theme_bw() +
  theme(legend.position = "bottom", strip.text.y = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow = TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_full_pois_sars)
```

```{r ci-example-sars-nb, echo = FALSE, fig.height=7, fig.width=6.5, fig.cap = "Example SARS epidemics with negative binomial incidence."}
fig_res_full_nb_sars <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Negative Binomial") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_ribbon(aes(ymax = Upper_bound, ymin = Lower_bound, fill = method), alpha = 0.5) +
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_grid(method ~ Rt_case, scales = "free_y") +
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for SARS epidemics with negative binomial incidence") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 5)) + 
  theme_bw() +
  theme(legend.position = "bottom", strip.text.y = element_blank()) +
  guides(color = guide_legend(title = NULL, byrow = TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_full_nb_sars)
```

## Experimental settings on coverage level comparison of confidence intervals

We focus on a specific $\calR_t$ scenario, the piecewise linear case, and only 
long epidemics to compare the coverage of 95\% confidence intervals across all 8 
methods. We use the "true" serial interval distributions, which are used to generate
the synthetic epidemics, in modelling. 
Table \@ref(tab:df-ci-exp) summarizes the experimental settings.

```{r df-ci-exp, echo=FALSE, message=FALSE, warning=FALSE}
df_ci_exp <- data.table(
  len = rep(300, 2),
  si = c("measles", "SARS"),
  rt = rep(3, 2),
  inci = rep("Poisson, NB", 2),
  si_pars = c("measles", "SARS"),
  method = rep("8 methods", 2)
)

kbl(df_ci_exp, booktabs = T, 
    caption = "Summary of experimental setting on coverage of confidence intervals",
    col.names = c("Length", "SI", "Rt scenario", "Incidence",
                  "SI for modelling", "Method")) %>%
  kable_styling(latex_options = c("repeat_header"))
```


## Experimental results on interval coverage comparison

Figures \@ref(fig:ci-coverage-measles) and \@ref(fig:ci-coverage-sars) displays 
the percentages of coverage of 95% CI per coordinate over 50 random samples for 
\meas and \sars epidemics respectively. 
Low Poisson incidence is the easiest for all methods, with coverage near 100\% at most
timepoints and 0 elsewhere. Large negative Binomial incidence is the "hardest",
while \EpiLPS\ does the best here with avearged coverage at all timepoints close to 1. 
It is consistent to the findings in the accuracy comparison (using KL values) 
and the illustration of sample epidemics in Figures
\@ref(fig:ci-example-measles-pois)--\@ref(fig:ci-example-sars-nb), where 
\EpiLPS\ is the most accurate in this case. 
\RtEstim\ with degrees k=1,2,3 have 100\% coverage at most timepoints except the
changepoints except in the "hardest" case, where larger degrees tend to have higher
percentages of coverge at most timepoints. 
\RtEstim\ with $k=0$ tends to produce overly narrow intervals, leading to lower coverage. 
\EpiEstim\ with weekly sliding windows has a higher chance to fail to cover the true 
$\calR_t$ give negative binomial incidence compared to Poisson and given larger incidence,
its point estimates are quite accurate, but since its 95\% confidence band is too 
narrow and the estimate curves are quite wiggly, it fails to cover the true values. 
\EpiEstim\ with monthly sliding windows has low percentages of interval coverage 
at more timepoints than other methods, especially given negative binomial 
incidence, which is also consistent to the findings in Section \@ref(sec:samples), 
where it misses to recover the $\calR_t$ values at many timepoints and also has 
narrow 95\% confidence intervals. \EpiFilter\ has lower percentages of coverage
given negative binomial incidence than given Poisson incidence, which is consistent
to its performance in accuracy of point estimation.

```{r ci-results, message=FALSE, warning=FALSE, include=FALSE}
CI_result <- readRDS(here::here("dat/rt_cluster_ci_results108.RDS"))
CI_result <- CI_result %>%
  mutate(method = fct_recode(
    method,
    "EpiEstim (weekly)" = "EpiEstim(week)",
    "EpiEstim (monthly)" = "EpiEstim(month)",
    "RtEstim (k=0)" = "RtEstim(k=0)",
    "RtEstim (k=1)" = "RtEstim(k=1)",
    "RtEstim (k=2)" = "RtEstim(k=2)",
    "RtEstim (k=3)" = "RtEstim(k=3)"
  )) %>%
  mutate(method = fct_relevel(
    method, 
    "EpiEstim (weekly)", 
    "EpiEstim (monthly)", 
    "EpiLPS", 
    "EpiFilter",
    "RtEstim (k=0)", 
    "RtEstim (k=1)", 
    "RtEstim (k=2)", 
    "RtEstim (k=3)"
  )) %>%
  mutate(dist = fct_recode(
    dist,
    "Poisson" = "poisson",
    "Negative Binomial" = "NB"
  )) %>%
  mutate(dist = fct_relevel(
    dist, 
    "Poisson", 
    "Negative Binomial"
  )) %>%
  mutate(Rt_case = as.character(Rt_case)) %>% 
  mutate(Rt_case = fct_recode(
    Rt_case,
    "Scenario 1" = "1",
    "Scenario 2" = "2",
    "Scenario 3" = "3",
    "Scenario 4" = "4"
  ))
ci_len <- double(nrow(CI_result))
for(i in 1:nrow(CI_result)) {
  ci_len[i] <- length(CI_result[i,]$ci_coverage[[1]])
}
CI_result <- data.table(CI_result, ci_len)
```

```{r ci-coverage-measles, echo=FALSE, fig.cap="Averaged coverage of CI per coordinate with measles epidemics.", fig.height=5, fig.width=6.5, message=FALSE, warning=FALSE}
Rt_res_ci <- CI_result %>%
  select(si_type, ci_len, dist, Rt_case, method, ci_coverage) %>%
  group_by(si_type, dist, Rt_case, method) %>%
  unnest_wider(ci_coverage, names_sep = "")

# replace NAs with negative values
columns_to_replace <- paste0("ci_coverage", 1:300)
Rt_res_ci_noNAs <- Rt_res_ci %>%
  mutate_at(vars(columns_to_replace), list(~ replace_na(., -1)))
Rt_res_ci_noNAs <- Rt_res_ci_noNAs %>% 
  group_by(si_type, ci_len, dist, Rt_case, method) %>%
  summarize(across(all_of(columns_to_replace), mean)) %>%
  pivot_longer(cols = columns_to_replace, names_to = "index", values_to = "ci_coverage") %>%
  mutate(index = as.numeric(gsub("\\D", "", index))) %>%
  mutate(time_index = 300 - ci_len + index) %>%
  group_by(dist, method, Rt_case) %>%
  filter(ci_coverage >= 0) 

ci_cover_measles <- Rt_res_ci_noNAs %>% # exclude NAs for certain methods
  filter(si_type == "measles") %>% # SARS
  ggplot(aes(y = ci_coverage, x = time_index)) +
  geom_line(aes(col = method)) + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) + 
  labs(y = "CI coverage for measles epidemics", x = "Methods") + 
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_cover_measles)
```

```{r ci-coverage-sars, echo=FALSE, fig.cap="Averaged coverage of CI per coordinate with SARS epidemics.", fig.height=5, fig.width=6.5, message=FALSE, warning=FALSE}
ci_cover_sars <- Rt_res_ci_noNAs %>% # exclude NAs for certain methods
  filter(si_type == "SARS") %>% # SARS
  ggplot(aes(y = ci_coverage, x = time_index)) +
  geom_line(aes(col = method)) + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) + 
  labs(y = "CI coverage for SARS epidemics", x = "Methods") + 
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_cover_sars)
```

Figures \@ref(fig:ci-percentage-measles) and \@ref(fig:ci-percentage-sars) 
displays the percentages of coverage of 95% CI across all timepoints averaged 
over 50 random measles and SARS epidemics respectively. 
CIs of \RtEstim\ with $k=1,2,3$ have nearly 100\% coverage across all timepoints 
for all random samples except in the "hardest" problem, where the incidence is 
large and overdispersed. The coverage of \RtEstim\ $k=0$ is lower than for other 
degrees, similar to the above. 
\EpiFilter\ has larger averaged percentages across all timepoints given Poisson 
incidence compared to negative binomial incidence. \EpiEstim with weekly sliding
windows has higher coverage compared to monthly windows, while their percentages 
of coverage across all timepoints are less than 95\% in most cases. 
The percentage of coverage of \EpiLPS\ is close to 95\% in most cases, and even
in the "hardest" problem, its coverage is roughly higher than 70\%. 


```{r ci-percentage-measles, echo=FALSE, fig.cap="Averaged percentages of CI coverage with measles epidemics. The orange dashed line represents 95\\% percentage of coverage across all timepoints.", fig.height=4.5, fig.width=6.2, message=FALSE, warning=FALSE}
ci_percent_measles <- CI_result %>% 
  filter(si_type == "measles", !is.na(ci_percentage)) %>%
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(y = ci_percentage, x = method)) + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  geom_boxplot(aes(col = method)) +
  scale_colour_manual(values = cbPalette[-5]) +
  geom_hline(yintercept = 0.95, color = "orange", linetype = "dashed") + 
  labs(x = "Methods", y = "CI percentages of coverage for measles epidemics") + 
  theme_bw() + 
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_percent_measles)
```


```{r ci-percentage-sars, echo=FALSE, fig.cap="Averaged percentages of CI coverage with SARS epidemics. The orange dashed line represents 95\\% percentage of coverage across all timepoints.", fig.height=4.5, fig.width=6.2, message=FALSE, warning=FALSE}
ci_percent_sars <- CI_result %>% 
  filter(si_type == "SARS", !is.na(ci_percentage)) %>%
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(y = ci_percentage, x = method)) + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  geom_boxplot(aes(col = method)) +
  scale_colour_manual(values = cbPalette[-5]) +
  geom_hline(yintercept = 0.95, color = "orange", linetype = "dashed") + 
  labs(x = "Methods", y = "CI percentages of coverage for SARS epidemics") + 
  theme_bw() + 
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_percent_sars)
```

We also output the interval score [@Bracher_2021] 
$$score_{\alpha}(\calR,u,l) = \frac{1}{n} \sum_{t=1}^n (u_t-l_t) + \frac{2}{\alpha}(l_t-\calR_t)\boldsymbol{1}_{(\calR_t<l_t)} + \frac{2}{\alpha}(\calR_t-u_t)\boldsymbol{1}_{(\calR_t>u_t)},$$ 
where $\alpha=0.05$ is the significance level, $l,u$ are the lower and upper 
bounds, $\calR_t$ is the true instantaneous reproduction number, and 
$\mathbf{1}_{X}$ is the indicator function of the condition $X$.
A confidence band that covers the true value more frequently with a	shorter 
interval width will have a lower interval score. 
Figures \@ref(fig:ci-score-measles) and \@ref(fig:ci-score-sars) displays the 
interval scores of 95% CI averaged over 50 random measles and SARS epidemics 
respectively. \RtEstim\ always has the lowest or close to the lowest interval
scores. In Poisson cases, \EpiFilter\ has the lowest interval scores (less than
1), and the scores of \RtEstim\ are always around 1.

```{r ci-score-measles, echo=FALSE, fig.cap="Averaged interval scores with measles epidemics. EpiLPS is excluded, since it's scores are always larger than 100.", fig.height=4.5, fig.width=6.2, message=FALSE, warning=FALSE}
ci_score_measles <- CI_result %>% 
  filter(method != "EpiLPS") %>% # too large
  filter(si_type == "measles") %>%
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(y = ci_score, x = method)) +
  facet_grid(dist ~ Rt_case, scales = "free") +
  geom_boxplot(aes(col = method)) +
  scale_colour_manual(values = cbPalette[-c(5)]) +
  scale_y_log10() +
  theme_bw() + 
  labs(y = "Interval scores for measles epidemics", x = "Methods") + 
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_score_measles)
```

```{r ci-score-sars, echo=FALSE, fig.cap="Averaged interval scores with SARS epidemics. EpiLPS is excluded, since it's scores are always larger than 100.", fig.height=4.5, fig.width=6.2, message=FALSE, warning=FALSE}
ci_score_sars <- CI_result %>% 
  filter(method != "EpiLPS") %>% # too large
  filter(si_type == "SARS") %>%
  group_by(dist, Rt_case, method) %>%
  ggplot(aes(y = ci_score, x = method)) +
  facet_grid(dist ~ Rt_case, scales = "free") +
  geom_boxplot(aes(col = method)) +
  scale_colour_manual(values = cbPalette[-c(3,5)]) +
  scale_y_log10() +
  theme_bw() + 
  labs(y = "Interval scores for SARS epidemics", x = "Methods") + 
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(ci_score_sars)
```


# Data examples and alternative visualizations of Figs 5 and 6

## More visualization of example epidemics

We generate \meas and \sars epidemics using Poisson and negative binomial
incidence distributions for each experimental settings. The condensed display of estimates 
for \meas with Poisson incidence and \sars with negative binomial incidence are provided in 
Fig 5 and Fig 6 in the manuscript. A full visualization of each case is provided 
in Section \@ref(sec:samples). 
Here, we provide the condensed visualization of other cases in Figures 
\@ref(fig:display-example-pois-sars) and \@ref(fig:display-example-nb-meas). 
All methods provide accurate point estimates given large incidence from Poisson 
distribution, while \EpiEstim\ (with weekly sliding window) and \EpiFilter\ are 
more wiggly given negative binomial incidence.

```{r display-example-pois-sars, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Example of instantaneous reproduction number estimation for SARS epidemics with Poisson observations."}
fig_res_pois_sars <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Poisson") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 8)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_pois_sars)
```

```{r  display-example-nb-meas, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Example of instantaneous reproduction number estimation for measles epidemics with negative binomial observations."}
fig_res_NB_measles <- res_dat_long %>%
  filter(si_type == "measles", dist == "Negative Binomial") %>% 
  #filter(method != "EpiLPS") %>% # EpiLPS is large for the first 14 days
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 8)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_NB_measles)
```

## Alternative view of difference between fitted and true Rt estimates

We also provide an alternative view of Fig 5 \& Fig 6 in the manuscript by 
plotting $\calR_t - \hat{\calR}_t$ per coordinate $t$ in Figures 
\@ref(fig:display-example-pois-meas-alter) and 
\@ref(fig:display-example-nb-sars-alter) respectively. 
Figures \@ref(fig:display-example-pois-sars-alter) and 
\@ref(fig:display-example-nb-meas-alter) provide the alternative view of 
\@ref(fig:display-example-pois-sars) and \@ref(fig:display-example-nb-meas) respectively.
We notice the different is larger at the changepoints for most methods. 
In the sinusoidal periodic scenario, the different of many method shows a 
periodic pattern, which implies that there is a periodic pattern that fails to 
be filtered by many methods. It makes sense since the "true" $calR_t$ curve is
sinusoidal, while each method only filters the curve to a fixed polynomial degree. 

```{r display-example-pois-meas-alter, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Difference between of the true instantaneous reproduction number and its estimation for measles epidemics with Poisson observations."}
fig_res_pois_meas_alter <- res_dat_long %>%
  filter(si_type == "measles", dist == "Poisson") %>% 
  mutate(dRt = Rt_value - trueRt) %>%
  group_by(Rt_case, method) %>%
  ggplot(aes(y = dRt, x = time)) + 
  geom_line(aes(col = method), lwd=0.5) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Different between true Rt and Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(-1.5, 1.5)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_pois_meas_alter)
```

```{r display-example-nb-sars-alter, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Difference between of the true instantaneous reproduction number and its estimation for SARS epidemics with negative binomial observations."}
fig_res_nb_sars_alter <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Negative Binomial") %>% 
  mutate(dRt = Rt_value - trueRt) %>%
  group_by(Rt_case, method) %>%
  ggplot(aes(y = dRt, x = time)) + 
  geom_line(aes(col = method), lwd=0.5) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Different between true Rt and Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(-1.5, 1.5)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_nb_sars_alter)
```

```{r display-example-pois-sars-alter, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Difference between of the true instantaneous reproduction number and its estimation for SARS epidemics with Poisson observations."}
fig_res_pois_meas_alter <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Poisson") %>% 
  mutate(dRt = Rt_value - trueRt) %>%
  group_by(Rt_case, method) %>%
  ggplot(aes(y = dRt, x = time)) + 
  geom_line(aes(col = method), lwd=0.5) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Different between true Rt and Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(-1.5, 1.5)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_pois_meas_alter)
```

```{r display-example-nb-meas-alter, echo=FALSE, fig.height=4, fig.width=6.2, fig.cap = "Difference between of the true instantaneous reproduction number and its estimation for measles epidemics with negative binomial observations."}
fig_res_nb_sars_alter <- res_dat_long %>%
  filter(si_type == "measles", dist == "Negative Binomial") %>% 
  mutate(dRt = Rt_value - trueRt) %>%
  group_by(Rt_case, method) %>%
  ggplot(aes(y = dRt, x = time)) + 
  geom_line(aes(col = method), lwd=0.5) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Different between true Rt and Rt estimates") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(-1.5, 1.5)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_nb_sars_alter)
```

# Application of RtEstim and all competitors on real epidemics 

We apply all methods on Covid19 incidence in Canada, and the estimated are displayed 
in \@ref(fig:plot-cancovid-ci). An alternative display which plots all estimated 
curves in one panel for an easier comparison is provided in 
\@ref(fig:plot-cancovid-one). All methods provide similar $\hat{\calR}_t$ curves
beyond the early stage. Many methods, including \RtEstim\ (k=1,2), \EpiLPS, and 
\EpiEstim\ (weekly sliding window), all have large estimates (larger than 3) at 
the early stage of the epidemic. \EpiFilter\ is much more wiggly than other estimated
$\hat{\calR}_t$ curves. All methods agree that the instantaneous reproduction number
of Covid19 in Canada decreases to below 1 near June 2021 and reaches a small peak 
afterwards, and then decreases slowly until an outbreak at the end of 2021. 
The instantaneous reproduction number decreases slowly, and remains close to but below 1.

```{r get-cancovid-si, eval=FALSE, include=FALSE}
# get serial interval distributions of different variants
can_pred_class <- readRDS(here::here("dat/can_pred_class.RDS"))
colnames(can_pred_class)[1] <- "date"
delay_dstns_can <- readRDS(here::here("dat/delay-distns-byvar.rds")) |>
  filter(para == "SI", type %in% unique(can_pred_class$var)) 

dates_index <- as.numeric(rtestim::cancovid$date - rtestim::cancovid$date[1] + 1)

# take weighted average of four variants based on the percentages in the time interval
delay_distn1 <- discretize_gamma(dates_index, delay_dstns_can[1,]$shape, delay_dstns_can[1,]$scale)
delay_distn2 <- discretize_gamma(dates_index, delay_dstns_can[2,]$shape, delay_dstns_can[2,]$scale)
delay_distn3 <- discretize_gamma(dates_index, delay_dstns_can[3,]$shape, delay_dstns_can[3,]$scale)
delay_distn4 <- discretize_gamma(dates_index, delay_dstns_can[4,]$shape, delay_dstns_can[4,]$scale)

len <- nrow(rtestim::cancovid)
prob_vec <- left_join(rtestim::cancovid, can_pred_class) %>%
  count(var) %>%
  mutate(percent = n / len)
prob_vec

covid_delay_distn <- prob_vec$percent[1]*delay_distn1 + 
                     prob_vec$percent[2]*delay_distn2 + 
                     prob_vec$percent[3]*delay_distn3 + 
                     prob_vec$percent[4]*delay_distn4
covid_delay_distn <- covid_delay_distn / sum(covid_delay_distn)
```

```{r fit-all-cancovid, eval=FALSE, include=FALSE}
alpha <- 0.05
# EpiNow2: cannot work for long epidemics

# RtEstim k = 0 ----
fit_cancovid_RtEstim0 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 0, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn)
ci_cancovid_RtEstim0 <- confband(fit_cancovid_RtEstim0, "lambda.min", 1-alpha)

RtEstim0_Rt_cancovid <- ci_cancovid_RtEstim0[,1][[1]]
RtEstim0_95Lo_cancovid <- ci_cancovid_RtEstim0[,2][[1]]
RtEstim0_95Up_cancovid <- ci_cancovid_RtEstim0[,3][[1]]

# RtEstim k=1,2,3 ----
fit_cancovid_RtEstim1 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 1, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn)
ci_cancovid_RtEstim1 <- confband(fit_cancovid_RtEstim1, "lambda.min", 1-alpha)

fit_cancovid_RtEstim2 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 2, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn)
ci_cancovid_RtEstim2 <- confband(fit_cancovid_RtEstim2, "lambda.min", 1-alpha)

fit_cancovid_RtEstim3 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 3, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn)
ci_cancovid_RtEstim3 <- confband(fit_cancovid_RtEstim3, "lambda.min", 1-alpha)

RtEstim1_Rt_cancovid <- ci_cancovid_RtEstim1[,1][[1]]
RtEstim1_95Lo_cancovid <- ci_cancovid_RtEstim1[,2][[1]]
RtEstim1_95Up_cancovid <- ci_cancovid_RtEstim1[,3][[1]]
RtEstim2_Rt_cancovid <- ci_cancovid_RtEstim2[,1][[1]]
RtEstim2_95Lo_cancovid <- ci_cancovid_RtEstim2[,2][[1]]
RtEstim2_95Up_cancovid <- ci_cancovid_RtEstim2[,3][[1]]
RtEstim3_Rt_cancovid <- ci_cancovid_RtEstim3[,1][[1]]
RtEstim3_95Lo_cancovid <- ci_cancovid_RtEstim3[,2][[1]]
RtEstim3_95Up_cancovid <- ci_cancovid_RtEstim3[,3][[1]]


# EpiEstim ----
incid <- data.frame(I = rtestim::cancovid$incident_cases, dates = rtestim::cancovid$date)
## weekly
t_start_week <- seq(2, len - 6)
t_end_week <- t_start_week + 6
config_week <- make_config(list(si_distr = c(0, covid_delay_distn), t_start = t_start_week, t_end = t_end_week))
# fit model
fit_week <- EpiEstim::estimate_R(incid = incid, config = config_week, 
                                 method = "non_parametric_si")
l <- which(colnames(fit_week$R) == paste0("Quantile.",alpha/2,"(R)"))
u <- which(colnames(fit_week$R) == paste0("Quantile.",1 - alpha/2,"(R)"))
fill_NAs <- function(Rt_fitted, n) {
    non_nas <- which(!is.na(Rt_fitted))
    m <- length(non_nas)
    return(c(rep(-1, n-m), Rt_fitted[non_nas]))
  }
EpiWeek_95Lo_cancovid <- fill_NAs(fit_week$R[,l], len) 
EpiWeek_95Up_cancovid <- fill_NAs(fit_week$R[,u], len)
EpiWeek_Rt_cancovid <- fill_NAs(fit_week$R$`Mean(R)`, len)

## monthly
t_start_month <- seq(2, len - 29)
t_end_month <- t_start_month + 29
config_month <- make_config(list(si_distr = c(0, covid_delay_distn), t_start = t_start_month, t_end = t_end_month))
# fit model
fit_month <- EpiEstim::estimate_R(incid = incid, config = config_month, 
                                  method = "non_parametric_si")
l <- which(colnames(fit_month$R) == paste0("Quantile.",alpha/2,"(R)"))
u <- which(colnames(fit_month$R) == paste0("Quantile.",1 - alpha/2,"(R)"))
fill_NAs <- function(Rt_fitted, n) {
    non_nas <- which(!is.na(Rt_fitted))
    m <- length(non_nas)
    return(c(rep(-1, n-m), Rt_fitted[non_nas]))
  }
EpiMonth_95Lo_cancovid <- fill_NAs(fit_month$R[,l], len) 
EpiMonth_95Up_cancovid <- fill_NAs(fit_month$R[,u], len)
EpiMonth_Rt_cancovid <- fill_NAs(fit_month$R$`Mean(R)`, len)


# EpiLPS ----
fitLPS <- EpiLPS::estimR(incidence = rtestim::cancovid$incident_cases, si = covid_delay_distn, K = 40, 
                         dates = rtestim::cancovid$date, CoriR = FALSE)
l <- which(colnames(fitLPS$RLPS) == paste0("Rq", alpha/2)) # alpha can only by 0.05, 0.1, 0.5
u <- which(colnames(fitLPS$RLPS) == paste0("Rq", 1 - alpha/2))
EpiLPS_Rt_cancovid <- fill_NAs(fitLPS$RLPS$R, len)
EpiLPS_95Lo_cancovid <- fill_NAs(fitLPS$RLPS[,l], len)
EpiLPS_95Up_cancovid <- fill_NAs(fitLPS$RLPS[,u], len)

# EpiFilter ----
# cannot use dates as input, but the delays are incorporated by the delay distribution
m <- 2000
pR0 <- (1 / m) * rep(1, m)
Rgrid <- seq(0.01, 10, length.out = m)
w <- delay_calculator(rtestim::cancovid$incident_cases, x = rtestim::cancovid$date, delay_distn = covid_delay_distn)
R_filter <- epiFilter(Rgrid, m, eta=0.1, pR0, len, w, rtestim::cancovid$incident_cases, a = alpha)
R_smoother <- epiSmoother(Rgrid, m, R_filter[[4]], R_filter[[5]], len, R_filter[[6]], a = alpha)
EpiFilter_Rt_cancovid <- R_smoother[[3]]
EpiFilter_95Lo_cancovid <- R_smoother[[2]][1,]
EpiFilter_95Up_cancovid <- R_smoother[[2]][2,]

# results ----
dat_cancovid_fitted_long <- data.table(
  Date = rep(rtestim::cancovid$date, 8),
  Method = rep(c("RtEstim (k=0)", "RtEstim (k=1)", "RtEstim (k=2)", "RtEstim (k=3)", 
                 "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter"), each = len),
  fitted_Rt = c(RtEstim0_Rt_cancovid, RtEstim1_Rt_cancovid, RtEstim2_Rt_cancovid, RtEstim3_Rt_cancovid, 
                EpiWeek_Rt_cancovid, EpiMonth_Rt_cancovid, EpiLPS_Rt_cancovid, EpiFilter_Rt_cancovid),
  ci_lo95 = c(RtEstim0_95Lo_cancovid, RtEstim1_95Lo_cancovid, RtEstim2_95Lo_cancovid, RtEstim3_95Lo_cancovid, 
              EpiWeek_95Lo_cancovid, EpiMonth_95Lo_cancovid, EpiLPS_95Lo_cancovid, EpiFilter_95Lo_cancovid),
  ci_hi95 = c(RtEstim0_95Up_cancovid, RtEstim1_95Up_cancovid, RtEstim2_95Up_cancovid, RtEstim3_95Up_cancovid,
              EpiWeek_95Up_cancovid, EpiMonth_95Up_cancovid, EpiLPS_95Up_cancovid, EpiFilter_95Up_cancovid)
)
dat_cancovid_fitted_long <- dat_cancovid_fitted_long %>%
  mutate(Method = fct_relevel(
    Method, 
    "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter",
    "RtEstim (k=0)", "RtEstim (k=1)",
    "RtEstim (k=2)", "RtEstim (k=3)"
  ))
saveRDS(dat_cancovid_fitted_long, here::here("dat/dat_cancovid_fitted.RDS"))
```

```{r plot-cancovid-ci, echo=FALSE, fig.height=8, fig.width=6.2, fig.cap = "Rt estimates with CIs for Covid19. Y-axes are truncated beyond 3 for a better display of the fluctuation in small values."}
dat_cancovid_fitted_long <- readRDS(here::here("dat/dat_cancovid_fitted.RDS"))
dat_cancovid_fitted_long %>%
  filter(fitted_Rt > 0) %>% # remove NAs marked by -1
  group_by(Method) %>%
  ggplot(aes(x = Date)) + 
  geom_line(aes(y = fitted_Rt, col = Method)) + 
  geom_ribbon(aes(ymin = ci_lo95, ymax = ci_hi95, fill = Method), alpha = 0.5) +
  facet_wrap(~ Method, scales = "free_y", ncol = 1) + 
  coord_cartesian(ylim=c(0, 3)) +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_x_date(date_breaks = "12 months", date_labels = "%Y", expand = expansion(c(0, 0.05))) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(x = "Year", y = "") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
```

```{r plot-cancovid-one, echo=FALSE, fig.height=3.5, fig.width=6.2, fig.cap = "Rt estimates for Covid19. Y-axis is truncated beyond 3 for a better display of the fluctuation in small values. EpiFilter is excluded here, because its estimates are too wiggly and make the plot less readable."}
dat_cancovid_fitted_long %>%
  filter(fitted_Rt > 0, Method != "EpiFilter") %>% # remove NAs marked by -1
  group_by(Method) %>%
  ggplot(aes(x = Date)) + 
  geom_line(aes(y = fitted_Rt, col = Method)) + 
  #facet_wrap(~ Method, scales = "free_y", ncol = 1) + 
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_x_date(date_breaks = "12 months", date_labels = "%Y", expand = expansion(c(0, 0.05))) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim=c(0, 3)) + 
  scale_colour_manual(values = cbPalette[-c(4,5)]) +
  labs(x = "Year", y = "") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
```

We also apply all methods on Flu in 1918 as well. The results are visualized in Figures 
\@ref(fig:plot-flu-ci) and \@ref(fig:plot-flu-one). 
\EpiEstim\ with weekly sliding windows, \EpiFilter\ and \RtEstim\ (k=0) captures 
the peak of $\calR_t$ (close to 3)
at around day 30 since the start of the epidemic. While \EpiEstim\ with monthly 
sliding windows, \EpiLPS, \RtEstim\ (k=2,3) captures the increase around day 30, 
but have smaller estimates. Most methods agree that after around day 50, the 
instantaneous reproduction number decreases to and remains below 1.

```{r fit-all-flu, eval=FALSE, include=FALSE}
data("Flu1918")

len <- length(Flu1918$incidence)
alpha <- 0.05
flu_delay_distn <- Flu1918$si_distr
time <- 1:len

# RtEstim k = 0
fit_flu_RtEstim0 <- cv_estimate_rt(
  observed_counts = Flu1918$incidence, korder = 0, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = flu_delay_distn)
ci_flu_RtEstim0 <- confband(fit_flu_RtEstim0, "lambda.min", 1-alpha)

RtEstim0_Rt_flu <- ci_flu_RtEstim0[,1][[1]]
RtEstim0_95Lo_flu <- ci_flu_RtEstim0[,2][[1]]
RtEstim0_95Up_flu <- ci_flu_RtEstim0[,3][[1]]

# RtEstim k=1,2,3
fit_flu_RtEstim1 <- cv_estimate_rt(
  observed_counts = Flu1918$incidence, korder = 1, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = flu_delay_distn)
ci_flu_RtEstim1 <- confband(fit_flu_RtEstim1, "lambda.min", 1-alpha)

fit_flu_RtEstim2 <- cv_estimate_rt(
  observed_counts = Flu1918$incidence, korder = 2, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = flu_delay_distn)
ci_flu_RtEstim2 <- confband(fit_flu_RtEstim2, "lambda.min", 1-alpha)

fit_flu_RtEstim3 <- cv_estimate_rt(
  observed_counts = Flu1918$incidence, korder = 3, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, delay_distn = flu_delay_distn)
ci_flu_RtEstim3 <- confband(fit_flu_RtEstim3, "lambda.min", 1-alpha)

RtEstim1_Rt_flu <- ci_flu_RtEstim1[,1][[1]]
RtEstim1_95Lo_flu <- ci_flu_RtEstim1[,2][[1]]
RtEstim1_95Up_flu <- ci_flu_RtEstim1[,3][[1]]
RtEstim2_Rt_flu <- ci_flu_RtEstim2[,1][[1]]
RtEstim2_95Lo_flu <- ci_flu_RtEstim2[,2][[1]]
RtEstim2_95Up_flu <- ci_flu_RtEstim2[,3][[1]]
RtEstim3_Rt_flu <- ci_flu_RtEstim3[,1][[1]]
RtEstim3_95Lo_flu <- ci_flu_RtEstim3[,2][[1]]
RtEstim3_95Up_flu <- ci_flu_RtEstim3[,3][[1]]

# EpiEstim (weekly)
t_start_week <- seq(2, len - 6)
t_end_week <- t_start_week + 6
config_week <- make_config(list(si_distr = c(0, flu_delay_distn), t_start = t_start_week, t_end = t_end_week))
# fit model
fit_week <- EpiEstim::estimate_R(incid = Flu1918$incidence, config = config_week, method = "non_parametric_si")
l <- which(colnames(fit_week$R) == paste0("Quantile.",alpha/2,"(R)"))
u <- which(colnames(fit_week$R) == paste0("Quantile.",1 - alpha/2,"(R)"))
fill_NAs <- function(Rt_fitted, n) {
    non_nas <- which(!is.na(Rt_fitted))
    m <- length(non_nas)
    return(c(rep(-1, n-m), Rt_fitted[non_nas]))
  }
EpiWeek_95Lo_flu <- fill_NAs(fit_week$R[,l], len) 
EpiWeek_95Up_flu <- fill_NAs(fit_week$R[,u], len)
EpiWeek_Rt_flu <- fill_NAs(fit_week$R$`Mean(R)`, len)

## monthly
t_start_month <- seq(2, len - 29)
t_end_month <- t_start_month + 29
config_month <- make_config(list(si_distr = c(0, flu_delay_distn), t_start = t_start_month, t_end = t_end_month))
# fit model
fit_month <- EpiEstim::estimate_R(incid = Flu1918$incidence, config = config_month, method = "non_parametric_si")
l <- which(colnames(fit_month$R) == paste0("Quantile.",alpha/2,"(R)"))
u <- which(colnames(fit_month$R) == paste0("Quantile.",1 - alpha/2,"(R)"))
EpiMonth_95Lo_flu <- fill_NAs(fit_month$R[,l], len) 
EpiMonth_95Up_flu <- fill_NAs(fit_month$R[,u], len)
EpiMonth_Rt_flu <- fill_NAs(fit_month$R$`Mean(R)`, len)


# EpiLPS
fitLPS <- EpiLPS::estimR(incidence = Flu1918$incidence, si = flu_delay_distn, K = 40, CoriR = FALSE)
l <- which(colnames(fitLPS$RLPS) == paste0("Rq", alpha/2)) # alpha can only by 0.05, 0.1, 0.5
u <- which(colnames(fitLPS$RLPS) == paste0("Rq", 1 - alpha/2))
EpiLPS_Rt_flu <- fill_NAs(fitLPS$RLPS$R, len)
EpiLPS_95Lo_flu <- fill_NAs(fitLPS$RLPS[,l], len)
EpiLPS_95Up_flu <- fill_NAs(fitLPS$RLPS[,u], len)

# EpiFilter
# cannot use dates as input, but the delays are incorporated by the delay distribution
m <- 2000
pR0 <- (1 / m) * rep(1, m)
Rgrid <- seq(0.01, 10, length.out = m)
w <- delay_calculator(Flu1918$incidence, x = 1:len, delay_distn = flu_delay_distn)
R_filter <- epiFilter(Rgrid, m, eta=0.1, pR0, len, w, Flu1918$incidence, a = alpha)
R_smoother <- epiSmoother(Rgrid, m, R_filter[[4]], R_filter[[5]], len, R_filter[[6]], a = alpha)
EpiFilter_Rt_flu <- R_smoother[[3]]
EpiFilter_95Lo_flu <- R_smoother[[2]][1,]
EpiFilter_95Up_flu <- R_smoother[[2]][2,]

# EpiNow2: use nonparametric SI distribution??
#dates <- seq(from = as.Date("2023-01-01"), by = "day", length.out = len)
#library(EpiNow2)
#out <- EpiNow2::epinow(reported_cases = Flu1918$incidence,
#              generation_time = flu_delay_distn,
#              rt = rt_opts(),
#              delays = delay_opts()
#)
#Rt_fitted <- out$estimates$summarised$mean[1:len]

# combine all results
dat_flu_fitted_long <- data.table(
  Date = rep(1:len, 8),
  Method = rep(c("RtEstim (k=0)", "RtEstim (k=1)", "RtEstim (k=2)", "RtEstim (k=3)", 
                 "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter"), each = len),
  fitted_Rt = c(RtEstim0_Rt_flu, RtEstim1_Rt_flu, RtEstim2_Rt_flu, RtEstim3_Rt_flu, 
                EpiWeek_Rt_flu, EpiMonth_Rt_flu, EpiLPS_Rt_flu, EpiFilter_Rt_flu),
  ci_lo95 = c(RtEstim0_95Lo_flu, RtEstim1_95Lo_flu, RtEstim2_95Lo_flu, RtEstim3_95Lo_flu, 
              EpiWeek_95Lo_flu, EpiMonth_95Lo_flu, EpiLPS_95Lo_flu, EpiFilter_95Lo_flu),
  ci_hi95 = c(RtEstim0_95Up_flu, RtEstim1_95Up_flu, RtEstim2_95Up_flu, RtEstim3_95Up_flu,
              EpiWeek_95Up_flu, EpiMonth_95Up_flu, EpiLPS_95Up_flu, EpiFilter_95Up_flu)
)
dat_flu_fitted_long <- dat_flu_fitted_long %>%
  mutate(Method = fct_relevel(
    Method, 
    "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter",
    "RtEstim (k=0)", "RtEstim (k=1)",
    "RtEstim (k=2)", "RtEstim (k=3)"
  ))
head(dat_flu_fitted_long)
saveRDS(dat_flu_fitted_long, here::here("dat/dat_flu_fitted_long.RDS"))
```

```{r plot-flu-ci,  echo=FALSE, fig.height=8, fig.width=6.2, fig.cap = "Rt estimates with CIs for Flu 1918. Y-axes are truncated beyond 3 for a better display of the fluctuation in small values."}
dat_flu_fitted_long <- readRDS(here::here("dat/dat_flu_fitted_long.RDS"))
dat_flu_fitted_long %>%
  filter(fitted_Rt > 0) %>% # remove NAs marked by -1
  group_by(Method) %>%
  ggplot(aes(x = Date)) + 
  geom_ribbon(aes(ymin = ci_lo95, ymax = ci_hi95, fill = Method), alpha = 0.5) +
  geom_line(aes(y = fitted_Rt, col = Method)) + 
  facet_wrap(~ Method, scales = "free_y", ncol = 1) + 
  coord_cartesian(ylim=c(0, 3)) +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_colour_manual(values = cbPalette[-5]) +
  scale_fill_manual(values = cbPalette[-5]) +
  labs(x = "Days since start of epidemic", y = "") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE), fill = guide_legend(title = NULL, byrow=TRUE))
```

```{r plot-flu-one,  echo=FALSE, fig.height=3.5, fig.width=6.2, fig.cap = "Rt estimates for Flu 1918. Y-axis is truncated beyond 3 for a better display of the fluctuation in small values."}
dat_flu_fitted_long %>%
  filter(fitted_Rt > 0) %>% # remove NAs marked by -1
  group_by(Method) %>%
  ggplot(aes(x = Date)) + 
  geom_line(aes(y = fitted_Rt, col = Method)) + 
  #facet_wrap(~ Method, scales = "free_y", ncol = 1) + 
  geom_hline(yintercept = 1, linetype = "dotted") +
  coord_cartesian(ylim=c(0, 3)) + 
  scale_colour_manual(values = cbPalette[-5]) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  labs(x = "Days since start of epidemic", y = "") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
```


#* References

<!-- The following R chunks reproduce figures in the main manuscript. -->

```{r Fig1-display-cancovid-example, eval = FALSE, include = FALSE}
library(rtestim)
library(Matrix)
n <- nrow(rtestim::cancovid)
can_pred_class <- readRDS(here::here("dat/can_pred_class.RDS"))
delay_dstns_can <- readRDS(here::here("dat/delay-distns-byvar.rds"))
delay_dstns_can <- delay_dstns_can|>
  filter(para == "SI", type %in% unique(can_pred_class$var)) 
delay_mat <- matrix(0, n, n)
delay_mat[1,1] <- 1
for (iter in 2:n) {
  current_var <- can_pred_class$var[iter]
  current_pars <- delay_dstns_can |> filter(type == current_var)
  delay <- discretize_gamma(0:(iter - 1), current_pars$shape, current_pars$scale)
  delay_mat[iter, 1:iter] <- rev(delay)
}
delay_mat <- drop0(as(delay_mat, "CsparseMatrix")) # make it sparse, not necessary
delay_mat <- delay_mat / rowSums(delay_mat) # renormalize


# plot fitted Rt with confidence band
cv_mod <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 2, nfold = 10, 
  delay_distn = delay_mat,
  maxiter = 1e7L, nsol = 50, error_measure = "deviance",
  lambda_min_ratio = 1e-6, init = configure_rt_admm(tol = 1e-6)
)
rt_ci_covid <- confband(cv_mod, "lambda.1se") # get 95% confidence band

start_dates <- data.table(rt_ci_covid, variant = can_pred_class$var, date = rtestim::cancovid$date) %>% group_by(variant) %>% summarize(min(date))

fig_rt_fit <- data.table(rt_ci_covid, variant = can_pred_class$var) %>%
  mutate(variant = fct_relevel(
    variant, 
    "Ancestral lineage", "Alpha", "Delta", "Omicron"
  )) %>%
  ggplot(aes(x = rtestim::cancovid$date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`, fill=variant), alpha = 0.5) +#fill = "gray", 
  geom_line(aes(y = Rt), col = "#0072B2", lwd = 1) +
  scale_x_date(date_breaks = "8 months", date_labels = "%Y-%m-%d", expand = expansion(c(0, 0.05))) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  geom_hline(yintercept = 1, linetype = "dotted") +
  #geom_vline(xintercept = start_dates$`min(date)`[-2], linetype = "dashed") +
  labs(x = "Date", y = "Piecewise quadratic Rt with 95% CIs") +
  coord_cartesian(ylim = c(0, 2)) + 
  theme_bw() +
  scale_fill_brewer(palette = "Dark2", name = "")
fig_rt_fit

# plot fitted incidence
mod_best <- estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 2, maxiter = 1e7L, delay_distn = delay_mat,
  lambda = cv_mod$lambda.1se
)
incidence_fit <- mod_best$weighted_past_counts * rt_ci_covid$Rt
inc_dat <- data.table(rtestim::cancovid, incidence_fit = incidence_fit)

fig_inc_fit <- data.table(inc_dat, variant = can_pred_class$var) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = incident_cases / 1000)) +
  geom_line(aes(y = incidence_fit / 1000), col = "#D55E00", lwd = 1) +
  scale_x_date(date_breaks = "8 months", date_labels = "%Y-%m-%d", expand = expansion(c(0, 0.05))) +
  labs(x = "Date", y = "Fitted incidence (thousands)") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  geom_vline(xintercept = start_dates$`min(date)`[-2], linetype = "dashed") +
  theme_bw()

CANCovid_fig <- ggpubr::ggarrange(
  fig_rt_fit, fig_inc_fit,
  ncol = 1, common.legend = TRUE, legend = "bottom"
)
print(CANCovid_fig)

ggsave(here::here("fig/intro-fig-new.png"), plot = CANCovid_fig, width = 6.5, height = 6.5)
```

```{r Fig2-display-data-examples, eval = FALSE, include = FALSE}
# generate a random sample for each setting (of Rt scenario, incidence distribution, serial interval distribution)
dat1_pois_measles <- data_generator(Rt_case = 1, dist = "poisson", si_type = "measles", seed = 9)
dat2_pois_measles <- data_generator(Rt_case = 2, dist = "poisson", si_type = "measles", seed = 4)
dat3_pois_measles <- data_generator(Rt_case = 3, dist = "poisson", si_type = "measles", seed = 9)
dat4_pois_measles <- data_generator(Rt_case = 4, dist = "poisson", si_type = "measles", seed = 4)
dat1_NB_measles <- data_generator(Rt_case = 1, dist = "NB", si_type = "measles", seed = 3)
dat2_NB_measles <- data_generator(Rt_case = 2, dist = "NB", si_type = "measles", seed = 5)
dat3_NB_measles <- data_generator(Rt_case = 3, dist = "NB", si_type = "measles", seed = 6)
dat4_NB_measles <- data_generator(Rt_case = 4, dist = "NB", si_type = "measles", seed = 4)
dat1_pois_SARS <- data_generator(Rt_case = 1, dist = "poisson", si_type = "SARS", seed = 792)
dat2_pois_SARS <- data_generator(Rt_case = 2, dist = "poisson", si_type = "SARS", seed = 1)
dat3_pois_SARS <- data_generator(Rt_case = 3, dist = "poisson", si_type = "SARS", seed = 792)
dat4_pois_SARS <- data_generator(Rt_case = 4, dist = "poisson", si_type = "SARS", seed = 3)
dat1_NB_SARS <- data_generator(Rt_case = 1, dist = "NB", si_type = "SARS", seed = 86)
dat2_NB_SARS <- data_generator(Rt_case = 2, dist = "NB", si_type = "SARS", seed = 48)
dat3_NB_SARS <- data_generator(Rt_case = 3, dist = "NB", si_type = "SARS", seed = 56)
dat4_NB_SARS <- data_generator(Rt_case = 4, dist = "NB", si_type = "SARS", seed = 29)

# combine the random samples as a data table  
dat_example <- data.table(incidence = c(dat1_pois_measles$incidence, 
                                        dat2_pois_measles$incidence,
                                        dat3_pois_measles$incidence,
                                        dat4_pois_measles$incidence,
                                        dat1_NB_measles$incidence,
                                        dat2_NB_measles$incidence,
                                        dat3_NB_measles$incidence,
                                        dat4_NB_measles$incidence,
                                        dat1_pois_SARS$incidence,
                                        dat2_pois_SARS$incidence,
                                        dat3_pois_SARS$incidence,
                                        dat4_pois_SARS$incidence,
                                        dat1_NB_SARS$incidence,
                                        dat2_NB_SARS$incidence,
                                        dat3_NB_SARS$incidence,
                                        dat4_NB_SARS$incidence),
                          mean = c(dat1_pois_measles$mean, 
                                   dat2_pois_measles$mean,
                                   dat3_pois_measles$mean,
                                   dat4_pois_measles$mean,
                                   dat1_NB_measles$mean,
                                   dat2_NB_measles$mean,
                                   dat3_NB_measles$mean,
                                   dat4_NB_measles$mean,
                                   dat1_pois_SARS$mean,
                                   dat2_pois_SARS$mean,
                                   dat3_pois_SARS$mean,
                                   dat4_pois_SARS$mean,
                                   dat1_NB_SARS$mean,
                                   dat2_NB_SARS$mean,
                                   dat3_NB_SARS$mean,
                                   dat4_NB_SARS$mean),
                          trueRt = rep(c(dat1_pois_measles$Rt,
                                         dat2_pois_measles$Rt,
                                         dat3_pois_measles$Rt,
                                         dat4_pois_measles$Rt), 4),
                          time = rep(1:300, 4 * 4),
                          Rt_case = rep(rep(1:4, each = 300), 4),
                          dist = rep(rep(c("Poisson", "Negative Binomial"), each = 300 * 4), 2),
                          si_type = rep(c("measles", "SARS"), each = 4 * 300 * 2)
                          )
dat_example <- dat_example %>%
  mutate(dist = fct_relevel(
    dist, 
    "Poisson", 
    "Negative Binomial"
  )) %>%
  mutate(Rt_case = as.character(Rt_case)) %>% 
  mutate(Rt_case = fct_recode(
    Rt_case,
    "Scenario 1" = "1",
    "Scenario 2" = "2",
    "Scenario 3" = "3",
    "Scenario 4" = "4"
  ))

# save examples
saveRDS(dat_example, here::here("dat/data_example.RDS"))

# display Rt
fig_true_rt <- dat_example %>%
  mutate(Rt_case = fct_recode(Rt_case,
    "Scenario 1: piecewise constant" = "Scenario 1",
    "Scenario 2: piecewise exponential" = "Scenario 2",
    "Scenario 3: piecewise linear" = "Scenario 3",
    "Scenario 4: periodic" = "Scenario 4"
  )) %>%
  group_by(Rt_case) %>%
  ggplot(aes(y = trueRt, x = time)) +
  geom_line() +
  facet_wrap(~ Rt_case, scales = "free") + #facet_grid
  labs(y = "True Rt curves", x = "Time") +
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  theme_bw()

# display each setting of the random samples
fig_measles_data <- dat_example %>%
  filter(si_type == "measles") %>%
  group_by(Rt_case, dist) %>%
  ggplot(aes(x = time, y = incidence)) +
  geom_line() + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  labs(y = "Incidence of measles epidemics", x = "Time") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  theme_bw()
  
fig_sars_data <- dat_example %>%
  filter(si_type == "SARS") %>%
  group_by(Rt_case, dist) %>%
  ggplot(aes(x = time, y = incidence)) +
  geom_line() + 
  facet_grid(dist ~ Rt_case, scales = "free") +
  labs(y = "Incidence of SARS epidemics", x = "Time") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  theme_bw()

# save plots
ggsave(plot=fig_true_rt, here::here("fig/fig_true_rt.png"), width = 6.5, height = 5)
ggsave(plot=fig_measles_data, here::here("fig/fig_measles_data.png"), width = 6.5, height = 5)
ggsave(plot=fig_sars_data, here::here("fig/fig_sars_data.png"), width = 6.5, height = 5)

# combine the three figures above
dat_example_Rt <- dat_example %>% select(trueRt, mean, time:dist)
dat_example_Rt <- data.table(dat_example_Rt, si_type = "true Rt")
colnames(dat_example_Rt)[1] <- "incidence"
dat_example_inci <- dat_example %>% select(!trueRt)
dat_example_long <- rbind(dat_example_Rt, dat_example_inci)
fig_sample <- dat_example_long %>%
  mutate(dist = fct_relevel(dist, "Negative Binomial", "Poisson")) %>%
  mutate(si_type = fct_relevel(si_type, "true Rt", "measles", "SARS")) %>%
  group_by(Rt_case, dist, si_type) %>%
  ggplot(aes(x = time, y = incidence)) +
  geom_line(aes(color = dist), alpha=0.8) + 
  facet_grid(si_type ~ Rt_case, scales = "free") +
  labs(y = "", x = "Time") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_color_brewer(palette = "Set1") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL))
ggsave(here::here("fig/fig_samples.png"), fig_sample, width = 7.5, height = 6)
```

```{r Figs3+4-display-simulation-results, eval = FALSE, include = FALSE}
# exclude the extremely large KL values (excluding the first weeks)
Rt_res_long_week_no_outliers <- Rt_result %>% 
  filter(len == 300, method != "EpiEstim (monthly)") %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) %>% 
  group_by(Rt_case, dist, method, si_type, si_pars) %>% 
  mutate(big_out = Rt_kl > fivenum(Rt_kl, na.rm = TRUE)[4] +
    1.5 * IQR(Rt_kl, na.rm = TRUE)) %>% 
  filter(!big_out) %>% 
  select(!big_out)

fig_kl_week_measles <- Rt_res_long_week_no_outliers %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) |> 
  group_by(Rt_case, dist, method, si_type, si_pars) |>
  filter(si_type == "measles", si_pars == "measles", method != "EpiNow2") |> # SARS, measles?
  ggplot(aes(x = method, y = Rt_kl)) + 
  stat_boxplot(aes(col = method)) +
  facet_grid(dist ~ Rt_case, scales = "free_y") +
  scale_y_log10() + #labels = scientific_10
  scale_colour_manual(values = cbPalette[c(1,3,4,6:9)]) +
  labs(x = "Method", y = "KL divergence for measles epidemics") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_week_measles)

fig_kl_week_sars <- Rt_res_long_week_no_outliers %>% 
  select(Rt_case, dist, method, Rt_kl, si_type, si_pars) |>
  group_by(Rt_case, dist, method, si_type, si_pars) |>
  filter(si_type == "SARS", si_pars == "SARS") |> # SARS, measles?
  ggplot(aes(x = method, y = Rt_kl)) + # averaged over KL per coordinate
  stat_boxplot(aes(col = method)) +
  facet_grid(dist ~ Rt_case, scales = "free") +
  scale_y_log10() +
  scale_colour_manual(values = cbPalette[c(1,3,4,6:9)]) +
  labs(x = "Method", y = "KL divergence for SARS epidemics") +
  theme_bw() +
  theme(axis.text.x = element_blank(), legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_kl_week_sars)

ggsave(plot=fig_kl_week_measles, here::here("fig/fig_kl_week_measles.png"), width = 6.5, height = 5)
ggsave(plot=fig_kl_week_sars, here::here("fig/fig_kl_week_sars.png"), width = 6.5, height = 5)
```

```{r save-fitted-rt, eval = FALSE, include = FALSE}
## Rt estimates for Poisson incidence by all models
## Commands used to visualize fitted Rt are commented.

# EpiEstim(week) ----
Pois1_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat1_pois_measles)
Pois1_measles_Epiweek_Rt <- Pois1_measles_Epiweek$fitted_Rt
Pois1_measles_Epiweek_95Lo <- Pois1_measles_Epiweek$lower_bound
Pois1_measles_Epiweek_95Up <- Pois1_measles_Epiweek$upper_bound
Pois2_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat2_pois_measles)
Pois2_measles_Epiweek_Rt <- Pois2_measles_Epiweek$fitted_Rt
Pois2_measles_Epiweek_95Lo <- Pois2_measles_Epiweek$lower_bound
Pois2_measles_Epiweek_95Up <- Pois2_measles_Epiweek$upper_bound
Pois3_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat3_pois_measles)
Pois3_measles_Epiweek_Rt <- Pois3_measles_Epiweek$fitted_Rt
Pois3_measles_Epiweek_95Lo <- Pois3_measles_Epiweek$lower_bound
Pois3_measles_Epiweek_95Up <- Pois3_measles_Epiweek$upper_bound
Pois4_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat4_pois_measles)
Pois4_measles_Epiweek_Rt <- Pois4_measles_Epiweek$fitted_Rt
Pois4_measles_Epiweek_95Lo <- Pois4_measles_Epiweek$lower_bound
Pois4_measles_Epiweek_95Up <- Pois4_measles_Epiweek$upper_bound
NB1_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat1_NB_measles)
NB1_measles_Epiweek_Rt <- NB1_measles_Epiweek$fitted_Rt
NB1_measles_Epiweek_95Lo <- NB1_measles_Epiweek$lower_bound
NB1_measles_Epiweek_95Up <- NB1_measles_Epiweek$upper_bound
NB2_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat2_NB_measles)
NB2_measles_Epiweek_Rt <- NB2_measles_Epiweek$fitted_Rt
NB2_measles_Epiweek_95Lo <- NB2_measles_Epiweek$lower_bound
NB2_measles_Epiweek_95Up <- NB2_measles_Epiweek$upper_bound
NB3_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat3_NB_measles)
NB3_measles_Epiweek_Rt <- NB3_measles_Epiweek$fitted_Rt
NB3_measles_Epiweek_95Lo <- NB3_measles_Epiweek$lower_bound
NB3_measles_Epiweek_95Up <- NB3_measles_Epiweek$upper_bound
NB4_measles_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat4_NB_measles)
NB4_measles_Epiweek_Rt <- NB4_measles_Epiweek$fitted_Rt
NB4_measles_Epiweek_95Lo <- NB4_measles_Epiweek$lower_bound
NB4_measles_Epiweek_95Up <- NB4_measles_Epiweek$upper_bound
Pois1_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat1_pois_SARS)
Pois1_SARS_Epiweek_Rt <- Pois1_SARS_Epiweek$fitted_Rt
Pois1_SARS_Epiweek_95Lo <- Pois1_SARS_Epiweek$lower_bound
Pois1_SARS_Epiweek_95Up <- Pois1_SARS_Epiweek$upper_bound
Pois2_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat2_pois_SARS)
Pois2_SARS_Epiweek_Rt <- Pois2_SARS_Epiweek$fitted_Rt
Pois2_SARS_Epiweek_95Lo <- Pois2_SARS_Epiweek$lower_bound
Pois2_SARS_Epiweek_95Up <- Pois2_SARS_Epiweek$upper_bound
Pois3_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat3_pois_SARS)
Pois3_SARS_Epiweek_Rt <- Pois3_SARS_Epiweek$fitted_Rt
Pois3_SARS_Epiweek_95Lo <- Pois3_SARS_Epiweek$lower_bound
Pois3_SARS_Epiweek_95Up <- Pois3_SARS_Epiweek$upper_bound
Pois4_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat4_pois_SARS)
Pois4_SARS_Epiweek_Rt <- Pois4_SARS_Epiweek$fitted_Rt
Pois4_SARS_Epiweek_95Lo <- Pois4_SARS_Epiweek$lower_bound
Pois4_SARS_Epiweek_95Up <- Pois4_SARS_Epiweek$upper_bound
NB1_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat1_NB_SARS)
NB1_SARS_Epiweek_Rt <- NB1_SARS_Epiweek$fitted_Rt
NB1_SARS_Epiweek_95Lo <- NB1_SARS_Epiweek$lower_bound
NB1_SARS_Epiweek_95Up <- NB1_SARS_Epiweek$upper_bound
NB2_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat2_NB_SARS)
NB2_SARS_Epiweek_Rt <- NB2_SARS_Epiweek$fitted_Rt
NB2_SARS_Epiweek_95Lo <- NB2_SARS_Epiweek$lower_bound
NB2_SARS_Epiweek_95Up <- NB2_SARS_Epiweek$upper_bound
NB3_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat3_NB_SARS)
NB3_SARS_Epiweek_Rt <- NB3_SARS_Epiweek$fitted_Rt
NB3_SARS_Epiweek_95Lo <- NB3_SARS_Epiweek$lower_bound
NB3_SARS_Epiweek_95Up <- NB3_SARS_Epiweek$upper_bound
NB4_SARS_Epiweek <- problem_solver(method = 'EpiEstim(week)', instance = dat4_NB_SARS)
NB4_SARS_Epiweek_Rt <- NB4_SARS_Epiweek$fitted_Rt
NB4_SARS_Epiweek_95Lo <- NB4_SARS_Epiweek$lower_bound
NB4_SARS_Epiweek_95Up <- NB4_SARS_Epiweek$upper_bound

Epiweek_res <- c(
  Pois1_measles_Epiweek_Rt, Pois1_measles_Epiweek_95Lo, Pois1_measles_Epiweek_95Up, 
  Pois2_measles_Epiweek_Rt, Pois2_measles_Epiweek_95Lo, Pois2_measles_Epiweek_95Up, 
  Pois3_measles_Epiweek_Rt, Pois3_measles_Epiweek_95Lo, Pois3_measles_Epiweek_95Up, 
  Pois4_measles_Epiweek_Rt, Pois4_measles_Epiweek_95Lo, Pois4_measles_Epiweek_95Up, 
  NB1_measles_Epiweek_Rt, NB1_measles_Epiweek_95Lo, NB1_measles_Epiweek_95Up, 
  NB2_measles_Epiweek_Rt, NB2_measles_Epiweek_95Lo, NB2_measles_Epiweek_95Up, 
  NB3_measles_Epiweek_Rt, NB3_measles_Epiweek_95Lo, NB3_measles_Epiweek_95Up, 
  NB4_measles_Epiweek_Rt, NB4_measles_Epiweek_95Lo, NB4_measles_Epiweek_95Up, 
  Pois1_SARS_Epiweek_Rt, Pois1_SARS_Epiweek_95Lo, Pois1_SARS_Epiweek_95Up, 
  Pois2_SARS_Epiweek_Rt, Pois2_SARS_Epiweek_95Lo, Pois2_SARS_Epiweek_95Up, 
  Pois3_SARS_Epiweek_Rt, Pois3_SARS_Epiweek_95Lo, Pois3_SARS_Epiweek_95Up, 
  Pois4_SARS_Epiweek_Rt, Pois4_SARS_Epiweek_95Lo, Pois4_SARS_Epiweek_95Up, 
  NB1_SARS_Epiweek_Rt, NB1_SARS_Epiweek_95Lo, NB1_SARS_Epiweek_95Up, 
  NB2_SARS_Epiweek_Rt, NB2_SARS_Epiweek_95Lo, NB2_SARS_Epiweek_95Up, 
  NB3_SARS_Epiweek_Rt, NB3_SARS_Epiweek_95Lo, NB3_SARS_Epiweek_95Up, 
  NB4_SARS_Epiweek_Rt, NB4_SARS_Epiweek_95Lo, NB4_SARS_Epiweek_95Up
)
Epiweek_res <- data.table(Res = Epiweek_res,
                         Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))

Epiweek_Rt <- Epiweek_res[Type == "Fitted_Rt", ]$Res
Epiweek_Lo95 <- Epiweek_res[Type == "Lo95", ]$Res
Epiweek_Up95 <- Epiweek_res[Type == "Up95", ]$Res


# EpiEstim(month) ----
Pois1_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat1_pois_measles)
Pois1_measles_Epimonth_Rt <- Pois1_measles_Epimonth$fitted_Rt
Pois1_measles_Epimonth_95Lo <- Pois1_measles_Epimonth$lower_bound
Pois1_measles_Epimonth_95Up <- Pois1_measles_Epimonth$upper_bound
Pois2_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat2_pois_measles)
Pois2_measles_Epimonth_Rt <- Pois2_measles_Epimonth$fitted_Rt
Pois2_measles_Epimonth_95Lo <- Pois2_measles_Epimonth$lower_bound
Pois2_measles_Epimonth_95Up <- Pois2_measles_Epimonth$upper_bound
Pois3_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat3_pois_measles)
Pois3_measles_Epimonth_Rt <- Pois3_measles_Epimonth$fitted_Rt
Pois3_measles_Epimonth_95Lo <- Pois3_measles_Epimonth$lower_bound
Pois3_measles_Epimonth_95Up <- Pois3_measles_Epimonth$upper_bound
Pois4_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat4_pois_measles)
Pois4_measles_Epimonth_Rt <- Pois4_measles_Epimonth$fitted_Rt
Pois4_measles_Epimonth_95Lo <- Pois4_measles_Epimonth$lower_bound
Pois4_measles_Epimonth_95Up <- Pois4_measles_Epimonth$upper_bound
NB1_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat1_NB_measles)
NB1_measles_Epimonth_Rt <- NB1_measles_Epimonth$fitted_Rt
NB1_measles_Epimonth_95Lo <- NB1_measles_Epimonth$lower_bound
NB1_measles_Epimonth_95Up <- NB1_measles_Epimonth$upper_bound
NB2_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat2_NB_measles)
NB2_measles_Epimonth_Rt <- NB2_measles_Epimonth$fitted_Rt
NB2_measles_Epimonth_95Lo <- NB2_measles_Epimonth$lower_bound
NB2_measles_Epimonth_95Up <- NB2_measles_Epimonth$upper_bound
NB3_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat3_NB_measles)
NB3_measles_Epimonth_Rt <- NB3_measles_Epimonth$fitted_Rt
NB3_measles_Epimonth_95Lo <- NB3_measles_Epimonth$lower_bound
NB3_measles_Epimonth_95Up <- NB3_measles_Epimonth$upper_bound
NB4_measles_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat4_NB_measles)
NB4_measles_Epimonth_Rt <- NB4_measles_Epimonth$fitted_Rt
NB4_measles_Epimonth_95Lo <- NB4_measles_Epimonth$lower_bound
NB4_measles_Epimonth_95Up <- NB4_measles_Epimonth$upper_bound
Pois1_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat1_pois_SARS)
Pois1_SARS_Epimonth_Rt <- Pois1_SARS_Epimonth$fitted_Rt
Pois1_SARS_Epimonth_95Lo <- Pois1_SARS_Epimonth$lower_bound
Pois1_SARS_Epimonth_95Up <- Pois1_SARS_Epimonth$upper_bound
Pois2_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat2_pois_SARS)
Pois2_SARS_Epimonth_Rt <- Pois2_SARS_Epimonth$fitted_Rt
Pois2_SARS_Epimonth_95Lo <- Pois2_SARS_Epimonth$lower_bound
Pois2_SARS_Epimonth_95Up <- Pois2_SARS_Epimonth$upper_bound
Pois3_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat3_pois_SARS)
Pois3_SARS_Epimonth_Rt <- Pois3_SARS_Epimonth$fitted_Rt
Pois3_SARS_Epimonth_95Lo <- Pois3_SARS_Epimonth$lower_bound
Pois3_SARS_Epimonth_95Up <- Pois3_SARS_Epimonth$upper_bound
Pois4_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat4_pois_SARS)
Pois4_SARS_Epimonth_Rt <- Pois4_SARS_Epimonth$fitted_Rt
Pois4_SARS_Epimonth_95Lo <- Pois4_SARS_Epimonth$lower_bound
Pois4_SARS_Epimonth_95Up <- Pois4_SARS_Epimonth$upper_bound
NB1_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat1_NB_SARS)
NB1_SARS_Epimonth_Rt <- NB1_SARS_Epimonth$fitted_Rt
NB1_SARS_Epimonth_95Lo <- NB1_SARS_Epimonth$lower_bound
NB1_SARS_Epimonth_95Up <- NB1_SARS_Epimonth$upper_bound
NB2_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat2_NB_SARS)
NB2_SARS_Epimonth_Rt <- NB2_SARS_Epimonth$fitted_Rt
NB2_SARS_Epimonth_95Lo <- NB2_SARS_Epimonth$lower_bound
NB2_SARS_Epimonth_95Up <- NB2_SARS_Epimonth$upper_bound
NB3_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat3_NB_SARS)
NB3_SARS_Epimonth_Rt <- NB3_SARS_Epimonth$fitted_Rt
NB3_SARS_Epimonth_95Lo <- NB3_SARS_Epimonth$lower_bound
NB3_SARS_Epimonth_95Up <- NB3_SARS_Epimonth$upper_bound
NB4_SARS_Epimonth <- problem_solver(method = 'EpiEstim(month)', instance = dat4_NB_SARS)
NB4_SARS_Epimonth_Rt <- NB4_SARS_Epimonth$fitted_Rt
NB4_SARS_Epimonth_95Lo <- NB4_SARS_Epimonth$lower_bound
NB4_SARS_Epimonth_95Up <- NB4_SARS_Epimonth$upper_bound

Epimonth_res <- c(
  Pois1_measles_Epimonth_Rt, Pois1_measles_Epimonth_95Lo, Pois1_measles_Epimonth_95Up,
  Pois2_measles_Epimonth_Rt, Pois2_measles_Epimonth_95Lo, Pois2_measles_Epimonth_95Up,
  Pois3_measles_Epimonth_Rt, Pois3_measles_Epimonth_95Lo, Pois3_measles_Epimonth_95Up,
  Pois4_measles_Epimonth_Rt, Pois4_measles_Epimonth_95Lo, Pois4_measles_Epimonth_95Up, 
  NB1_measles_Epimonth_Rt, NB1_measles_Epimonth_95Lo, NB1_measles_Epimonth_95Up, 
  NB2_measles_Epimonth_Rt, NB2_measles_Epimonth_95Lo, NB2_measles_Epimonth_95Up, 
  NB3_measles_Epimonth_Rt, NB3_measles_Epimonth_95Lo, NB3_measles_Epimonth_95Up, 
  NB4_measles_Epimonth_Rt, NB4_measles_Epimonth_95Lo, NB4_measles_Epimonth_95Up, 
  Pois1_SARS_Epimonth_Rt, Pois1_SARS_Epimonth_95Lo, Pois1_SARS_Epimonth_95Up, 
  Pois2_SARS_Epimonth_Rt, Pois2_SARS_Epimonth_95Lo, Pois2_SARS_Epimonth_95Up, 
  Pois3_SARS_Epimonth_Rt, Pois3_SARS_Epimonth_95Lo, Pois3_SARS_Epimonth_95Up, 
  Pois4_SARS_Epimonth_Rt, Pois4_SARS_Epimonth_95Lo, Pois4_SARS_Epimonth_95Up, 
  NB1_SARS_Epimonth_Rt, NB1_SARS_Epimonth_95Lo, NB1_SARS_Epimonth_95Up, 
  NB2_SARS_Epimonth_Rt, NB2_SARS_Epimonth_95Lo, NB2_SARS_Epimonth_95Up, 
  NB3_SARS_Epimonth_Rt, NB3_SARS_Epimonth_95Lo, NB3_SARS_Epimonth_95Up, 
  NB4_SARS_Epimonth_Rt, NB4_SARS_Epimonth_95Lo, NB4_SARS_Epimonth_95Up
)
Epimonth_res <- data.table(Res = Epimonth_res,
                         Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))

Epimonth_Rt <- Epimonth_res[Type == "Fitted_Rt", ]$Res
Epimonth_Lo95 <- Epimonth_res[Type == "Lo95", ]$Res
Epimonth_Up95 <- Epimonth_res[Type == "Up95", ]$Res



# EpiLPS ----
Pois1_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat1_pois_measles)
Pois1_measles_EpiLPS_Rt <- Pois1_measles_EpiLPS$fitted_Rt
Pois1_measles_EpiLPS_95Lo <- Pois1_measles_EpiLPS$lower_bound
Pois1_measles_EpiLPS_95Up <- Pois1_measles_EpiLPS$upper_bound
Pois2_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat2_pois_measles)
Pois2_measles_EpiLPS_Rt <- Pois2_measles_EpiLPS$fitted_Rt
Pois2_measles_EpiLPS_95Lo <- Pois2_measles_EpiLPS$lower_bound
Pois2_measles_EpiLPS_95Up <- Pois2_measles_EpiLPS$upper_bound
Pois3_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat3_pois_measles)
Pois3_measles_EpiLPS_Rt <- Pois3_measles_EpiLPS$fitted_Rt
Pois3_measles_EpiLPS_95Lo <- Pois3_measles_EpiLPS$lower_bound
Pois3_measles_EpiLPS_95Up <- Pois3_measles_EpiLPS$upper_bound
Pois4_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat4_pois_measles)
Pois4_measles_EpiLPS_Rt <- Pois4_measles_EpiLPS$fitted_Rt
Pois4_measles_EpiLPS_95Lo <- Pois4_measles_EpiLPS$lower_bound
Pois4_measles_EpiLPS_95Up <- Pois4_measles_EpiLPS$upper_bound
NB1_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat1_NB_measles)
NB1_measles_EpiLPS_Rt <- NB1_measles_EpiLPS$fitted_Rt
NB1_measles_EpiLPS_95Lo <- NB1_measles_EpiLPS$lower_bound
NB1_measles_EpiLPS_95Up <- NB1_measles_EpiLPS$upper_bound
NB2_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat2_NB_measles)
NB2_measles_EpiLPS_Rt <- NB2_measles_EpiLPS$fitted_Rt
NB2_measles_EpiLPS_95Lo <- NB2_measles_EpiLPS$lower_bound
NB2_measles_EpiLPS_95Up <- NB2_measles_EpiLPS$upper_bound
NB3_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat3_NB_measles)
NB3_measles_EpiLPS_Rt <- NB3_measles_EpiLPS$fitted_Rt
NB3_measles_EpiLPS_95Lo <- NB3_measles_EpiLPS$lower_bound
NB3_measles_EpiLPS_95Up <- NB3_measles_EpiLPS$upper_bound
NB4_measles_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat4_NB_measles)
NB4_measles_EpiLPS_Rt <- NB4_measles_EpiLPS$fitted_Rt
NB4_measles_EpiLPS_95Lo <- NB4_measles_EpiLPS$lower_bound
NB4_measles_EpiLPS_95Up <- NB4_measles_EpiLPS$upper_bound
Pois1_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat1_pois_SARS)
Pois1_SARS_EpiLPS_Rt <- Pois1_SARS_EpiLPS$fitted_Rt
Pois1_SARS_EpiLPS_95Lo <- Pois1_SARS_EpiLPS$lower_bound
Pois1_SARS_EpiLPS_95Up <- Pois1_SARS_EpiLPS$upper_bound
Pois2_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat2_pois_SARS)
Pois2_SARS_EpiLPS_Rt <- Pois2_SARS_EpiLPS$fitted_Rt
Pois2_SARS_EpiLPS_95Lo <- Pois2_SARS_EpiLPS$lower_bound
Pois2_SARS_EpiLPS_95Up <- Pois2_SARS_EpiLPS$upper_bound
Pois3_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat3_pois_SARS)
Pois3_SARS_EpiLPS_Rt <- Pois3_SARS_EpiLPS$fitted_Rt
Pois3_SARS_EpiLPS_95Lo <- Pois3_SARS_EpiLPS$lower_bound
Pois3_SARS_EpiLPS_95Up <- Pois3_SARS_EpiLPS$upper_bound
Pois4_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat4_pois_SARS)
Pois4_SARS_EpiLPS_Rt <- Pois4_SARS_EpiLPS$fitted_Rt
Pois4_SARS_EpiLPS_95Lo <- Pois4_SARS_EpiLPS$lower_bound
Pois4_SARS_EpiLPS_95Up <- Pois4_SARS_EpiLPS$upper_bound
NB1_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat1_NB_SARS)
NB1_SARS_EpiLPS_Rt <- NB1_SARS_EpiLPS$fitted_Rt
NB1_SARS_EpiLPS_95Lo <- NB1_SARS_EpiLPS$lower_bound
NB1_SARS_EpiLPS_95Up <- NB1_SARS_EpiLPS$upper_bound
NB2_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat2_NB_SARS)
NB2_SARS_EpiLPS_Rt <- NB2_SARS_EpiLPS$fitted_Rt
NB2_SARS_EpiLPS_95Lo <- NB2_SARS_EpiLPS$lower_bound
NB2_SARS_EpiLPS_95Up <- NB2_SARS_EpiLPS$upper_bound
NB3_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat3_NB_SARS)
NB3_SARS_EpiLPS_Rt <- NB3_SARS_EpiLPS$fitted_Rt
NB3_SARS_EpiLPS_95Lo <- NB3_SARS_EpiLPS$lower_bound
NB3_SARS_EpiLPS_95Up <- NB3_SARS_EpiLPS$upper_bound
NB4_SARS_EpiLPS <- problem_solver(method = 'EpiLPS', instance = dat4_NB_SARS)
NB4_SARS_EpiLPS_Rt <- NB4_SARS_EpiLPS$fitted_Rt
NB4_SARS_EpiLPS_95Lo <- NB4_SARS_EpiLPS$lower_bound
NB4_SARS_EpiLPS_95Up <- NB4_SARS_EpiLPS$upper_bound

EpiLPS_res <- c(
  Pois1_measles_EpiLPS_Rt, Pois1_measles_EpiLPS_95Lo, Pois1_measles_EpiLPS_95Up, 
  Pois2_measles_EpiLPS_Rt, Pois2_measles_EpiLPS_95Lo, Pois2_measles_EpiLPS_95Up, 
  Pois3_measles_EpiLPS_Rt, Pois3_measles_EpiLPS_95Lo, Pois3_measles_EpiLPS_95Up, 
  Pois4_measles_EpiLPS_Rt, Pois4_measles_EpiLPS_95Lo, Pois4_measles_EpiLPS_95Up, 
  NB1_measles_EpiLPS_Rt, NB1_measles_EpiLPS_95Lo, NB1_measles_EpiLPS_95Up, 
  NB2_measles_EpiLPS_Rt, NB2_measles_EpiLPS_95Lo, NB2_measles_EpiLPS_95Up, 
  NB3_measles_EpiLPS_Rt, NB3_measles_EpiLPS_95Lo, NB3_measles_EpiLPS_95Up, 
  NB4_measles_EpiLPS_Rt, NB4_measles_EpiLPS_95Lo, NB4_measles_EpiLPS_95Up, 
  Pois1_SARS_EpiLPS_Rt, Pois1_SARS_EpiLPS_95Lo, Pois1_SARS_EpiLPS_95Up, 
  Pois2_SARS_EpiLPS_Rt, Pois2_SARS_EpiLPS_95Lo, Pois2_SARS_EpiLPS_95Up, 
  Pois3_SARS_EpiLPS_Rt, Pois3_SARS_EpiLPS_95Lo, Pois3_SARS_EpiLPS_95Up, 
  Pois4_SARS_EpiLPS_Rt, Pois4_SARS_EpiLPS_95Lo, Pois4_SARS_EpiLPS_95Up, 
  NB1_SARS_EpiLPS_Rt, NB1_SARS_EpiLPS_95Lo, NB1_SARS_EpiLPS_95Up, 
  NB2_SARS_EpiLPS_Rt, NB2_SARS_EpiLPS_95Lo, NB2_SARS_EpiLPS_95Up, 
  NB3_SARS_EpiLPS_Rt, NB3_SARS_EpiLPS_95Lo, NB3_SARS_EpiLPS_95Up, 
  NB4_SARS_EpiLPS_Rt, NB4_SARS_EpiLPS_95Lo, NB4_SARS_EpiLPS_95Up
)
EpiLPS_res <- data.table(Res = EpiLPS_res,
                         Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))

EpiLPS_Rt <- EpiLPS_res[Type == "Fitted_Rt", ]$Res
EpiLPS_Lo95 <- EpiLPS_res[Type == "Lo95", ]$Res
EpiLPS_Up95 <- EpiLPS_res[Type == "Up95", ]$Res


# EpiFilter ----
Pois1_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat1_pois_measles)
Pois1_measles_EpiFilter_Rt <- Pois1_measles_EpiFilter$fitted_Rt
Pois1_measles_EpiFilter_95Lo <- Pois1_measles_EpiFilter$lower_bound
Pois1_measles_EpiFilter_95Up <- Pois1_measles_EpiFilter$upper_bound
Pois2_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat2_pois_measles)
Pois2_measles_EpiFilter_Rt <- Pois2_measles_EpiFilter$fitted_Rt
Pois2_measles_EpiFilter_95Lo <- Pois2_measles_EpiFilter$lower_bound
Pois2_measles_EpiFilter_95Up <- Pois2_measles_EpiFilter$upper_bound
Pois3_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat3_pois_measles)
Pois3_measles_EpiFilter_Rt <- Pois3_measles_EpiFilter$fitted_Rt
Pois3_measles_EpiFilter_95Lo <- Pois3_measles_EpiFilter$lower_bound
Pois3_measles_EpiFilter_95Up <- Pois3_measles_EpiFilter$upper_bound
Pois4_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat4_pois_measles)
Pois4_measles_EpiFilter_Rt <- Pois4_measles_EpiFilter$fitted_Rt
Pois4_measles_EpiFilter_95Lo <- Pois4_measles_EpiFilter$lower_bound
Pois4_measles_EpiFilter_95Up <- Pois4_measles_EpiFilter$upper_bound
NB1_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat1_NB_measles)
NB1_measles_EpiFilter_Rt <- NB1_measles_EpiFilter$fitted_Rt
NB1_measles_EpiFilter_95Lo <- NB1_measles_EpiFilter$lower_bound
NB1_measles_EpiFilter_95Up <- NB1_measles_EpiFilter$upper_bound
NB2_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat2_NB_measles)
NB2_measles_EpiFilter_Rt <- NB2_measles_EpiFilter$fitted_Rt
NB2_measles_EpiFilter_95Lo <- NB2_measles_EpiFilter$lower_bound
NB2_measles_EpiFilter_95Up <- NB2_measles_EpiFilter$upper_bound
NB3_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat3_NB_measles)
NB3_measles_EpiFilter_Rt <- NB3_measles_EpiFilter$fitted_Rt
NB3_measles_EpiFilter_95Lo <- NB3_measles_EpiFilter$lower_bound
NB3_measles_EpiFilter_95Up <- NB3_measles_EpiFilter$upper_bound
NB4_measles_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat4_NB_measles)
NB4_measles_EpiFilter_Rt <- NB4_measles_EpiFilter$fitted_Rt
NB4_measles_EpiFilter_95Lo <- NB4_measles_EpiFilter$lower_bound
NB4_measles_EpiFilter_95Up <- NB4_measles_EpiFilter$upper_bound
Pois1_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat1_pois_SARS)
Pois1_SARS_EpiFilter_Rt <- Pois1_SARS_EpiFilter$fitted_Rt
Pois1_SARS_EpiFilter_95Lo <- Pois1_SARS_EpiFilter$lower_bound
Pois1_SARS_EpiFilter_95Up <- Pois1_SARS_EpiFilter$upper_bound
Pois2_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat2_pois_SARS)
Pois2_SARS_EpiFilter_Rt <- Pois2_SARS_EpiFilter$fitted_Rt
Pois2_SARS_EpiFilter_95Lo <- Pois2_SARS_EpiFilter$lower_bound
Pois2_SARS_EpiFilter_95Up <- Pois2_SARS_EpiFilter$upper_bound
Pois3_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat3_pois_SARS)
Pois3_SARS_EpiFilter_Rt <- Pois3_SARS_EpiFilter$fitted_Rt
Pois3_SARS_EpiFilter_95Lo <- Pois3_SARS_EpiFilter$lower_bound
Pois3_SARS_EpiFilter_95Up <- Pois3_SARS_EpiFilter$upper_bound
Pois4_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat4_pois_SARS)
Pois4_SARS_EpiFilter_Rt <- Pois4_SARS_EpiFilter$fitted_Rt
Pois4_SARS_EpiFilter_95Lo <- Pois4_SARS_EpiFilter$lower_bound
Pois4_SARS_EpiFilter_95Up <- Pois4_SARS_EpiFilter$upper_bound
NB1_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat1_NB_SARS)
NB1_SARS_EpiFilter_Rt <- NB1_SARS_EpiFilter$fitted_Rt
NB1_SARS_EpiFilter_95Lo <- NB1_SARS_EpiFilter$lower_bound
NB1_SARS_EpiFilter_95Up <- NB1_SARS_EpiFilter$upper_bound
NB2_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat2_NB_SARS)
NB2_SARS_EpiFilter_Rt <- NB2_SARS_EpiFilter$fitted_Rt
NB2_SARS_EpiFilter_95Lo <- NB2_SARS_EpiFilter$lower_bound
NB2_SARS_EpiFilter_95Up <- NB2_SARS_EpiFilter$upper_bound
NB3_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat3_NB_SARS)
NB3_SARS_EpiFilter_Rt <- NB3_SARS_EpiFilter$fitted_Rt
NB3_SARS_EpiFilter_95Lo <- NB3_SARS_EpiFilter$lower_bound
NB3_SARS_EpiFilter_95Up <- NB3_SARS_EpiFilter$upper_bound
NB4_SARS_EpiFilter <- problem_solver(method = 'EpiFilter', instance = dat4_NB_SARS)
NB4_SARS_EpiFilter_Rt <- NB4_SARS_EpiFilter$fitted_Rt
NB4_SARS_EpiFilter_95Lo <- NB4_SARS_EpiFilter$lower_bound
NB4_SARS_EpiFilter_95Up <- NB4_SARS_EpiFilter$upper_bound

EpiFilter_res <- c(
  Pois1_measles_EpiFilter_Rt, Pois1_measles_EpiFilter_95Lo, Pois1_measles_EpiFilter_95Up,
  Pois2_measles_EpiFilter_Rt, Pois2_measles_EpiFilter_95Lo, Pois2_measles_EpiFilter_95Up,
  Pois3_measles_EpiFilter_Rt, Pois3_measles_EpiFilter_95Lo, Pois3_measles_EpiFilter_95Up,
  Pois4_measles_EpiFilter_Rt, Pois4_measles_EpiFilter_95Lo, Pois4_measles_EpiFilter_95Up,
  NB1_measles_EpiFilter_Rt, NB1_measles_EpiFilter_95Lo, NB1_measles_EpiFilter_95Up, 
  NB2_measles_EpiFilter_Rt, NB2_measles_EpiFilter_95Lo, NB2_measles_EpiFilter_95Up, 
  NB3_measles_EpiFilter_Rt, NB3_measles_EpiFilter_95Lo, NB3_measles_EpiFilter_95Up, 
  NB4_measles_EpiFilter_Rt, NB4_measles_EpiFilter_95Lo, NB4_measles_EpiFilter_95Up, 
  Pois1_SARS_EpiFilter_Rt, Pois1_SARS_EpiFilter_95Lo, Pois1_SARS_EpiFilter_95Up, 
  Pois2_SARS_EpiFilter_Rt, Pois2_SARS_EpiFilter_95Lo, Pois2_SARS_EpiFilter_95Up, 
  Pois3_SARS_EpiFilter_Rt, Pois3_SARS_EpiFilter_95Lo, Pois3_SARS_EpiFilter_95Up, 
  Pois4_SARS_EpiFilter_Rt, Pois4_SARS_EpiFilter_95Lo, Pois4_SARS_EpiFilter_95Up, 
  NB1_SARS_EpiFilter_Rt, NB1_SARS_EpiFilter_95Lo, NB1_SARS_EpiFilter_95Up, 
  NB2_SARS_EpiFilter_Rt, NB2_SARS_EpiFilter_95Lo, NB2_SARS_EpiFilter_95Up, 
  NB3_SARS_EpiFilter_Rt, NB3_SARS_EpiFilter_95Lo, NB3_SARS_EpiFilter_95Up, 
  NB4_SARS_EpiFilter_Rt, NB4_SARS_EpiFilter_95Lo, NB4_SARS_EpiFilter_95Up
)
EpiFilter_res <- data.table(Res = EpiFilter_res,
                           Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))
EpiFilter_Rt <- EpiFilter_res[Type == "Fitted_Rt", ]$Res
EpiFilter_Lo95 <- EpiFilter_res[Type == "Lo95", ]$Res
EpiFilter_Up95 <- EpiFilter_res[Type == "Up95", ]$Res


# RtEstim(k=0) ----
Pois1_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat1_pois_measles)
Pois1_measles_RtEstim0_Rt <- Pois1_measles_RtEstim0$fitted_Rt
Pois1_measles_RtEstim0_95Lo <- Pois1_measles_RtEstim0$lower_bound
Pois1_measles_RtEstim0_95Up <- Pois1_measles_RtEstim0$upper_bound
Pois2_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat2_pois_measles)
Pois2_measles_RtEstim0_Rt <- Pois2_measles_RtEstim0$fitted_Rt
Pois2_measles_RtEstim0_95Lo <- Pois2_measles_RtEstim0$lower_bound
Pois2_measles_RtEstim0_95Up <- Pois2_measles_RtEstim0$upper_bound
Pois3_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat3_pois_measles)
Pois3_measles_RtEstim0_Rt <- Pois3_measles_RtEstim0$fitted_Rt
Pois3_measles_RtEstim0_95Lo <- Pois3_measles_RtEstim0$lower_bound
Pois3_measles_RtEstim0_95Up <- Pois3_measles_RtEstim0$upper_bound
Pois4_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat4_pois_measles)
Pois4_measles_RtEstim0_Rt <- Pois4_measles_RtEstim0$fitted_Rt
Pois4_measles_RtEstim0_95Lo <- Pois4_measles_RtEstim0$lower_bound
Pois4_measles_RtEstim0_95Up <- Pois4_measles_RtEstim0$upper_bound
NB1_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat1_NB_measles)
NB1_measles_RtEstim0_Rt <- NB1_measles_RtEstim0$fitted_Rt
NB1_measles_RtEstim0_95Lo <- NB1_measles_RtEstim0$lower_bound
NB1_measles_RtEstim0_95Up <- NB1_measles_RtEstim0$upper_bound
NB2_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat2_NB_measles)
NB2_measles_RtEstim0_Rt <- NB2_measles_RtEstim0$fitted_Rt
NB2_measles_RtEstim0_95Lo <- NB2_measles_RtEstim0$lower_bound
NB2_measles_RtEstim0_95Up <- NB2_measles_RtEstim0$upper_bound
NB3_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat3_NB_measles)
NB3_measles_RtEstim0_Rt <- NB3_measles_RtEstim0$fitted_Rt
NB3_measles_RtEstim0_95Lo <- NB3_measles_RtEstim0$lower_bound
NB3_measles_RtEstim0_95Up <- NB3_measles_RtEstim0$upper_bound
NB4_measles_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat4_NB_measles)
NB4_measles_RtEstim0_Rt <- NB4_measles_RtEstim0$fitted_Rt
NB4_measles_RtEstim0_95Lo <- NB4_measles_RtEstim0$lower_bound
NB4_measles_RtEstim0_95Up <- NB4_measles_RtEstim0$upper_bound
Pois1_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat1_pois_SARS)
Pois1_SARS_RtEstim0_Rt <- Pois1_SARS_RtEstim0$fitted_Rt
Pois1_SARS_RtEstim0_95Lo <- Pois1_SARS_RtEstim0$lower_bound
Pois1_SARS_RtEstim0_95Up <- Pois1_SARS_RtEstim0$upper_bound
Pois2_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat2_pois_SARS)
Pois2_SARS_RtEstim0_Rt <- Pois2_SARS_RtEstim0$fitted_Rt
Pois2_SARS_RtEstim0_95Lo <- Pois2_SARS_RtEstim0$lower_bound
Pois2_SARS_RtEstim0_95Up <- Pois2_SARS_RtEstim0$upper_bound
Pois3_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat3_pois_SARS)
Pois3_SARS_RtEstim0_Rt <- Pois3_SARS_RtEstim0$fitted_Rt
Pois3_SARS_RtEstim0_95Lo <- Pois3_SARS_RtEstim0$lower_bound
Pois3_SARS_RtEstim0_95Up <- Pois3_SARS_RtEstim0$upper_bound
Pois4_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat4_pois_SARS)
Pois4_SARS_RtEstim0_Rt <- Pois4_SARS_RtEstim0$fitted_Rt
Pois4_SARS_RtEstim0_95Lo <- Pois4_SARS_RtEstim0$lower_bound
Pois4_SARS_RtEstim0_95Up <- Pois4_SARS_RtEstim0$upper_bound
NB1_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat1_NB_SARS)
NB1_SARS_RtEstim0_Rt <- NB1_SARS_RtEstim0$fitted_Rt
NB1_SARS_RtEstim0_95Lo <- NB1_SARS_RtEstim0$lower_bound
NB1_SARS_RtEstim0_95Up <- NB1_SARS_RtEstim0$upper_bound
NB2_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat2_NB_SARS)
NB2_SARS_RtEstim0_Rt <- NB2_SARS_RtEstim0$fitted_Rt
NB2_SARS_RtEstim0_95Lo <- NB2_SARS_RtEstim0$lower_bound
NB2_SARS_RtEstim0_95Up <- NB2_SARS_RtEstim0$upper_bound
NB3_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat3_NB_SARS)
NB3_SARS_RtEstim0_Rt <- NB3_SARS_RtEstim0$fitted_Rt
NB3_SARS_RtEstim0_95Lo <- NB3_SARS_RtEstim0$lower_bound
NB3_SARS_RtEstim0_95Up <- NB3_SARS_RtEstim0$upper_bound
NB4_SARS_RtEstim0 <- problem_solver(method = 'RtEstim(k=0)', instance = dat4_NB_SARS)
NB4_SARS_RtEstim0_Rt <- NB4_SARS_RtEstim0$fitted_Rt
NB4_SARS_RtEstim0_95Lo <- NB4_SARS_RtEstim0$lower_bound
NB4_SARS_RtEstim0_95Up <- NB4_SARS_RtEstim0$upper_bound

RtEstim0_res <- c(
  Pois1_measles_RtEstim0_Rt, Pois1_measles_RtEstim0_95Lo, Pois1_measles_RtEstim0_95Up,
  Pois2_measles_RtEstim0_Rt, Pois2_measles_RtEstim0_95Lo, Pois2_measles_RtEstim0_95Up,
  Pois3_measles_RtEstim0_Rt, Pois3_measles_RtEstim0_95Lo, Pois3_measles_RtEstim0_95Up,
  Pois4_measles_RtEstim0_Rt, Pois4_measles_RtEstim0_95Lo, Pois4_measles_RtEstim0_95Up, 
  NB1_measles_RtEstim0_Rt, NB1_measles_RtEstim0_95Lo, NB1_measles_RtEstim0_95Up, 
  NB2_measles_RtEstim0_Rt, NB2_measles_RtEstim0_95Lo, NB2_measles_RtEstim0_95Up, 
  NB3_measles_RtEstim0_Rt, NB3_measles_RtEstim0_95Lo, NB3_measles_RtEstim0_95Up, 
  NB4_measles_RtEstim0_Rt, NB4_measles_RtEstim0_95Lo, NB4_measles_RtEstim0_95Up, 
  Pois1_SARS_RtEstim0_Rt, Pois1_SARS_RtEstim0_95Lo, Pois1_SARS_RtEstim0_95Up, 
  Pois2_SARS_RtEstim0_Rt, Pois2_SARS_RtEstim0_95Lo, Pois2_SARS_RtEstim0_95Up, 
  Pois3_SARS_RtEstim0_Rt, Pois3_SARS_RtEstim0_95Lo, Pois3_SARS_RtEstim0_95Up, 
  Pois4_SARS_RtEstim0_Rt, Pois4_SARS_RtEstim0_95Lo, Pois4_SARS_RtEstim0_95Up, 
  NB1_SARS_RtEstim0_Rt, NB1_SARS_RtEstim0_95Lo, NB1_SARS_RtEstim0_95Up, 
  NB2_SARS_RtEstim0_Rt, NB2_SARS_RtEstim0_95Lo, NB2_SARS_RtEstim0_95Up, 
  NB3_SARS_RtEstim0_Rt, NB3_SARS_RtEstim0_95Lo, NB3_SARS_RtEstim0_95Up, 
  NB4_SARS_RtEstim0_Rt, NB4_SARS_RtEstim0_95Lo, NB4_SARS_RtEstim0_95Up
)
RtEstim0_res <- data.table(Res = RtEstim0_res,
                           Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))
RtEstim0_Rt <- RtEstim0_res[Type == "Fitted_Rt", ]$Res
RtEstim0_Lo95 <- RtEstim0_res[Type == "Lo95", ]$Res
RtEstim0_Up95 <- RtEstim0_res[Type == "Up95", ]$Res


# RtEstim(k=1) ----
Pois1_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat1_pois_measles)
Pois1_measles_RtEstim1_Rt <- Pois1_measles_RtEstim1$fitted_Rt
Pois1_measles_RtEstim1_95Lo <- Pois1_measles_RtEstim1$lower_bound
Pois1_measles_RtEstim1_95Up <- Pois1_measles_RtEstim1$upper_bound
Pois2_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat2_pois_measles)
Pois2_measles_RtEstim1_Rt <- Pois2_measles_RtEstim1$fitted_Rt
Pois2_measles_RtEstim1_95Lo <- Pois2_measles_RtEstim1$lower_bound
Pois2_measles_RtEstim1_95Up <- Pois2_measles_RtEstim1$upper_bound
Pois3_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat3_pois_measles)
Pois3_measles_RtEstim1_Rt <- Pois3_measles_RtEstim1$fitted_Rt
Pois3_measles_RtEstim1_95Lo <- Pois3_measles_RtEstim1$lower_bound
Pois3_measles_RtEstim1_95Up <- Pois3_measles_RtEstim1$upper_bound
Pois4_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat4_pois_measles)
Pois4_measles_RtEstim1_Rt <- Pois4_measles_RtEstim1$fitted_Rt
Pois4_measles_RtEstim1_95Lo <- Pois4_measles_RtEstim1$lower_bound
Pois4_measles_RtEstim1_95Up <- Pois4_measles_RtEstim1$upper_bound
NB1_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat1_NB_measles)
NB1_measles_RtEstim1_Rt <- NB1_measles_RtEstim1$fitted_Rt
NB1_measles_RtEstim1_95Lo <- NB1_measles_RtEstim1$lower_bound
NB1_measles_RtEstim1_95Up <- NB1_measles_RtEstim1$upper_bound
NB2_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat2_NB_measles)
NB2_measles_RtEstim1_Rt <- NB2_measles_RtEstim1$fitted_Rt
NB2_measles_RtEstim1_95Lo <- NB2_measles_RtEstim1$lower_bound
NB2_measles_RtEstim1_95Up <- NB2_measles_RtEstim1$upper_bound
NB3_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat3_NB_measles)
NB3_measles_RtEstim1_Rt <- NB3_measles_RtEstim1$fitted_Rt
NB3_measles_RtEstim1_95Lo <- NB3_measles_RtEstim1$lower_bound
NB3_measles_RtEstim1_95Up <- NB3_measles_RtEstim1$upper_bound
NB4_measles_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat4_NB_measles)
NB4_measles_RtEstim1_Rt <- NB4_measles_RtEstim1$fitted_Rt
NB4_measles_RtEstim1_95Lo <- NB4_measles_RtEstim1$lower_bound
NB4_measles_RtEstim1_95Up <- NB4_measles_RtEstim1$upper_bound
Pois1_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat1_pois_SARS)
Pois1_SARS_RtEstim1_Rt <- Pois1_SARS_RtEstim1$fitted_Rt
Pois1_SARS_RtEstim1_95Lo <- Pois1_SARS_RtEstim1$lower_bound
Pois1_SARS_RtEstim1_95Up <- Pois1_SARS_RtEstim1$upper_bound
Pois2_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat2_pois_SARS)
Pois2_SARS_RtEstim1_Rt <- Pois2_SARS_RtEstim1$fitted_Rt
Pois2_SARS_RtEstim1_95Lo <- Pois2_SARS_RtEstim1$lower_bound
Pois2_SARS_RtEstim1_95Up <- Pois2_SARS_RtEstim1$upper_bound
Pois3_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat3_pois_SARS)
Pois3_SARS_RtEstim1_Rt <- Pois3_SARS_RtEstim1$fitted_Rt
Pois3_SARS_RtEstim1_95Lo <- Pois3_SARS_RtEstim1$lower_bound
Pois3_SARS_RtEstim1_95Up <- Pois3_SARS_RtEstim1$upper_bound
Pois4_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat4_pois_SARS)
Pois4_SARS_RtEstim1_Rt <- Pois4_SARS_RtEstim1$fitted_Rt
Pois4_SARS_RtEstim1_95Lo <- Pois4_SARS_RtEstim1$lower_bound
Pois4_SARS_RtEstim1_95Up <- Pois4_SARS_RtEstim1$upper_bound
NB1_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat1_NB_SARS)
NB1_SARS_RtEstim1_Rt <- NB1_SARS_RtEstim1$fitted_Rt
NB1_SARS_RtEstim1_95Lo <- NB1_SARS_RtEstim1$lower_bound
NB1_SARS_RtEstim1_95Up <- NB1_SARS_RtEstim1$upper_bound
NB2_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat2_NB_SARS)
NB2_SARS_RtEstim1_Rt <- NB2_SARS_RtEstim1$fitted_Rt
NB2_SARS_RtEstim1_95Lo <- NB2_SARS_RtEstim1$lower_bound
NB2_SARS_RtEstim1_95Up <- NB2_SARS_RtEstim1$upper_bound
NB3_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat3_NB_SARS)
NB3_SARS_RtEstim1_Rt <- NB3_SARS_RtEstim1$fitted_Rt
NB3_SARS_RtEstim1_95Lo <- NB3_SARS_RtEstim1$lower_bound
NB3_SARS_RtEstim1_95Up <- NB3_SARS_RtEstim1$upper_bound
NB4_SARS_RtEstim1 <- problem_solver(method = 'RtEstim(k=1)', instance = dat4_NB_SARS)
NB4_SARS_RtEstim1_Rt <- NB4_SARS_RtEstim1$fitted_Rt
NB4_SARS_RtEstim1_95Lo <- NB4_SARS_RtEstim1$lower_bound
NB4_SARS_RtEstim1_95Up <- NB4_SARS_RtEstim1$upper_bound


RtEstim1_res <- c(
  Pois1_measles_RtEstim1_Rt, Pois1_measles_RtEstim1_95Lo, Pois1_measles_RtEstim1_95Up,
  Pois2_measles_RtEstim1_Rt, Pois2_measles_RtEstim1_95Lo, Pois2_measles_RtEstim1_95Up,
  Pois3_measles_RtEstim1_Rt, Pois3_measles_RtEstim1_95Lo, Pois3_measles_RtEstim1_95Up,
  Pois4_measles_RtEstim1_Rt, Pois4_measles_RtEstim1_95Lo, Pois4_measles_RtEstim1_95Up, 
  NB1_measles_RtEstim1_Rt, NB1_measles_RtEstim1_95Lo, NB1_measles_RtEstim1_95Up, 
  NB2_measles_RtEstim1_Rt, NB2_measles_RtEstim1_95Lo, NB2_measles_RtEstim1_95Up, 
  NB3_measles_RtEstim1_Rt, NB3_measles_RtEstim1_95Lo, NB3_measles_RtEstim1_95Up, 
  NB4_measles_RtEstim1_Rt, NB4_measles_RtEstim1_95Lo, NB4_measles_RtEstim1_95Up, 
  Pois1_SARS_RtEstim1_Rt, Pois1_SARS_RtEstim1_95Lo, Pois1_SARS_RtEstim1_95Up, 
  Pois2_SARS_RtEstim1_Rt, Pois2_SARS_RtEstim1_95Lo, Pois2_SARS_RtEstim1_95Up, 
  Pois3_SARS_RtEstim1_Rt, Pois3_SARS_RtEstim1_95Lo, Pois3_SARS_RtEstim1_95Up, 
  Pois4_SARS_RtEstim1_Rt, Pois4_SARS_RtEstim1_95Lo, Pois4_SARS_RtEstim1_95Up, 
  NB1_SARS_RtEstim1_Rt, NB1_SARS_RtEstim1_95Lo, NB1_SARS_RtEstim1_95Up, 
  NB2_SARS_RtEstim1_Rt, NB2_SARS_RtEstim1_95Lo, NB2_SARS_RtEstim1_95Up, 
  NB3_SARS_RtEstim1_Rt, NB3_SARS_RtEstim1_95Lo, NB3_SARS_RtEstim1_95Up, 
  NB4_SARS_RtEstim1_Rt, NB4_SARS_RtEstim1_95Lo, NB4_SARS_RtEstim1_95Up
)
RtEstim1_res <- data.table(Res = RtEstim1_res,
                           Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))
RtEstim1_Rt <- RtEstim1_res[Type == "Fitted_Rt", ]$Res
RtEstim1_Lo95 <- RtEstim1_res[Type == "Lo95", ]$Res
RtEstim1_Up95 <- RtEstim1_res[Type == "Up95", ]$Res


# RtEstim(k=2) ----
Pois1_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat1_pois_measles)
Pois1_measles_RtEstim2_Rt <- Pois1_measles_RtEstim2$fitted_Rt
Pois1_measles_RtEstim2_95Lo <- Pois1_measles_RtEstim2$lower_bound
Pois1_measles_RtEstim2_95Up <- Pois1_measles_RtEstim2$upper_bound
Pois2_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat2_pois_measles)
Pois2_measles_RtEstim2_Rt <- Pois2_measles_RtEstim2$fitted_Rt
Pois2_measles_RtEstim2_95Lo <- Pois2_measles_RtEstim2$lower_bound
Pois2_measles_RtEstim2_95Up <- Pois2_measles_RtEstim2$upper_bound
Pois3_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat3_pois_measles)
Pois3_measles_RtEstim2_Rt <- Pois3_measles_RtEstim2$fitted_Rt
Pois3_measles_RtEstim2_95Lo <- Pois3_measles_RtEstim2$lower_bound
Pois3_measles_RtEstim2_95Up <- Pois3_measles_RtEstim2$upper_bound
Pois4_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat4_pois_measles)
Pois4_measles_RtEstim2_Rt <- Pois4_measles_RtEstim2$fitted_Rt
Pois4_measles_RtEstim2_95Lo <- Pois4_measles_RtEstim2$lower_bound
Pois4_measles_RtEstim2_95Up <- Pois4_measles_RtEstim2$upper_bound
NB1_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat1_NB_measles)
NB1_measles_RtEstim2_Rt <- NB1_measles_RtEstim2$fitted_Rt
NB1_measles_RtEstim2_95Lo <- NB1_measles_RtEstim2$lower_bound
NB1_measles_RtEstim2_95Up <- NB1_measles_RtEstim2$upper_bound
NB2_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat2_NB_measles)
NB2_measles_RtEstim2_Rt <- NB2_measles_RtEstim2$fitted_Rt
NB2_measles_RtEstim2_95Lo <- NB2_measles_RtEstim2$lower_bound
NB2_measles_RtEstim2_95Up <- NB2_measles_RtEstim2$upper_bound
NB3_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat3_NB_measles)
NB3_measles_RtEstim2_Rt <- NB3_measles_RtEstim2$fitted_Rt
NB3_measles_RtEstim2_95Lo <- NB3_measles_RtEstim2$lower_bound
NB3_measles_RtEstim2_95Up <- NB3_measles_RtEstim2$upper_bound
NB4_measles_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat4_NB_measles)
NB4_measles_RtEstim2_Rt <- NB4_measles_RtEstim2$fitted_Rt
NB4_measles_RtEstim2_95Lo <- NB4_measles_RtEstim2$lower_bound
NB4_measles_RtEstim2_95Up <- NB4_measles_RtEstim2$upper_bound
Pois1_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat1_pois_SARS)
Pois1_SARS_RtEstim2_Rt <- Pois1_SARS_RtEstim2$fitted_Rt
Pois1_SARS_RtEstim2_95Lo <- Pois1_SARS_RtEstim2$lower_bound
Pois1_SARS_RtEstim2_95Up <- Pois1_SARS_RtEstim2$upper_bound
Pois2_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat2_pois_SARS)
Pois2_SARS_RtEstim2_Rt <- Pois2_SARS_RtEstim2$fitted_Rt
Pois2_SARS_RtEstim2_95Lo <- Pois2_SARS_RtEstim2$lower_bound
Pois2_SARS_RtEstim2_95Up <- Pois2_SARS_RtEstim2$upper_bound
Pois3_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat3_pois_SARS)
Pois3_SARS_RtEstim2_Rt <- Pois3_SARS_RtEstim2$fitted_Rt
Pois3_SARS_RtEstim2_95Lo <- Pois3_SARS_RtEstim2$lower_bound
Pois3_SARS_RtEstim2_95Up <- Pois3_SARS_RtEstim2$upper_bound
Pois4_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat4_pois_SARS)
Pois4_SARS_RtEstim2_Rt <- Pois4_SARS_RtEstim2$fitted_Rt
Pois4_SARS_RtEstim2_95Lo <- Pois4_SARS_RtEstim2$lower_bound
Pois4_SARS_RtEstim2_95Up <- Pois4_SARS_RtEstim2$upper_bound
NB1_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat1_NB_SARS)
NB1_SARS_RtEstim2_Rt <- NB1_SARS_RtEstim2$fitted_Rt
NB1_SARS_RtEstim2_95Lo <- NB1_SARS_RtEstim2$lower_bound
NB1_SARS_RtEstim2_95Up <- NB1_SARS_RtEstim2$upper_bound
NB2_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat2_NB_SARS)
NB2_SARS_RtEstim2_Rt <- NB2_SARS_RtEstim2$fitted_Rt
NB2_SARS_RtEstim2_95Lo <- NB2_SARS_RtEstim2$lower_bound
NB2_SARS_RtEstim2_95Up <- NB2_SARS_RtEstim2$upper_bound
NB3_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat3_NB_SARS)
NB3_SARS_RtEstim2_Rt <- NB3_SARS_RtEstim2$fitted_Rt
NB3_SARS_RtEstim2_95Lo <- NB3_SARS_RtEstim2$lower_bound
NB3_SARS_RtEstim2_95Up <- NB3_SARS_RtEstim2$upper_bound
NB4_SARS_RtEstim2 <- problem_solver(method = 'RtEstim(k=2)', instance = dat4_NB_SARS)
NB4_SARS_RtEstim2_Rt <- NB4_SARS_RtEstim2$fitted_Rt
NB4_SARS_RtEstim2_95Lo <- NB4_SARS_RtEstim2$lower_bound
NB4_SARS_RtEstim2_95Up <- NB4_SARS_RtEstim2$upper_bound

RtEstim2_res <- c( 
  Pois1_measles_RtEstim2_Rt, Pois1_measles_RtEstim2_95Lo, Pois1_measles_RtEstim2_95Up,
  Pois2_measles_RtEstim2_Rt, Pois2_measles_RtEstim2_95Lo, Pois2_measles_RtEstim2_95Up,
  Pois3_measles_RtEstim2_Rt, Pois3_measles_RtEstim2_95Lo, Pois3_measles_RtEstim2_95Up,
  Pois4_measles_RtEstim2_Rt, Pois4_measles_RtEstim2_95Lo, Pois4_measles_RtEstim2_95Up,
  NB1_measles_RtEstim2_Rt, NB1_measles_RtEstim2_95Lo, NB1_measles_RtEstim2_95Up,
  NB2_measles_RtEstim2_Rt, NB2_measles_RtEstim2_95Lo, NB2_measles_RtEstim2_95Up, 
  NB3_measles_RtEstim2_Rt, NB3_measles_RtEstim2_95Lo, NB3_measles_RtEstim2_95Up,
  NB4_measles_RtEstim2_Rt, NB4_measles_RtEstim2_95Lo, NB4_measles_RtEstim2_95Up,
  Pois1_SARS_RtEstim2_Rt, Pois1_SARS_RtEstim2_95Lo, Pois1_SARS_RtEstim2_95Up,
  Pois2_SARS_RtEstim2_Rt, Pois2_SARS_RtEstim2_95Lo, Pois2_SARS_RtEstim2_95Up,
  Pois3_SARS_RtEstim2_Rt, Pois3_SARS_RtEstim2_95Lo, Pois3_SARS_RtEstim2_95Up,
  Pois4_SARS_RtEstim2_Rt, Pois4_SARS_RtEstim2_95Lo, Pois4_SARS_RtEstim2_95Up, 
  NB1_SARS_RtEstim2_Rt, NB1_SARS_RtEstim2_95Lo, NB1_SARS_RtEstim2_95Up, 
  NB2_SARS_RtEstim2_Rt, NB2_SARS_RtEstim2_95Lo, NB2_SARS_RtEstim2_95Up, 
  NB3_SARS_RtEstim2_Rt, NB3_SARS_RtEstim2_95Lo, NB3_SARS_RtEstim2_95Up, 
  NB4_SARS_RtEstim2_Rt, NB4_SARS_RtEstim2_95Lo, NB4_SARS_RtEstim2_95Up
)

RtEstim2_res <- data.table(Res = RtEstim2_res,
                           Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))
RtEstim2_Rt <- RtEstim2_res[Type == "Fitted_Rt", ]$Res
RtEstim2_Lo95 <- RtEstim2_res[Type == "Lo95", ]$Res
RtEstim2_Up95 <- RtEstim2_res[Type == "Up95", ]$Res

# RtEstim(k=3) ----
Pois1_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat1_pois_measles)
Pois1_measles_RtEstim3_Rt <- Pois1_measles_RtEstim3$fitted_Rt
Pois1_measles_RtEstim3_95Lo <- Pois1_measles_RtEstim3$lower_bound
Pois1_measles_RtEstim3_95Up <- Pois1_measles_RtEstim3$upper_bound
Pois2_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat2_pois_measles)
Pois2_measles_RtEstim3_Rt <- Pois2_measles_RtEstim3$fitted_Rt
Pois2_measles_RtEstim3_95Lo <- Pois2_measles_RtEstim3$lower_bound
Pois2_measles_RtEstim3_95Up <- Pois2_measles_RtEstim3$upper_bound
Pois3_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat3_pois_measles)
Pois3_measles_RtEstim3_Rt <- Pois3_measles_RtEstim3$fitted_Rt
Pois3_measles_RtEstim3_95Lo <- Pois3_measles_RtEstim3$lower_bound
Pois3_measles_RtEstim3_95Up <- Pois3_measles_RtEstim3$upper_bound
Pois4_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat4_pois_measles)
Pois4_measles_RtEstim3_Rt <- Pois4_measles_RtEstim3$fitted_Rt
Pois4_measles_RtEstim3_95Lo <- Pois4_measles_RtEstim3$lower_bound
Pois4_measles_RtEstim3_95Up <- Pois4_measles_RtEstim3$upper_bound
NB1_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat1_NB_measles)
NB1_measles_RtEstim3_Rt <- NB1_measles_RtEstim3$fitted_Rt
NB1_measles_RtEstim3_95Lo <- NB1_measles_RtEstim3$lower_bound
NB1_measles_RtEstim3_95Up <- NB1_measles_RtEstim3$upper_bound
NB2_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat2_NB_measles)
NB2_measles_RtEstim3_Rt <- NB2_measles_RtEstim3$fitted_Rt
NB2_measles_RtEstim3_95Lo <- NB2_measles_RtEstim3$lower_bound
NB2_measles_RtEstim3_95Up <- NB2_measles_RtEstim3$upper_bound
NB3_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat3_NB_measles)
NB3_measles_RtEstim3_Rt <- NB3_measles_RtEstim3$fitted_Rt
NB3_measles_RtEstim3_95Lo <- NB3_measles_RtEstim3$lower_bound
NB3_measles_RtEstim3_95Up <- NB3_measles_RtEstim3$upper_bound
NB4_measles_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat4_NB_measles)
NB4_measles_RtEstim3_Rt <- NB4_measles_RtEstim3$fitted_Rt
NB4_measles_RtEstim3_95Lo <- NB4_measles_RtEstim3$lower_bound
NB4_measles_RtEstim3_95Up <- NB4_measles_RtEstim3$upper_bound
Pois1_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat1_pois_SARS)
Pois1_SARS_RtEstim3_Rt <- Pois1_SARS_RtEstim3$fitted_Rt
Pois1_SARS_RtEstim3_95Lo <- Pois1_SARS_RtEstim3$lower_bound
Pois1_SARS_RtEstim3_95Up <- Pois1_SARS_RtEstim3$upper_bound
Pois2_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat2_pois_SARS)
Pois2_SARS_RtEstim3_Rt <- Pois2_SARS_RtEstim3$fitted_Rt
Pois2_SARS_RtEstim3_95Lo <- Pois2_SARS_RtEstim3$lower_bound
Pois2_SARS_RtEstim3_95Up <- Pois2_SARS_RtEstim3$upper_bound
Pois3_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat3_pois_SARS)
Pois3_SARS_RtEstim3_Rt <- Pois3_SARS_RtEstim3$fitted_Rt
Pois3_SARS_RtEstim3_95Lo <- Pois3_SARS_RtEstim3$lower_bound
Pois3_SARS_RtEstim3_95Up <- Pois3_SARS_RtEstim3$upper_bound
Pois4_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat4_pois_SARS)
Pois4_SARS_RtEstim3_Rt <- Pois4_SARS_RtEstim3$fitted_Rt
Pois4_SARS_RtEstim3_95Lo <- Pois4_SARS_RtEstim3$lower_bound
Pois4_SARS_RtEstim3_95Up <- Pois4_SARS_RtEstim3$upper_bound
NB1_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat1_NB_SARS)
NB1_SARS_RtEstim3_Rt <- NB1_SARS_RtEstim3$fitted_Rt
NB1_SARS_RtEstim3_95Lo <- NB1_SARS_RtEstim3$lower_bound
NB1_SARS_RtEstim3_95Up <- NB1_SARS_RtEstim3$upper_bound
NB2_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat2_NB_SARS)
NB2_SARS_RtEstim3_Rt <- NB2_SARS_RtEstim3$fitted_Rt
NB2_SARS_RtEstim3_95Lo <- NB2_SARS_RtEstim3$lower_bound
NB2_SARS_RtEstim3_95Up <- NB2_SARS_RtEstim3$upper_bound
NB3_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat3_NB_SARS)
NB3_SARS_RtEstim3_Rt <- NB3_SARS_RtEstim3$fitted_Rt
NB3_SARS_RtEstim3_95Lo <- NB3_SARS_RtEstim3$lower_bound
NB3_SARS_RtEstim3_95Up <- NB3_SARS_RtEstim3$upper_bound
NB4_SARS_RtEstim3 <- problem_solver(method = 'RtEstim(k=3)', instance = dat4_NB_SARS)
NB4_SARS_RtEstim3_Rt <- NB4_SARS_RtEstim3$fitted_Rt
NB4_SARS_RtEstim3_95Lo <- NB4_SARS_RtEstim3$lower_bound
NB4_SARS_RtEstim3_95Up <- NB4_SARS_RtEstim3$upper_bound

RtEstim3_res <- c(Pois1_measles_RtEstim3_Rt, Pois1_measles_RtEstim3_95Lo, Pois1_measles_RtEstim3_95Up, 
                  Pois2_measles_RtEstim3_Rt, Pois2_measles_RtEstim3_95Lo, Pois2_measles_RtEstim3_95Up, 
                  Pois3_measles_RtEstim3_Rt, Pois3_measles_RtEstim3_95Lo, Pois3_measles_RtEstim3_95Up, 
                  Pois4_measles_RtEstim3_Rt, Pois4_measles_RtEstim3_95Lo, Pois4_measles_RtEstim3_95Up, 
                  NB1_measles_RtEstim3_Rt, NB1_measles_RtEstim3_95Lo, NB1_measles_RtEstim3_95Up, 
                  NB2_measles_RtEstim3_Rt, NB2_measles_RtEstim3_95Lo, NB2_measles_RtEstim3_95Up, 
                  NB3_measles_RtEstim3_Rt, NB3_measles_RtEstim3_95Lo, NB3_measles_RtEstim3_95Up, 
                  NB4_measles_RtEstim3_Rt, NB4_measles_RtEstim3_95Lo, NB4_measles_RtEstim3_95Up,
                  Pois1_SARS_RtEstim3_Rt, Pois1_SARS_RtEstim3_95Lo, Pois1_SARS_RtEstim3_95Up, 
                  Pois2_SARS_RtEstim3_Rt, Pois2_SARS_RtEstim3_95Lo, Pois2_SARS_RtEstim3_95Up, 
                  Pois3_SARS_RtEstim3_Rt, Pois3_SARS_RtEstim3_95Lo, Pois3_SARS_RtEstim3_95Up, 
                  Pois4_SARS_RtEstim3_Rt, Pois4_SARS_RtEstim3_95Lo, Pois4_SARS_RtEstim3_95Up,
                  NB1_SARS_RtEstim3_Rt, NB1_SARS_RtEstim3_95Lo, NB1_SARS_RtEstim3_95Up, 
                  NB2_SARS_RtEstim3_Rt, NB2_SARS_RtEstim3_95Lo, NB2_SARS_RtEstim3_95Up, 
                  NB3_SARS_RtEstim3_Rt, NB3_SARS_RtEstim3_95Lo, NB3_SARS_RtEstim3_95Up, 
                  NB4_SARS_RtEstim3_Rt, NB4_SARS_RtEstim3_95Lo, NB4_SARS_RtEstim3_95Up
)
RtEstim3_res <- data.table(Res = RtEstim3_res,
                           Type = rep(rep(c("Fitted_Rt", "Lo95", "Up95"), each = 300), 16))
RtEstim3_Rt <- RtEstim3_res[Type == "Fitted_Rt", ]$Res
RtEstim3_Lo95 <- RtEstim3_res[Type == "Lo95", ]$Res
RtEstim3_Up95 <- RtEstim3_res[Type == "Up95", ]$Res

# save results ----
res_dat <- rbind(dat_example, dat_example, dat_example, dat_example,
                 dat_example, dat_example, dat_example, dat_example)
res_dat <- data.table(
  res_dat,
  EpiWeek_Rt = Epiweek_Rt, 
  EpiMonth_Rt = Epimonth_Rt,
  EpiLPS_Rt = EpiLPS_Rt,
  EpiFilter_Rt = EpiFilter_Rt,
  RtEstim0_Rt = RtEstim0_Rt,
  RtEstim1_Rt = RtEstim1_Rt,
  RtEstim2_Rt = RtEstim2_Rt,
  RtEstim3_Rt = RtEstim3_Rt,
  EpiWeek_Lo95 = Epiweek_Lo95,
  EpiMonth_Lo95 = Epimonth_Lo95,
  EpiLPS_Lo95 = EpiLPS_Lo95,
  EpiFilter_Lo95 = EpiFilter_Lo95,
  RtEstim0_Lo95 = RtEstim0_Lo95,
  RtEstim1_Lo95 = RtEstim1_Lo95,
  RtEstim2_Lo95 = RtEstim2_Lo95,
  RtEstim3_Lo95 = RtEstim3_Lo95,
  EpiWeek_Up95 = Epiweek_Up95,
  EpiMonth_Up95 = Epimonth_Up95,
  EpiLPS_Up95 = EpiLPS_Up95,
  EpiFilter_Up95 = EpiFilter_Up95,
  RtEstim0_Up95 = RtEstim0_Up95,
  RtEstim1_Up95 = RtEstim1_Up95,
  RtEstim2_Up95 = RtEstim2_Up95,
  RtEstim3_Up95 = RtEstim3_Up95
)
  
saveRDS(res_dat, here::here("dat/data_example_estimates.RDS"))
```

```{r Figs5+6-display-fitted-rt, eval = FALSE, include = FALSE}
#res_dat_long <- res_dat %>%
#  pivot_longer(EpiWeek_Rt:RtEstim3_Rt, names_to = "method", values_to = "Rt_value") %>%
#  mutate(method = fct_recode(method,
#    "EpiEstim (weekly)" = "EpiWeek_Rt",
#    "EpiEstim (monthly)" = "EpiMonth_Rt",
#    "EpiLPS" = "EpiLPS_Rt",
#    "EpiFilter" = "EpiFilter_Rt",
#    "RtEstim (k=0)" = "RtEstim0_Rt",
#    "RtEstim (k=1)" = "RtEstim1_Rt",
#    "RtEstim (k=2)" = "RtEstim2_Rt",
#    "RtEstim (k=3)" = "RtEstim3_Rt"
#  )) %>%
#  mutate(method = fct_relevel(
#    method, 
#    "EpiEstim (weekly)", "EpiEstim (monthly)", "EpiLPS", "EpiFilter", 
#    "RtEstim (k=0)", "RtEstim (k=1)", "RtEstim (k=2)", "RtEstim (k=3)"
#  )) %>%
#  filter(Rt_value > 0)

fig_res_pois_measles <- res_dat_long %>%
  filter(si_type == "measles", dist == "Poisson") %>% 
  #filter(method != "EpiLPS") %>% # EpiLPS is large for the first 14 days
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for measles epidemics with Poisson incidence", x="Time") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  coord_cartesian(ylim = c(0, 4)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_pois_measles)

fig_res_NB_sars <- res_dat_long %>%
  filter(si_type == "SARS", dist == "Negative Binomial") %>% 
  group_by(Rt_case, method) %>%
  ggplot(aes(y = Rt_value, x = time)) + 
  geom_line(aes(y = trueRt)) + 
  geom_line(aes(col = method)) +
  facet_wrap( ~ Rt_case, scales = "free") +
  scale_colour_manual(values = cbPalette[-5]) +
  labs(y = "Rt estimates for SARS epidemics with NB incidence", x="Time") + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 4)) + 
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = NULL, byrow=TRUE))
print(fig_res_NB_sars)

ggsave(plot=fig_res_pois_measles, here::here("fig/fig_res_pois_measles.png"), width = 6.5, height = 5)
ggsave(plot=fig_res_NB_sars, here::here("fig/fig_res_NB_sars.png"), width = 6.5, height = 5)
```

```{r Fig7-display-cancovid-Rt-est, eval=FALSE, include=FALSE}
## estimate Rt using RtEstim
# k=1
cv_mod_covid1 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 1, nfold = 10,
  maxiter = 1e7L, nsol = 50, error_measure = "deviance", 
  lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn
)
rt_ci_covid1 <- confband(cv_mod_covid1, "lambda.1se") # get 95% confidence band

# k=2
cv_mod_covid2 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 2, nfold = 10,
  maxiter = 1e7L, nsol = 50, error_measure = "deviance", 
  lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn
)
rt_ci_covid2 <- confband(cv_mod_covid2, "lambda.1se") # get 95% confidence band

# k=3
cv_mod_covid3 <- cv_estimate_rt(
  x = rtestim::cancovid$date, rtestim::cancovid$incident_cases, korder = 3, nfold = 10,
  maxiter = 1e7L, nsol = 50, error_measure = "deviance", 
  lambda_min_ratio = 1e-6, delay_distn = covid_delay_distn
)
rt_ci_covid3 <- confband(cv_mod_covid3, "lambda.1se") # get 95% confidence band

# Rt values with a candidate set
len <- nrow(rtestim::cancovid) #1253
k <- 50
df_covid_Rt <- data.table(
  Rt = c(as.vector(cv_mod_covid1$full_fit$Rt), as.vector(cv_mod_covid2$full_fit$Rt), as.vector(cv_mod_covid3$full_fit$Rt)),
  lambda = c(rep(cv_mod_covid1$lambda, each = len), rep(cv_mod_covid2$lambda, each = len), rep(cv_mod_covid3$lambda, each = len)),
  Time = rep(rtestim::cancovid$date, k*3),
  degree = rep(c("1","2","3"), each = len*k)
)

fig_covid_Rt <- df_covid_Rt %>%
  mutate(degree = fct_recode(degree,
    "Piecewise linear Rt" = "1",
    "Piecewise quadratic Rt" = "2",
    "Piecewise cubic Rt" = "3"
  )) %>%
  group_by(degree, lambda) %>%
  ggplot(aes(Time, Rt, colour = lambda, group = lambda)) +
  geom_line() +
  facet_wrap(~ degree, scales = "free_y", ncol = 1) + 
  labs(y ="Rt estimates for Covid-19 pandemic in Canada", x = "Year", color = "\u03BB") +
  scale_colour_viridis_c(trans = "log10", breaks = c(1, 1e2L, 1e4L, 1e6L), labels = c(1, expression(10^2), expression(10^4), expression(10^6))) +
  guides(color = guide_colorbar(barwidth = 10, barheight = 1)) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y", expand = expansion(c(0, 0.05))) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  coord_cartesian(ylim = c(0, 2)) + 
  geom_hline(yintercept = 1) +
  theme_bw() +
  theme(legend.position = "bottom")
  
# CV-chosen Rt with CB
rt_ci_covid3 <- as.data.table(rt_ci_covid3) %>% 
  mutate(degree = "3", lambda = cv_mod_covid3$lambda.1se) %>% 
  mutate(Date = rtestim::cancovid$date) 
rt_ci_covid2 <- as.data.table(rt_ci_covid2) %>% 
  mutate(degree = "2", lambda = cv_mod_covid2$lambda.1se) %>% 
  mutate(Date = rtestim::cancovid$date) 
rt_ci_covid1 <- as.data.table(rt_ci_covid1) %>% 
  mutate(degree = "1", lambda = cv_mod_covid1$lambda.1se) %>% 
  mutate(Date = rtestim::cancovid$date) 
df_covid_ci <- rbind(rt_ci_covid1, rt_ci_covid2, rt_ci_covid3)


fig_covid_ci <- df_covid_ci %>% 
  mutate(degree = fct_recode(degree,
    "Piecewise linear Rt" = "1",
    "Piecewise quadratic Rt" = "2",
    "Piecewise cubic Rt" = "3"
  )) %>%
  group_by(degree, lambda) %>%
  ggplot(aes(x = Date)) +
  geom_line(aes(y = Rt)) +
  facet_wrap(~ degree, scales = "free_y", ncol = 1) + 
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`),
    fill = "gray", alpha = 0.5
  ) +
  geom_hline(yintercept = 1, linetype = "dotted") +
  coord_cartesian(ylim = c(0, 2)) + 
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y", expand = expansion(c(0, 0.05))) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  labs(x = "Year", y = "") +
  theme_bw()

CanCovid_Rt_all <- ggpubr::ggarrange(fig_covid_Rt, fig_covid_ci,
                         ncol = 2, nrow = 1, 
                         common.legend = TRUE, legend = "bottom"
)
print(CanCovid_Rt_all)

ggsave(here::here("fig/covid_full_res.png"), CanCovid_Rt_all, width = 8.34, height = 6.19)
```

```{r Fig8-display-flu1918-data, eval=FALSE, include=FALSE}
# get the flu data from `EpiEstim` package
data("Flu1918")
len <- length(Flu1918$incidence)
flu_dat <- data.frame(incidence = Flu1918$incidence, Day = 1:len)
flu_delay_distn <- Flu1918$si_distr

Flu_case <- flu_dat %>%
  ggplot(aes(x = Day)) +
  geom_line(aes(y = incidence)) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  labs(x = "Days since start of epidemic", y = "1918 Influenza daily incidence in Baltimore") +
  theme_bw()
print(Flu_case)

ggsave(here::here("fig/flu_dat.png"), Flu_case, width = 6.5, height = 5)
```

```{r Fig9-display-flu1918-Rt-est, eval=FALSE, include=FALSE}
cv_mod_flu3 <- cv_estimate_rt(
  observed_counts = flu_dat$incidence, korder = 3, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, 
  delay_distn = flu_delay_distn)
rt_ci_flu3 <- confband(cv_mod_flu3, "lambda.1se") 

cv_mod_flu2 <- cv_estimate_rt(
  observed_counts = flu_dat$incidence, korder = 2, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, 
  delay_distn = flu_delay_distn)
rt_ci_flu2 <- confband(cv_mod_flu2, "lambda.1se") 

cv_mod_flu1 <- cv_estimate_rt(
  observed_counts = flu_dat$incidence, korder = 1, nfold = 10, maxiter = 1e7L, 
  nsol = 50, error_measure = "deviance", lambda_min_ratio = 1e-6, 
  delay_distn = flu_delay_distn)
rt_ci_flu1 <- confband(cv_mod_flu1, "lambda.1se") 

# Rt values with a candidate set
k <- length(cv_mod_flu1$lambda)
df_flu_Rt <- data.table(
  Rt = c(as.vector(cv_mod_flu1$full_fit$Rt), as.vector(cv_mod_flu2$full_fit$Rt), as.vector(cv_mod_flu3$full_fit$Rt)),
  lambda = c(rep(cv_mod_flu1$lambda, each=len), rep(cv_mod_flu2$lambda, each=len), rep(cv_mod_flu3$lambda, each=len)),
  Time = rep(1:len, k*3),
  degree = rep(c("1","2","3"), each = len*k)
)

fig_flu_Rt <- df_flu_Rt %>%
  mutate(degree = fct_recode(degree,
    "Piecewise linear Rt" = "1",
    "Piecewise quadratic Rt" = "2",
    "Piecewise cubic Rt" = "3"
  )) %>%
  group_by(degree, lambda) %>%
  ggplot(aes(Time, Rt, colour = lambda, group = lambda)) +
  geom_line() +
  facet_wrap(~ degree, scales = "free_y", ncol = 1) + 
  labs(y ="Rt estimates for 1918 Influenza in Baltimore", x = "Days since start of epidemic", color = "\u03BB") +
  scale_colour_viridis_c(trans = "log10", breaks = c(0.1, 1, 10, 100, 1000), labels = c(0.1, 1, 10, expression(10^2), expression(10^3))) +
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  guides(color = guide_colorbar(barwidth = 10, barheight = 1)) +
  geom_hline(yintercept = 1) +
  theme_bw() +
  theme(legend.position = "bottom")

rt_ci_flu3 <- as.data.table(rt_ci_flu3) %>% 
  mutate(degree = "3", lambda = cv_mod_flu3$lambda.1se) %>% 
  mutate(Day = flu_dat$Day) 
rt_ci_flu2 <- as.data.table(rt_ci_flu2) %>% 
  mutate(degree = "2", lambda = cv_mod_flu2$lambda.1se) %>% 
  mutate(Day = flu_dat$Day) 
rt_ci_flu1 <- as.data.table(rt_ci_flu1) %>% 
  mutate(degree = "1", lambda = cv_mod_flu1$lambda.1se) %>% 
  mutate(Day = flu_dat$Day) 
df_flu_ci <- rbind(rt_ci_flu1, rt_ci_flu2, rt_ci_flu3)

fig_flu_ci <- df_flu_ci %>% 
  mutate(degree = fct_recode(degree,
    "Piecewise linear Rt" = "1",
    "Piecewise quadratic Rt" = "2",
    "Piecewise cubic Rt" = "3"
  )) %>%
  group_by(degree, lambda) %>%
  ggplot(aes(x = Day)) +
  geom_line(aes(y = Rt)) +
  facet_wrap(~ degree, scales = "free_y", ncol = 1) + 
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`),
    fill = "gray", alpha = 0.5
  ) +
  scale_x_continuous(expand = expansion(c(0, 0.05))) + 
  scale_y_continuous(expand = expansion(c(0, 0.05))) + 
  geom_hline(yintercept = 1, linetype = "dotted") +
  labs(x = "Days since start of epidemic", y = "") +
  theme_bw()

Flu_Rt <- ggarrange(fig_flu_Rt, fig_flu_ci,
  ncol = 2, nrow = 1, common.legend = TRUE, legend = "bottom"
)
print(Flu_Rt)

ggsave(here::here("fig/flu_full_res.png"), Flu_Rt, width = 6.5, height = 5)
```

